---
title: "Serology: MFI spaghetti plots"
date: "updated: `r Sys.time()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
---

## File Overview

This file generates spaghetti plots of log10MFI IgG for individual childsets, across all antigen targets.

Important to note:

1.  Childsets are coded PCR/IgG status at the **antigen** level

2.  Childsets are coded conditional on a PCR+ result occurring before the *last occurrence* of an IgG result. This means if the PCR result was positive only at 24 months, the childset will be coded as PCR-. It also means if the only PCR+ result was at or after the last occurrence of an IgG result, the childset will be coded as PCR- (see table below). In other words, to classify children, PCR+ results were required to occur before the final available serology measurement. PCR+ detections observed only at the last serology timepoint were excluded, regardless if they were IgG+ or IgG-, because no subsequent IgG follow-up was available to assess seroconversion.

| PCR status | IgG status | Coded as... |
|------------------------|------------------------|------------------------|
| +, at any timepoint before last IgG result | +, at any timepoint  | PCR+IgG+ |
| +, at any timepoint before last IgG result | -, at all timepoints | PCR+IgG- |
| -, at all timepoints before last IgG result | +, at any timepoint | PCR-IgG+ |
| -, at all timepoints before last IgG result | -, at all timepoints | PCR-IgG- |

: *PCR+ only counts if it occurs strictly before the last recorded IgG timepoint, or if a PCR+ at the last IgG timepoint is accompanied by at least one earlier PCR+ event*

3.  For plotting purposes, if the PCR+ result occurred at a time-point where IgG was not measured (3 or 15 months), or at a timepoint where IgG was not available for that childset -\> the PCR+ result was pushed to the next occurrence of an IgG result in the time series so it would be visualized on the plot.

## System setup

```{r}
# Read in files 
library(here)
source(here("R", "0-config.R"))
library(purrr)
```

## Read in files

```{r}
# seroconversion individual level
sero_conv <- readRDS(here("output/ecomid_sero_conv_tbl.rds"))
sero_conv$house_ID <- str_sub(sero_conv$childid, -4)


# seropositivity individual level
sero_pos <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_2026-05-19.csv")) %>% filter(antigen != "hyle")

# seropositivity individual level, updated cutoffs
sero_pos2 <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_enteric_cutoffs_2026-05-19.csv")) %>% filter(antigen != "hyle")

# pcr individual level
pcr <- readRDS(here("data/ecomid_pcr_analysis_data_2026-06-11.rds")) %>% ungroup() %>% 
  rename(hhid = childid, 
         sampleid = sample_ID) %>% 
  mutate(childid = substr(sampleid, start = 1, stop = 6))
                          

pcr$community_type <- as.character(pcr$community_type)
pcr <- pcr %>% mutate(community_type = ifelse(str_detect(community_type, "Rural"), "Rural", community_type))
pcr$community_type <- factor(pcr$community_type, levels = c("Esmeraldas", "Borbon", "Rural"))
pcr <- pcr %>% filter(childid %in% sero_pos$childid)

```

## 1) PCR/IgG: Coding

Recode the childsets into various PCR, IgG categorizaiton. This is done on the **antigen** level


```{r}
# list of enteric pathogens for analysis
pathogens <- c("Campylobacter", "Cryptosporidium", "E. histolytica", 
               "LT-ETEC", "Giardia", "Norovirus GI", "Norovirus GII", 
               "Shigella/EIEC", "Salmonella" )

# need to code PCR data at antigen level.
pcr2 <- pcr %>% 
  mutate(
    leca      = ifelse(pathogen == "E. histolytica", 1, 0), 
    cp23      = ifelse(pathogen == "Cryptosporidium", 1, 0), 
    cp17      = ifelse(pathogen == "Cryptosporidium", 1, 0),
    norogii6  = ifelse(pathogen == "Norovirus GII", 1, 0), 
    norogii17 = ifelse(pathogen == "Norovirus GII", 1, 0), 
    norogii4  = ifelse(pathogen == "Norovirus GII", 1, 0), 
    norogii3  = ifelse(pathogen == "Norovirus GII", 1, 0), 
    norogi5   = ifelse(pathogen == "Norovirus GI", 1, 0),
    p18       = ifelse(pathogen == "Campylobacter", 1, 0), 
    p39       = ifelse(pathogen == "Campylobacter", 1, 0), 
    vsp5      = ifelse(pathogen == "Giardia", 1, 0), 
    etec      = ifelse(pathogen == "LT-ETEC", 1, 0), 
    ipab      = ifelse(pathogen == "Shigella/EIEC", 1, 0), 
    salb      = ifelse(pathogen == "Salmonella", 1, 0), 
    sald      = ifelse(pathogen == "Salmonella", 1, 0)
  ) %>% 
  filter(pathogen %in% pathogens) %>%
  pivot_longer(
    cols      = c(leca, cp23, cp17, norogii6, norogii17, norogii4, norogii3,
                  norogi5, p18, p39, vsp5, etec, ipab, salb, sald),
    names_to  = "antigen", 
    values_to = "presence"
  ) %>% 
  filter(presence == 1) %>% 
  rename(timepoint = visit)


# make dataframe for all antigen and seropositivity/mfi 
# since this is now on the ANTIGEN level, need to code seropos/pcr agreement on the antigen level over all timepoints 
all_mfi <- sero_pos %>% select(sampleid, childid, timepoint, 
                               pathogen, antigen, log10mfi,
                               mficut, seropos) %>% filter(pathogen %in% pathogens) %>% 
  full_join(pcr2 %>% select(childid, timepoint, pathogen, antigen, pcrpos), by = c("childid", "timepoint", "pathogen", "antigen")) 


# 1) Recode the grouping according to the above table 
# filter out PCR results that are positive and occur only at the last seropos result timepoint 
child_antigen_summary <- all_mfi %>%
  group_by(childid, pathogen, antigen) %>%
  summarise(
    last_sero_tp  = max(timepoint[!is.na(seropos)], na.rm = TRUE),
    any_pcr_before = any(pcrpos == 1 & timepoint < last_sero_tp, na.rm = TRUE),
    any_sero      = any(seropos == 1, na.rm = TRUE),
    assay_pos     = case_when(
       any_pcr_before &  any_sero ~ "PCR+ IgG+",
       any_pcr_before & !any_sero ~ "PCR+ IgG-",
      !any_pcr_before &  any_sero ~ "PCR- IgG+",
      TRUE                        ~ "PCR- IgG-"
    ),
    .groups = "drop"
  ) %>%
  select(childid, pathogen, antigen, assay_pos)


# merge the child level back into the raw data we made earlier 
df <- all_mfi %>% left_join(child_antigen_summary %>% select(childid, pathogen, antigen, assay_pos), 
                               by = c("childid", "pathogen", "antigen"))

# export for further downstream analyses
write.csv(df, "../output/ecomid_seropos_pcr_alltargets.csv")
write_rds(df, "../output/ecomid_seropos_pcr_alltargets.rds")

# summarise table of findings for all pathogens 
t1 <- df %>% 
  group_by(pathogen, antigen, assay_pos) %>% 
  select(childid, pathogen, antigen, assay_pos) %>% 
  unique() %>% 
  summarise(n_children = n(), .groups = "drop") %>% 
  pivot_wider(
    names_from = assay_pos, 
    values_from = n_children,
    values_fill = 0
  ) %>% 
  select("pathogen", "antigen", 
         "PCR+ IgG+", "PCR+ IgG-", 
         "PCR- IgG+", "PCR- IgG-")

kable(t1, caption = "Child-set categorization") %>% kable_styling() %>% 
  footnote(general = "PCR results excluded if only positive result is after or at the last occurence of an IgG result")

# Count of how many child-sets are coded as PCR- but have a PCR+ at the last recorded IgG timepoint 
t2 <- df %>% 
  group_by(pathogen, antigen, childid) %>%
  filter(assay_pos == "PCR- IgG+") %>%
  summarise(
    any_pcrpos = any(pcrpos == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(any_pcrpos) %>%
  group_by(pathogen, antigen) %>%
  summarise(n_children = n(), .groups = "drop")

kable(t2, caption = "Child-set categorized as PCR- due to PCR exclusion criteria") %>% kable_styling() %>%
  footnote(general = "PCR+ results occuring only at or after the last recorded IgG result were excluded")

```

Data clean

```{r, plots for age-based assay groupings, width = 12, height = 9}
# calculate line coloring
# Set antigen-specific thresholds
df <- df %>% mutate(threshold = log10(mficut)) 

# recode the PCR result according to following:
# If PCRPOS == 1 result occurs where there is a NA for seropos result, it shifts the PCRPOS == 1 to the the nearest timepoint where seropos is not NA 
# Otherwise, use original pcrpos == 1 value at the timepoint where there is a seropos result

df <- df %>%
  group_by(childid, antigen) %>%
  arrange(timepoint, .by_group = TRUE) %>%
  group_modify(~{
    dat <- .x
    dat$row_id <- seq_len(nrow(dat))

    # for each row, index of the next row with non-missing seropos
    dat$next_sero_row <- rev(cummin(rev(ifelse(!is.na(dat$seropos), dat$row_id, Inf))))

    dat$pcrpos_adjusted <- dat$pcrpos

    # rows where PCR is positive and the seropos result is NA, move to a future seropos result row
    idx <- which(dat$pcrpos == 1 & is.na(dat$seropos) & is.finite(dat$next_sero_row))

    if (length(idx) > 0) {
      target_rows <- dat$next_sero_row[idx]

      # if multiple PCR+ rows point to the same target, keep the most recent one
      for (tr in unique(target_rows)) {
        src <- max(idx[target_rows == tr])
        dat$pcrpos_adjusted[src] <- 0
        dat$pcrpos_adjusted[tr]  <- 1
      }
    }

    dat
  }) %>%
  ungroup() %>%
  select(-row_id, -next_sero_row)


# Compute next_timepoint, next_log10mfi, next_sero, segment_color
df <- df %>% filter(timepoint != 3 & timepoint != 15, !is.na(seropos)) %>% 
    arrange(childid, pathogen, antigen, timepoint) %>%
    group_by(childid, pathogen, antigen) %>%
    mutate(
      next_timepoint = lead(timepoint),
      next_log10mfi = lead(log10mfi),
      next_sero = lead(seropos)
    ) %>%
    ungroup() %>%
    mutate(
      segment_color = case_when(
        is.na(log10mfi) | is.na(next_log10mfi) ~ NA_character_,
        log10mfi < threshold  & next_log10mfi > threshold ~ "Seroconversion / Sero+",
        log10mfi > threshold & next_log10mfi > threshold ~ "Seroconversion / Sero+",
        log10mfi > threshold & next_log10mfi < threshold  ~ "Seroreversion",
        log10mfi < threshold  & next_log10mfi < threshold  ~ "Sero-"
      )
    ) 


# add in age in days
df_days <- df %>% left_join(sero_conv %>% select(childid, timepoint, pathogen, agedays_best), 
                            by = c("childid", "timepoint", "pathogen")) %>% 
 # filter(!is.na(seropos), 
  #       !is.na(agedays_best)) %>%
    arrange(childid, agedays_best) %>%
    group_by(childid, pathogen, antigen) %>%
    mutate(
      agemonths_best = agedays_best/30.417,
      next_age_days = lead(agedays_best),
      next_age_months = lead(agemonths_best),
    ) %>%
    ungroup()

# add levels for the assays category for plot panels 
df_days$assay_pos <- factor(df_days$assay_pos, levels = 
                              c("PCR+ IgG-", "PCR+ IgG+", "PCR- IgG+"))

# remove the neg-neg group for plots
df_days <- df_days %>% filter(assay_pos != "PCR- IgG-") 

# make a long format of the table so can insert N values to plots 
t1_long <- t1 %>%
  pivot_longer(
    cols = c(`PCR+ IgG+`, `PCR+ IgG-`, `PCR- IgG+`, `PCR- IgG-`),
    names_to = "assay_pos",
    values_to = "N"
  ) %>%
  mutate(
    label = paste0("N=", N)
  ) %>% filter(assay_pos != "PCR- IgG-")

t1_long$assay_pos <- factor(t1_long$assay_pos , levels = c("PCR+ IgG-", "PCR+ IgG+", "PCR- IgG+"))

```

## 2) Spaghetti Plots

Spaghetti plots are built individually for each antigen to account for more flexibility in aggregate plot building.

### A) Bacteria

#### Campylobacter (p18, p39)

```{r}
# Limit to one antigen per pathogen
#--------------------------------------#
# Campy; p18
#--------------------------------------#
df_campy_p18 <- df_days %>% filter(antigen == "p18")
first_assay <- unique(df_campy_p18$assay_pos)[1] # used for plot wrapping purposes 

p_campy_p18 <- ggplot(df_campy_p18, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_campy_p18 %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_campy_p18$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",  
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Campylobacter, p18"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
 geom_text(
  data = t1_long %>% filter(antigen == "p18"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),
        strip.background = element_rect(fill = "lightblue", color = "lightblue"))

p_campy_p18

#--------------------------------------#
# Campy; p39
#--------------------------------------#

df_campy_p39 <- df_days %>% filter(antigen == "p39")
p_campy_p39 <- ggplot(df_campy_p39, 
                      aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE
  ) + 
  
  geom_point(
    data = df_campy_p39 %>%
      filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
    shape = 21,
    fill = "red",
    color = "red", 
    size = 0.4
  ) +
  
  geom_hline(
    yintercept = df_campy_p39$threshold,
    linetype = "dashed",
    size = 0.25
  ) +
  
  scale_color_manual(
    name = "",
    values = c(
      "Seroconversion / Sero+" = "#F2A900",
      "Seroreversion" = "#08306B",
      "Sero-" = "grey70"
    ),
    na.translate = FALSE
  ) +
  
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  ) +
  
  labs(
    x = NULL,
    y = NULL,
    title = NULL
  ) +
  
  scale_x_continuous(
    limits = c(6, 24), 
    breaks = seq(0, 24, by = 3)
  ) + 
  scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  facet_grid(~ assay_pos) + 
  
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Campylobacter, p39"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) +
  geom_text(
  data = t1_long %>% filter(antigen == "p39"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) +
  coord_cartesian(ylim = c(0, 5), clip = "off") +
  
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2),
    axis.title.y = element_markdown(), 
    strip.background = element_rect(fill = "lightblue", color = "lightblue")
  )

p_campy_p39

```

#### ETEC

```{r}
# Limit to one antigen per pathogen

#--------------------------------------#
# LT-ETEC
#--------------------------------------#
df_etec <- df_days %>% filter(antigen == "etec")
p_etec <- ggplot(df_etec, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
  geom_point(
    data = df_etec %>%
      filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
    shape = 21,
    fill = "red",
    color = "red", 
    size = 0.4) +
  
    geom_hline(yintercept = df_etec$threshold, linetype = "dashed", size = 0.25) +
    
    scale_color_manual(
      name = "",
      values = c(
        "Seroconversion / Sero+" = "#F2A900",
        "Seroreversion" = "#08306B",
        "Sero-" = "grey70"
      ),
      na.translate = FALSE
    ) +
  
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none", #c(0.15,0.9),
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) +
  
  labs(x = NULL, 
      y = "log10 Luminex IgG (MFI-bg)",
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
       breaks = seq(0, 24, by = 3)
    )  + 
  scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) +
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos)  + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "LT-ETEC, LTB subunit"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + geom_text(
  data = t1_long %>% filter(antigen == "etec"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),
        strip.text = element_blank(), 
        strip.background = element_blank())
        #strip.text = element_blank(), 
        #strip.background = element_rect(fill = "lightblue", color = "lightblue")) 

p_etec

```

#### Shigella/EIEC

```{r}
# Limit to one antigen per pathogen
#--------------------------------------#
# Shigella/EIEC
#--------------------------------------#
df_shig <- df_days %>% filter(antigen == "ipab")
p_shig <- ggplot(df_shig, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
  geom_point(
    data = df_shig %>%
      filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
    shape = 21,
    fill = "red",
    color = "red", 
    size = 0.4) +
  
    geom_hline(yintercept = df_shig$threshold, linetype = "dashed", size = 0.25) +
    
    scale_color_manual(
      name = "",
      values = c(
        "Seroconversion / Sero+" = "#F2A900",
        "Seroreversion" = "#08306B",
        "Sero-" = "grey70"
      ),
      na.translate = FALSE
    ) +
  
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) +
  
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
       breaks = seq(0, 24, by = 3)
    )  + 
  scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) +
  coord_cartesian(ylim=c(0,5))  +
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Shigella/EIEC, IpaB"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + geom_text(
  data = t1_long %>% filter(antigen == "ipab"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  
  theme(axis.title.y = element_markdown(), 
        strip.text = element_blank(), 
        strip.background = element_blank())

p_shig
```

#### Salmonella (LPS-B, LPS-D)

```{r}
# Limit to one antigen per pathogen

#--------------------------------------#
# Salmonella LPS-B
#--------------------------------------#
df_sal_lpsb <- df_days %>% filter(antigen == "salb")
p_sal_lpsb <- ggplot(df_sal_lpsb, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
  geom_point(
  data = df_sal_lpsb %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
guides(
  color = guide_legend(
    override.aes = list(
      linetype = c(1, 1, 1, 0),
      shape = c(NA, NA, NA, 20),
      size = c(NA, NA, NA, 2), 
      linewidth = c(1, 1, 1, NA)
    )
  )
) + 
  
    geom_hline(yintercept = df_sal_lpsb$threshold, linetype = "dashed", size = 0.25) + 
  
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "bottom") +  
  #axis.text.x = element_blank(),
 # axis.ticks.x = element_blank(),
  #axis.title.x = element_blank()) + 
  
  labs(x = " ",
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
       breaks = seq(0, 24, by = 3)
    )  +
  scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  coord_cartesian(clip = "off", ylim=c(0,5)) +
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos)  + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Salmonella, LPS-B"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 0, 
    vjust = 1.5, 
    size = 5
  ) + 
  geom_text(
  data = t1_long %>% filter(antigen == "salb"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.05,   
  vjust = 2.5,
  size = 3
) + 
  theme(axis.title.y = element_markdown(),  
        strip.text = element_blank(), 
        strip.background = element_blank())

p_sal_lpsb

#--------------------------------------#
# Salmonella LPS-D
#--------------------------------------#
df_sal_lpsd <- df_days %>% filter(antigen == "sald")
p_sal_lpsd <- ggplot(df_sal_lpsd, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
  geom_point(
    data = df_sal_lpsd %>%
      filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
    shape = 21,
    fill = "red",
    color = "red", 
    size = 0.4) +
  
    geom_hline(yintercept = df_sal_lpsd$threshold, linetype = "dashed", size = 0.25) +
    
    scale_color_manual(
      name = "",
      values = c(
        "Seroconversion / Sero+" = "#F2A900",
        "Seroreversion" = "#08306B",
        "Sero-" = "grey70"
      ),
      na.translate = FALSE
    ) +
  
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none") + 
  #axis.text.x = element_blank(),
  #axis.ticks.x = element_blank(),
  #axis.title.x = element_blank()) + 
  labs(x = " ",
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
       breaks = seq(0, 24, by = 3)
    )  +
  scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) +
  coord_cartesian(clip = "off", ylim=c(0,5)) +
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Salmonella, LPS-D"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
  geom_text(
  data = t1_long %>% filter(antigen == "sald"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  theme(axis.title.y = element_markdown(),  
        strip.text = element_blank(), 
        strip.background = element_blank())

p_sal_lpsd
```

#### Aggregate plot- Bacteria

```{r}
##-----------------------###
## Aggregate plot, bacteria ##
##-----------------------###

bact <- (p_campy_p18 + p_campy_p39) / 
  (p_etec + p_shig) / 
  (p_sal_lpsb + p_sal_lpsd) + 
  theme(plot.margin = margin(0, 0, 0, 0)) + 
  plot_annotation( caption = "Age (months)" ) & 
  theme( plot.caption = element_text(hjust = 0.5, vjust = -0.1, size = 14, margin = margin(t = -25)) )

bact

ggsave(here("figures/aggregate_spaghetti_plot_bacteria.tiff"), bact, device="tiff", height = 230, width = 410, units = "mm")
```

### B) Protozoa

#### Cryptosporidium (cp17, cp23)

```{r}
#--------------------------------------#
# Cryptosporidium, cp17
#--------------------------------------#
df_cryp_cp17 <- df_days %>% filter(antigen == "cp17")

p_cryp_cp17 <- ggplot(df_cryp_cp17, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_cryp_cp17 %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_cryp_cp17$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",  
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = "log10 Luminex IgG (MFI-bg)", # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Cryptosporidium, Cp17"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
 geom_text(
  data = t1_long %>% filter(antigen == "cp17"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),
        strip.background = element_rect(fill = "lightblue", color = "lightblue"))
p_cryp_cp17 



#--------------------------------------#
# Cryptosporidium, cp23
#--------------------------------------#
df_cryp_cp23 <- df_days %>% filter(antigen == "cp23")

p_cryp_cp23 <- ggplot(df_cryp_cp23, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_cryp_cp23 %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_cryp_cp23$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",  
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Cryptosporidium, Cp23"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
 geom_text(
  data = t1_long %>% filter(antigen == "cp23"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),
        strip.background = element_rect(fill = "lightblue", color = "lightblue"))
p_cryp_cp23

```

#### Giardia

```{r}
#--------------------------------------#
# Giarida
#--------------------------------------#
df_giardia <- df_days %>% filter(antigen == "vsp5")

p_giardia <- ggplot(df_giardia, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_giardia %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_giardia$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
guides(
  color = guide_legend(
    override.aes = list(
      linetype = c(1, 1, 1, 0),
      shape = c(NA, NA, NA, 20),
      size = c(NA, NA, NA, 2), 
      linewidth = c(1, 1, 1, NA)
    )
  )
)  +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "bottom") + 
  #axis.text.x = element_blank(),
  #axis.ticks.x = element_blank(),
  #axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Giardia, VSP-5"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
 geom_text(
  data = t1_long %>% filter(antigen == "vsp5"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),  
        strip.text = element_blank(), 
        strip.background = element_blank())
p_giardia

```

#### E. histolytica

```{r}
#--------------------------------------#
# E. histolytica
#--------------------------------------#
df_ehist <- df_days %>% filter(antigen == "leca")

p_ehist <- ggplot(df_ehist, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_ehist %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_ehist$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none") +  
  #axis.text.x = element_blank(),
  #axis.ticks.x = element_blank(),
  #axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "E. histolytica, Lec-A"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
 geom_text(
  data = t1_long %>% filter(antigen == "leca"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),  
        strip.text = element_blank(), 
        strip.background = element_blank())
p_ehist
```

#### Aggregate plot- Protozoa

```{r}
##-----------------------###
## Aggrate plot, Protozoa ##
##-----------------------###

prot <- (p_cryp_cp17 + p_cryp_cp23) / 
  (p_giardia + p_ehist)  + 
  theme(plot.margin = margin(0, 0, 0, 0)) + 
  plot_annotation( caption = "Age (months)" ) & 
  theme( plot.caption = element_text(hjust = 0.5, vjust = -0.1, size = 14, margin = margin(t = -25)) )

prot

ggsave(here("figures/aggregate_spaghetti_plot_protozoa.tiff"), prot, device="tiff", height = 190, width = 410, units = "mm")

```

### C) Norovirus

#### Noro GI.5

```{r}
#--------------------------------------#
# Norovirus, GI
#--------------------------------------#
df_noro1 <- df_days %>% filter(antigen == "norogi5")
p_noro1 <- ggplot(df_noro1, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_noro1 %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_noro1$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "bottom") +
  #axis.text.x = element_blank(),
  #axis.ticks.x = element_blank(),
  #axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Norovirus GI.5"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
 geom_text(
  data = t1_long %>% filter(antigen == "norogi5"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),
        strip.background = element_blank(), 
        strip.text = element_blank())

p_noro1

```

#### Noro GII (3, 4, 6, 17)

```{r}
#--------------------------------------#
# Norovirus GII.3
#--------------------------------------#
df_noro2_3 <- df_days %>% filter(antigen == "norogii3")
p_noro2_3 <- ggplot(df_noro2_3, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_noro2_3 %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_noro2_3$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",  
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Norovirus, GII.3"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
 geom_text(
  data = t1_long %>% filter(antigen == "norogii3"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),
        strip.background = element_rect(fill = "lightblue", color = "lightblue"))

p_noro2_3

#--------------------------------------#
# Norovirus GII.4
#--------------------------------------#
df_noro2_4 <- df_days %>% filter(antigen == "norogii4")
p_noro2_4 <- ggplot(df_noro2_4, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_noro2_4 %>%
    filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_noro2_4$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",  
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Norovirus, GII.4"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + 
 geom_text(
  data = t1_long %>% filter(antigen == "norogii4"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5), clip = "off") +
  theme(axis.title.y = element_markdown(),
        strip.background = element_rect(fill = "lightblue", color = "lightblue"))

p_noro2_4

#--------------------------------------#
# Norovirus GII.6
#--------------------------------------#
df_noro2_6 <- df_days %>% filter(antigen == "norogii6")
p_noro2_6 <- ggplot(df_noro2_6, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
  geom_point(
    data = df_noro2_6 %>%
      filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
    shape = 21,
    fill = "red",
    color = "red", 
    size = 0.4) +
  
    geom_hline(yintercept = df_noro2_6$threshold, linetype = "dashed", size = 0.25) +
    
    scale_color_manual(
      name = "",
      values = c(
        "Seroconversion / Sero+" = "#F2A900",
        "Seroreversion" = "#08306B",
        "Sero-" = "grey70"
      ),
      na.translate = FALSE
    ) +
  
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) +
  
  labs(x = NULL,
    y = "log10 Luminex IgG (MFI-bg)", # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
       breaks = seq(0, 24, by = 3)
    )  + 
  scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) +
  coord_cartesian(ylim=c(0,5))  +
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Norovirus, GII.6"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + geom_text(
  data = t1_long %>% filter(antigen == "norogii6"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  
  theme(axis.title.y = element_markdown(), 
        strip.text = element_blank(), 
        strip.background = element_blank())

p_noro2_6

#--------------------------------------#
# Norovirus GII.17
#--------------------------------------#
df_noro2_17 <- df_days %>% filter(antigen == "norogii17")
p_noro2_17 <- ggplot(df_noro2_17, 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
  geom_point(
    data = df_noro2_17 %>%
      filter(pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
    shape = 21,
    fill = "red",
    color = "red", 
    size = 0.4) +
  
    geom_hline(yintercept = df_noro2_17$threshold, linetype = "dashed", size = 0.25) +
    
    scale_color_manual(
      name = "",
      values = c(
        "Seroconversion / Sero+" = "#F2A900",
        "Seroreversion" = "#08306B",
        "Sero-" = "grey70"
      ),
      na.translate = FALSE
    ) +
  
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none")+ 
  #axis.text.x = element_blank(),
  #axis.ticks.x = element_blank(),
  #axis.title.x = element_blank()) +
  
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
       breaks = seq(0, 24, by = 3)
    )  + 
  scale_y_continuous(limits = c(0, 5), 
       breaks = seq(0, 4, by = 1)
    ) +
  coord_cartesian(ylim=c(0,5))  +
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Norovirus, GII.17"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 5
  ) + geom_text(
  data = t1_long %>% filter(antigen == "norogii17"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = Inf,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  
  theme(axis.title.y = element_markdown(), 
        strip.text = element_blank(), 
        strip.background = element_blank())


p_noro2_17

```

#### Aggregate plot- Virus

```{r}
##-----------------------###
## Aggrate plot, Virus    ##
##-----------------------###

vir <- (p_noro2_3 + p_noro2_4) / 
  (p_noro2_6 + p_noro2_17) / 
  (p_noro1 + plot_spacer())  + 
  theme(plot.margin = margin(0, 0, 0, 0)) + 
  plot_annotation( caption = "Age (months)" ) & 
  theme( plot.caption = element_text(hjust = 0.5, vjust = -0.1, size = 14, margin = margin(t = -25)) )

vir

ggsave(here("figures/aggregate_spaghetti_plot_virus.tiff"), vir, device="tiff", height = 230, width = 410, units = "mm")

```

### D) Main text figure

Assemble a 90 mm plot of 3-4 targets as a main text figure. We chose Campylobacter p39, shigella ipaB, Norovirus GII.4, and Cryptosporidium, cp17 as representative targets because they demonstrate the variety of pcr/igg patterns

```{r}
#---------------------#
# Reset plot settings #
#---------------------#

# create a new count dataframe for just the two panes
t1_main <- t1_long %>% filter(assay_pos != "PCR+ IgG-")
t1_main$assay_pos <- factor(t1_main$assay_pos , levels = c("PCR+ IgG+", "PCR- IgG+"))

# reset the first assay label for the facet panel plotting
first_assay <- unique(t1_main$assay_pos)[1] 

```

Updated plots\
Campy

```{r}
p_campy_p39_2 <- ggplot(df_campy_p39 %>% filter(assay_pos != "PCR+ IgG-"), 
                      aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE
  ) + 
  geom_point(
    data = df_campy_p39 %>%
      filter(assay_pos != "PCR+ IgG-", 
             pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
    shape = 21,
    fill = "red",
    color = "red", 
    size = 0.4
  ) +
  
  geom_hline(
    yintercept = df_campy_p39$threshold,
    linetype = "dashed",
    size = 0.25
  ) +
  
  scale_color_manual(
    name = "",
    values = c(
      "Seroconversion / Sero+" = "#F2A900",
      "Seroreversion" = "#08306B",
      "Sero-" = "grey70"
    ),
    na.translate = FALSE
  ) +
  
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  ) +
  
  labs(
    x = NULL,
    y = NULL,
    title = NULL
  ) +
  
  scale_x_continuous(
    limits = c(6, 24), 
    breaks = seq(0, 24, by = 3)
  ) + 
  scale_y_continuous(limits = c(0, 5.2), 
       breaks = seq(0, 4, by = 1)
    ) + 
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = 6.0,
      y = Inf,
      assay_pos = first_assay,
      label = "Campylobacter, p39"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 3.5
  ) +
  geom_text(
  data = t1_main %>% filter(antigen == "p39"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = 1,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) +
  coord_cartesian(ylim = c(0, 5.2), clip = "off") +
  
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2),
    axis.title.y = element_markdown(), 
    strip.background = element_rect(fill = "lightblue", color = "lightblue")
  )

p_campy_p39_2
```

Crypto

```{r}
# CRYPTO #
# update crypto to remove the legend at the top, keep the y-axis label 
p_cryp_cp17_2 <- ggplot(df_cryp_cp17 %>% filter(assay_pos != "PCR+ IgG-"), 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_cryp_cp17 %>%
    filter(assay_pos != "PCR+ IgG-", 
           pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_cryp_cp17$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",  
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = "log10 Luminex IgG (MFI-bg)", # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5.2), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = 6.0,
      y = Inf,
      assay_pos = first_assay,
      label = "Cryptosporidium, Cp17"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 3.5
  ) + 
 geom_text(
  data = t1_main %>% filter(antigen == "cp17"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = 1,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5.2), clip = "off") +
  theme(axis.title.y = element_markdown(),
        strip.text = element_blank(), 
        strip.background = element_blank())
p_cryp_cp17_2
```

Noro

```{r}
# NORO #

# update noro GII.4 to remove the legend at the top, keep the y-axis label 
p_noro2_4_2 <- ggplot(df_noro2_4 %>% filter(assay_pos != "PCR+ IgG-"), 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
 geom_point(
  data = df_noro2_4 %>%
    filter(assay_pos != "PCR+ IgG-", 
           pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_noro2_4$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
  legend.position = "none",  
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()) + 
  labs(x = NULL,
    y = NULL, # Change to log10(MFI) for last panel 
    title = NULL) +
  scale_x_continuous(limits = c(6, 24), 
      breaks = seq(0, 24, by = 3)
    )  + scale_y_continuous(limits = c(0, 5.2), 
       breaks = seq(0, 4, by = 1)
    ) + 
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  geom_text(
    data = data.frame(
      x = 6.0,
      y = Inf,
      assay_pos = first_assay,
      label = "Norovirus, GII.4"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 3.5
  ) + 
 geom_text(
  data = t1_main %>% filter(antigen == "norogii4"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = 1,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  coord_cartesian(ylim=c(0,5.2), clip = "off") +
  theme(axis.title.y = element_markdown(),
    strip.text = element_blank(),
    strip.background = element_blank(),
    legend.position = "none"
  )
p_noro2_4_2
```

Shigella

```{r}
# SHIGELLA #

# update p_shig to include x-axis (Just rerun code because difficult to add back in)
p_shig_2 <- ggplot(df_shig %>% filter(assay_pos != "PCR+ IgG-"), 
                  aes(x = agemonths_best, y = log10mfi, group = childid)) + 
  
  geom_segment(
    aes(xend = next_age_months, yend = next_log10mfi, color = segment_color),
    linewidth = 0.25, alpha = 0.4, na.rm = TRUE) + 
  
  geom_point(
  data = df_shig %>%
    filter(assay_pos != "PCR+ IgG-", 
           pcrpos_adjusted == 1,
             timepoint != 24, 
             !is.na(next_age_months)),
  aes(color = "PCR+"),
  shape = 16,
  size = 0.6
) +
  geom_hline(yintercept = df_shig$threshold, linetype = "dashed", size = 0.25) + # threshold for specific antigen
  
scale_color_manual(
  name = NULL,
  breaks = c("Seroconversion / Sero+", "Seroreversion", "Sero-", "PCR+"),
  values = c(
    "Seroconversion / Sero+" = "#F2A900",
    "Seroreversion" = "#08306B",
    "Sero-" = "grey70",
    "PCR+" = "red"
  ),
  na.translate = FALSE
) +
guides(
  color = guide_legend(
    nrow = 2, 
    byrow = TRUE, 
    override.aes = list(
      linetype = c(1, 1, 1, 0),
      shape = c(NA, NA, NA, 20),
      size = c(NA, NA, NA, 2), 
      linewidth = c(1, 1, 1, NA)
    )
  )
) +
  theme_minimal() +
  theme(
  panel.grid = element_blank(),
    legend.position = "bottom",
    legend.justification = "left") + 
  #axis.text.x = element_blank(),
  #axis.ticks.x = element_blank(),
  #axis.title.x = element_blank()) +
  scale_x_continuous(limits = c(6, 24), 
       breaks = seq(0, 24, by = 3)
    )  + 
  scale_y_continuous(limits = c(0, 5.2), 
       breaks = seq(0, 4, by = 1)
    ) +
  coord_cartesian(ylim=c(0,5.2), clip = "off")  +
  theme(
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2)
  ) +
  facet_grid(~ assay_pos) + 
  labs(x = "Age (months)", 
       y = NULL) + 
  geom_text(
    data = data.frame(
      x = -Inf,
      y = Inf,
      assay_pos = first_assay,
      label = "Shigella/EIEC, IpaB"
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust =0, 
    vjust = 1.5, 
    size = 3.5
  ) + geom_text(
  data = t1_main %>% filter(antigen == "ipab"), #need to adjust per antigen plot
  aes(
    x = Inf,
    y = 1,
    label = label
  ),
  inherit.aes = FALSE,
   hjust = 1.1,   
  vjust = 2.5,
  size = 3
) + 
  theme(axis.title.y = element_markdown(), 
        strip.text = element_blank(), 
        strip.background = element_blank(), 
         legend.spacing.y = unit(-2, "pt"), 
        legend.box.margin = margin(t = -5),  # negative top margin pulls it upward
        legend.key.height = unit(8, "pt"), 
        axis.line.x = element_line(color = "grey", linewidth = 0.3), 
        axis.ticks.x = element_line(color = "grey", linewidth = 0.3)) 

p_shig_2

```

Aggregate figure

```{r}
# make plot 
agg_maintext <-
  (p_campy_p39_2 / p_cryp_cp17_2 / p_noro2_4_2 / p_shig_2)  

agg_maintext

# save
ggsave(here("figures/aggregate_spaghetti_plot_maintext.tiff"), agg_maintext, device="tiff", height = 190, width = 99, units = "mm")
```
