---
title: "Serology and PCR: summary-measure dot plots and Spearman correlations"
subtitle: "Objective 3 of the PCR-vs-serology cleanup"
author: "Nikolina Walas (nikolina.walas@ucsf.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute:
  echo: true
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

## File overview

This file generates the pathogen x community-type dot-plot correlations.

Four comparisons are produced:

-   **(A) SCR vs. PCR prevalence (pooled).** Giardia excluded from the Spearman ρ shown on
    the plot; Giardia points are still plotted, and the plot is
    annotated to say so. The Giardia-included ρ is computed and printed
    separately (not plotted). Used as the first (leftmost) panel of the
    combined figure below.
-   **(B) SCR vs. PCR Kaplan-Meier cumulative incidence at 24 months.**
    Same Giardia handling as (A). Used as the second panel of the
    combined figure below. Its confidence interval now uses the same
    matched-sample bootstrap as A, C, and D (see below), rather than
    the row-level bootstrap `10-ecomid_dotplot_correlations.qmd` uses
    for this panel.
-   **(C) SCR vs. seroprevalence (pooled).** Giardia included in ρ.
    Used as the third (rightmost) panel of the combined figure below.
-   **(D) SCR vs. seroreversion rate (SRR).** Giardia included in ρ.
    Standalone only -- not part of the combined figure below -- but
    additionally exported as a saved ggplot object
    (`output/dotplot_scr_vs_srr.rds`) so it can be read into a
    separate R file for a supplemental figure there.

A, B, C, and D use a matched-sample bootstrap (resample 370 children with
replacement, recompute pathogen-level summaries, take Spearman ρ across
pathogens) described in CLAUDE.md's Statistical Conventions and
originally implemented in `9-ecomid_sero_pcr_comp.qmd`. D resamples from
the union of `sero_conv` and `sero_rev`.

B resamples children and refitting a Kaplan-Meier curve on each replicate. The
output file used is generated in 
`7-ecomid_cumulative_incidence_data.qmd` (exported as
`ecomid_km_intervals_pcr_by_community.rds`) and just resampled by
`childid` here, the same way A/C/D already resample `sero_conv`/
`pcr_all`. Only the `survfit()` refit itself runs inside the loop, and
reading cumulative incidence at 24 months via `summary(fit, times = 24,
extend = TRUE)` (instead of `broom::tidy()` + the `km_at_rounds()`
forward-fill grid used elsewhere in this workflow) keeps each
replicate's per-pathogen extraction lean.

A combined 3-panel figure (single row) is exported. Its three panels
are document panels A, B, and C, left to right, in that order -- the
combined figure's panel order matches the document's panel order
exactly, so no separate composite lettering is needed. D (SCR vs. SRR)
is not part of the combined figure; it remains standalone, and is
additionally exported on its own as a saved ggplot object (see its
bullet above).


### Load packages

```{r}
library(here)
source(here("R", "0-config.R"))
library(survival)
```

### Read in files

```{r}
# --- Community-level, pre-computed summary tables (population-level; same
#     source as 9-ecomid_sero_pcr_summary.qmd). "Borbon" is relabeled to the
#     accented "Borbón" immediately on read-in, to match comm_n_lookup and
#     the plot legend labels below -- without this, Borbón rows would fail
#     to match comm_n_lookup and silently drop out of the size legend. ---
sero_inc_type <- readRDS(here("output/ecomid_pathogen_comm_sero_inc_3_levels.rds")) %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium") %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
sero_inc_type$community_type2 <- factor(sero_inc_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

sero_prev_type <- readRDS(here("output/ecomid_sero_prev_rates_3_levels.rds")) %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
sero_prev_type$community_type2 <- factor(sero_prev_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

pcr_prev_type <- readRDS(here("output/ecomid_pcr_prev_by_community_3_levels.rds")) %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
pcr_prev_type$community_type2 <- factor(pcr_prev_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

sero_rev_type <- readRDS(here("output/ecomid_pathogen_comm_sero_rev_3_levels.rds")) %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium") %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
sero_rev_type$community_type2 <- factor(sero_rev_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

# --- Individual-level data, needed for the matched-sample bootstrap (A, C, D) ---
# "Borbon" relabeled to accented "Borbón" here too, for the same reason as
# pcr_all/sero_all below -- confirmed against the real data that sero_conv
# and sero_rev carry the raw unaccented "Borbon". Without this, every join
# against pcr_all/sero_all by community_type (comparisons A and C) would
# silently drop all Borbón rows from those bootstraps.
sero_conv <- readRDS(here("output/ecomid_sero_conv_tbl.rds")) %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium") %>%
  mutate(community_type = as.character(community_type),
         community_type = ifelse(community_type == "Borbon", "Borbón", community_type))

sero_rev <- readRDS(here("output/ecomid_sero_rev_tbl.rds")) %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium") %>%
  mutate(community_type = as.character(community_type),
         community_type = ifelse(community_type == "Borbon", "Borbón", community_type))

pcr_all <- readRDS(here("data","ecomid_pcr_analysis_data_2026-08-12.rds"))
pcr_all$community_type <- as.character(pcr_all$community_type)
pcr_all <- pcr_all %>% mutate(community_type = ifelse(str_detect(community_type, "Rural"), "Rural", community_type))
pcr_all <- pcr_all %>% mutate(community_type = ifelse(community_type == "Borbon", "Borbón", community_type))
pcr_all$community_type <- factor(pcr_all$community_type, levels = c("Esmeraldas", "Borbón", "Rural"))
pcr_all <- pcr_all %>% filter(pathogen %in% sero_conv$pathogen)

sero_all <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_2026-08-26.csv")) %>%
  mutate(community_type = ifelse(str_detect(community_type, "Rural"), "Rural", community_type)) %>%
  mutate(community_type = ifelse(community_type == "Borbon", "Borbón", community_type)) %>%
  filter(antigen_lab != "S. enterica HlyE") %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium")

# --- PCR KM cumulative incidence by community, every nominal round -- for B ---
# Exported by 7-ecomid_cumulative_incidence_data.qmd; relabeled to accented
# "Borbón" here for the same reason as the tables above.
km_pcr_by_comm_round <- readRDS(here("output/ecomid_km_pcr_by_community_round.rds")) %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2)) %>% filter(time == 24) %>% filter(pathogen != "S. enterica")

# --- Per-child PCR Kaplan-Meier survival intervals, by community -- for B's
#     matched-sample bootstrap (section 2 below). Exported by
#     7-ecomid_cumulative_incidence_data.qmd as km_data_by_comm: one row per
#     child x pathogen x community_type2, giving that child's own
#     (tstart, tstop, event) interval. A child's interval doesn't depend on
#     which other children land in a given bootstrap resample, so it's built
#     once upstream in file 7 rather than rebuilt inside every one of B's
#     2000 replicates -- that per-child rebuild is what made an earlier
#     attempt at this too slow to run in practice. Relabeled to accented
#     "Borbón" and renamed to plain "community_type" to match sero_conv's
#     join key (same treatment as pcr_all above); S. enterica dropped here
#     too, matching km_pcr_by_comm_round's exclusion above, so B's
#     confidence interval is computed over the same pathogen set as its plot.
km_data_pcr_by_comm <- readRDS(here("output/ecomid_km_intervals_pcr_by_community.rds")) %>%
  mutate(community_type = ifelse(community_type2 == "Borbon", "Borbón", community_type2)) %>%
  filter(pathogen != "S. enterica") %>%
  select(childid, pathogen, community_type, tstart, tstop, event)
```

# 1) Build pathogen x community-type comparison data-frames (A, C, D)

```{r}
d1 <- sero_prev_type %>%
  select(pathogen, community_type2, sero_prevalence, lower_ci, upper_ci, n_total) %>%
  filter(pathogen %in% pcr_prev_type$pathogen)

d1b <- sero_inc_type %>%
  select(pathogen, community_type2, ir_per_100) %>%
  filter(pathogen %in% pcr_prev_type$pathogen)

d2 <- pcr_prev_type %>%
  select(pathogen, community_type2, pcr_prevalence, lower_ci, upper_ci, n_total) %>%
  rename(lower_ci_pcr = lower_ci,
         upper_ci_pcr = upper_ci,
         n_total_pcr = n_total) %>%
  filter(pathogen %in% d1$pathogen)

pathogen_levels <- c("Campylobacter", "Cryptosporidium", "E. histolytica", "Giardia",
                      "LT-ETEC", "Norovirus GI", "Norovirus GII", "S. enterica", "Shigella/EIEC")

comm_n_lookup <- c(Esmeraldas = 85, "Borbón" = 108, Rural = 177)

# --- A) SCR (pooled) + PCR prevalence (pooled) ---
d3 <- left_join(d1b, d2, by = c("pathogen", "community_type2")) %>%
  mutate(n_children = comm_n_lookup[as.character(community_type2)])
d3$pathogen <- factor(d3$pathogen, levels = pathogen_levels)

# --- C) SCR + seroprevalence ---
d4 <- left_join(d1, d1b, by = c("pathogen", "community_type2")) %>%
  mutate(n_children = comm_n_lookup[as.character(community_type2)])
d4$pathogen <- factor(d4$pathogen, levels = pathogen_levels)

# --- D) SCR + SRR ---
d5 <- left_join(d1b %>% rename(scr_per_100 = ir_per_100),
                sero_rev_type %>% select(pathogen, community_type2, ir_per_100) %>%
                  rename(srr_per_100 = ir_per_100), by = c("pathogen", "community_type2")) %>%
  mutate(n_children = comm_n_lookup[as.character(community_type2)])
d5$pathogen <- factor(d5$pathogen, levels = pathogen_levels)
```

# 1b) PCR Kaplan-Meier cumulative incidence at 24 months (for B's plot points)

Reuses `km_pcr_by_comm_round` (read in above), exported by
`7-ecomid_cumulative_incidence_data.qmd` -- no need to refit anything
here. That export already applies the correct "hold the last KM step
forward onto the nominal round grid" logic (including the EPS-rounding
fix for baseline-positive event times), which is more careful than
independently re-deriving a 24-month snapshot in this file would be.

**Data note (resolved):** `sero_conv`/`sero_rev` were confirmed to carry
the raw unaccented "Borbon" (like every raw-data table in this file
before relabeling), while `pcr_all`/`sero_all` are relabeled to accented
"Borbón" on read-in. `sero_conv`/`sero_rev` are now relabeled the same
way immediately after read-in (see above), so the matched-sample
bootstraps for A and C correctly include all three community types when
joining against `pcr_all`/`sero_all` by `community_type` (D's inputs,
`sero_conv` and `sero_rev`, are relabeled consistently with each other
regardless).

```{r}
# --- B) SCR + PCR KM cumulative incidence, 24mo (plot + bootstrap data) ---
pcr_cuminc24_type <- km_pcr_by_comm_round %>%
  filter(time == 24) %>%
  select(pathogen, community_type2, pcr_cum_inc_24 = cum_inc)

d9 <- left_join(d1b, pcr_cuminc24_type, by = c("pathogen", "community_type2")) %>%
  mutate(n_children = comm_n_lookup[as.character(community_type2)])
d9$pathogen <- factor(d9$pathogen, levels = pathogen_levels)
```

# 2) Matched-sample bootstrap Spearman correlations (A, B, C, D)

```{r}
# Format to exactly 2 significant figures, keeping trailing zeros
sig2 <- function(x) formatC(x, digits = 2, format = "fg", flag = "#")

# Used by C: individual-level seroprevalence (IgG positivity), from
# sero_all.
summarize_serop <- function(df) {
  df %>%
    group_by(pathogen, community_type) %>%
    summarise(
      n_pos = sum(seropos, na.rm = TRUE),
      n_total = sum(!is.na(seropos)),
      prev = n_pos / n_total,
      .groups = "drop"
    )
}

# Used by A, B, C, and D: individual-level seroconversion rate (SCR), from
# sero_conv. All four of this file's matched-sample-bootstrap panels are
# anchored on SCR.
summarize_sero <- function(df) {
  df %>%
    group_by(pathogen, community_type) %>%
    summarise(
      total_seroconversions = sum(seroconversion, na.rm = TRUE),
      total_pt_years = sum(person_time_years, na.rm = TRUE),
      incidence_rate_per_year = total_seroconversions / total_pt_years,
      .groups = "drop"
    )
}

# Used by B: refits a per-pathogen, per-community Kaplan-Meier curve on
# whichever children a bootstrap replicate resampled (via km_data_pcr_by_comm,
# read in above), then reads off cumulative incidence at exactly 24 months.
# summary(..., times = 24, extend = TRUE) is used instead of broom::tidy() +
# the km_at_rounds() forward-fill grid used elsewhere in this workflow --
# both read the same fitted step function, but this skips formatting work
# broom::tidy() does that isn't needed for a single time point per stratum,
# which matters since this runs on every one of B's 2000 replicates.
# extend = TRUE reproduces km_at_rounds()'s forward-fill behavior for a
# pathogen whose last observed round falls before 24 months.
fit_km_cuminc24 <- function(df) {
  df %>%
    group_by(pathogen) %>%
    group_modify(~ {
      fit <- survfit(Surv(tstart, tstop, event) ~ community_type, data = .x)
      s <- summary(fit, times = 24, extend = TRUE)
      tibble(
        community_type = sub("^community_type=", "", as.character(s$strata)),
        cum_inc_24 = 1 - s$surv
      )
    }) %>%
    ungroup()
}
```

## A) SCR + PCR prevalence

```{r}
###############################################################
##### Matched-sample bootstrap: 370 children resampled with  ##
##### replacement from the union of the two datasets          #
###############################################################

## --- Giardia EXCLUDED (this is the version used ON the plot) ---
sero_conv2 <- sero_conv %>% filter(pathogen != "Giardia")
pcr2 <- pcr_all %>% filter(pathogen != "Giardia")

summarize_pcr <- function(df) {
  df %>%
    group_by(pathogen, community_type) %>%
    summarise(
      n_pos = sum(pcrpos, na.rm = TRUE),
      n_total = sum(!is.na(pcrpos)),
      pcr_prev = n_pos / n_total,
      .groups = "drop"
    )
}

compute_corr_boot_p1 <- function(sero_df, pcr_df) {
  all_ids <- union(sero_df$childid, pcr_df$childid)
  boot_ids <- sample(all_ids, size = 370, replace = TRUE)
  boot_tbl <- tibble(childid = boot_ids)

  sero_resampled <- boot_tbl %>% left_join(sero_df, by = "childid", relationship = "many-to-many")
  pcr_resampled  <- boot_tbl %>% left_join(pcr_df,  by = "childid", relationship = "many-to-many")

  sero_sum <- summarize_sero(sero_resampled)
  pcr_sum  <- summarize_pcr(pcr_resampled)

  merged <- inner_join(sero_sum, pcr_sum, by = c("pathogen", "community_type"))
  if (nrow(merged) < 3) return(NA_real_)
  cor(merged$incidence_rate_per_year, merged$pcr_prev, method = "spearman")
}

run_corr_ci_p1 <- function(sero_df, pcr_df) {
  set.seed(239)
  B <- 2000
  boot_r <- replicate(B, compute_corr_boot_p1(sero_df, pcr_df))

  merged_obs <- inner_join(summarize_sero(sero_df), summarize_pcr(pcr_df),
                            by = c("pathogen", "community_type"))
  observed_rho <- cor(merged_obs$incidence_rate_per_year, merged_obs$pcr_prev, method = "spearman")

  tibble(
    rs    = observed_rho,
    rs_lo = quantile(boot_r, 0.025, na.rm = TRUE),
    rs_hi = quantile(boot_r, 0.975, na.rm = TRUE)
  ) %>%
    mutate(final_rho = paste0(sig2(rs), " (", sig2(rs_lo), ", ", sig2(rs_hi), ")"))
}

# Giardia excluded -- used ON plot A
corr_ci_p1_excl <- run_corr_ci_p1(sero_conv2, pcr2)
corr_ci_p1_excl

# Giardia included -- reported separately, NOT plotted (supplement)
corr_ci_p1_incl <- run_corr_ci_p1(sero_conv, pcr_all)
corr_ci_p1_incl
```

## B) SCR + PCR KM cumulative incidence, 24 months

Uses the same matched-sample bootstrap as A, C, and D (370 children
resampled with replacement, pathogen-level summaries recomputed on each
resample) rather than the row-level bootstrap over the ~27 aggregated
pathogen x community points that `10-ecomid_dotplot_correlations.qmd`
uses for this panel. The only structural difference from A/C/D is that
the PCR side needs an actual model refit -- `fit_km_cuminc24()`, defined
above -- instead of a simple resampled sum, since cumulative incidence
at 24 months comes from a Kaplan-Meier curve rather than a raw
proportion. `km_data_pcr_by_comm` (read in above) is what makes that
refit affordable inside a 2000-replicate loop: it holds every child's
own survival interval, precomputed once in
`7-ecomid_cumulative_incidence_data.qmd`, so each replicate below only
resamples children and refits `survfit()` -- it never rebuilds
intervals from raw testing rounds, which is what made an earlier
attempt at this too slow to run in practice.

Giardia handling matches (A): excluded from the ρ shown on the plot,
with the all-pathogen ρ computed separately for reference.
`sero_conv2` (Giardia excluded) is reused from (A)'s section above.
(Internal variable names below keep their historical `_f` suffix from
when this was panel F in the seven-panel version of this file.)

```{r}
km_data_pcr_by_comm2 <- km_data_pcr_by_comm %>% filter(pathogen != "Giardia")

compute_corr_boot_p4 <- function(sero_df, km_df) {
  all_ids <- union(sero_df$childid, km_df$childid)
  boot_ids <- sample(all_ids, size = 370, replace = TRUE)
  boot_tbl <- tibble(childid = boot_ids)

  scr_resampled <- boot_tbl %>% left_join(sero_df, by = "childid", relationship = "many-to-many")
  km_resampled  <- boot_tbl %>% left_join(km_df,  by = "childid", relationship = "many-to-many")

  scr_sum <- summarize_sero(scr_resampled)
  km_sum  <- fit_km_cuminc24(km_resampled)

  merged <- inner_join(scr_sum, km_sum, by = c("pathogen", "community_type"))
  if (nrow(merged) < 3) return(NA_real_)
  cor(merged$incidence_rate_per_year, merged$cum_inc_24, method = "spearman")
}

run_corr_ci_p4 <- function(sero_df, km_df) {
  set.seed(239)
  B <- 2000
  boot_r <- replicate(B, compute_corr_boot_p4(sero_df, km_df))

  merged_obs <- inner_join(summarize_sero(sero_df), fit_km_cuminc24(km_df),
                            by = c("pathogen", "community_type"))
  observed_rho <- cor(merged_obs$incidence_rate_per_year, merged_obs$cum_inc_24, method = "spearman")

  tibble(
    rs    = observed_rho,
    rs_lo = quantile(boot_r, 0.025, na.rm = TRUE),
    rs_hi = quantile(boot_r, 0.975, na.rm = TRUE)
  ) %>%
    mutate(final_rho = paste0(sig2(rs), " (", sig2(rs_lo), ", ", sig2(rs_hi), ")"))
}

# Giardia excluded -- used ON plot B
corr_ci_f <- run_corr_ci_p4(sero_conv2, km_data_pcr_by_comm2)
corr_ci_f

# Giardia included -- reported separately, NOT plotted (supplement)
corr_ci_f_incl <- run_corr_ci_p4(sero_conv, km_data_pcr_by_comm)
corr_ci_f_incl
```

## C) SCR + seroprevalence

```{r}
sero_prev_indiv <- sero_all

# summarize_serop() and summarize_sero() are both defined above, in
# section 2's shared preamble.

compute_corr_boot_p2 <- function() {
  all_ids <- union(sero_conv$childid, sero_prev_indiv$childid)
  boot_ids <- sample(all_ids, size = 370, replace = TRUE)
  boot_tbl <- tibble(childid = boot_ids)

  scr_resampled   <- boot_tbl %>% left_join(sero_conv, by = "childid", relationship = "many-to-many")
  serop_resampled <- boot_tbl %>% left_join(sero_prev_indiv, by = "childid", relationship = "many-to-many")

  scr_sum   <- summarize_sero(scr_resampled)
  serop_sum <- summarize_serop(serop_resampled)

  merged <- inner_join(serop_sum, scr_sum, by = c("pathogen", "community_type"))
  if (nrow(merged) < 3) return(NA_real_)
  cor(merged$incidence_rate_per_year, merged$prev, method = "spearman")
}

set.seed(239)
boot_r <- replicate(2000, compute_corr_boot_p2())

merged_obs <- inner_join(summarize_sero(sero_conv), summarize_serop(sero_prev_indiv),
                          by = c("pathogen", "community_type"))
observed_rho <- cor(merged_obs$incidence_rate_per_year, merged_obs$prev, method = "spearman")

corr_ci_p2 <- tibble(
  rs    = observed_rho,
  rs_lo = quantile(boot_r, 0.025, na.rm = TRUE),
  rs_hi = quantile(boot_r, 0.975, na.rm = TRUE)
) %>%
  mutate(final_rho = paste0(sig2(rs), " (", sig2(rs_lo), ", ", sig2(rs_hi), ")"))

corr_ci_p2
```

## D) SCR + SRR

```{r}
summarize_srr <- function(df) {
  df %>%
    group_by(pathogen, community_type) %>%
    summarise(
      total_seroreversions = sum(seroreversion, na.rm = TRUE),
      total_pt_years = sum(person_time_years, na.rm = TRUE),
      sr_rate_per_year = total_seroreversions / total_pt_years,
      .groups = "drop"
    )
}

compute_corr_boot_p3 <- function() {
  all_ids <- union(sero_conv$childid, sero_rev$childid)
  boot_ids <- sample(all_ids, size = 370, replace = TRUE)
  boot_tbl <- tibble(childid = boot_ids)

  scr_resampled <- boot_tbl %>% left_join(sero_conv, by = "childid", relationship = "many-to-many")
  srr_resampled <- boot_tbl %>% left_join(sero_rev,  by = "childid", relationship = "many-to-many")

  scr_sum <- summarize_sero(scr_resampled)
  srr_sum <- summarize_srr(srr_resampled)

  merged <- inner_join(srr_sum, scr_sum, by = c("pathogen", "community_type"))
  if (nrow(merged) < 3) return(NA_real_)
  cor(merged$incidence_rate_per_year, merged$sr_rate_per_year, method = "spearman")
}

set.seed(239)
boot_r <- replicate(2000, compute_corr_boot_p3())

merged_obs <- inner_join(summarize_sero(sero_conv), summarize_srr(sero_rev),
                          by = c("pathogen", "community_type"))
observed_rho <- cor(merged_obs$incidence_rate_per_year, merged_obs$sr_rate_per_year, method = "spearman")

corr_ci_p3 <- tibble(
  rs    = observed_rho,
  rs_lo = quantile(boot_r, 0.025, na.rm = TRUE),
  rs_hi = quantile(boot_r, 0.975, na.rm = TRUE)
) %>%
  mutate(final_rho = paste0(sig2(rs), " (", sig2(rs_lo), ", ", sig2(rs_hi), ")"))

corr_ci_p3
```

# 3) Plots

```{r}
my_colors <- c(
  "Campylobacter" = "darkgoldenrod1",
  "Cryptosporidium" = "cornflowerblue",
  "S. enterica" = "darkseagreen",
  "Shigella/EIEC" = "darkgreen",
  "E. histolytica" = "chocolate",
  "Norovirus GI" = "tan",
  "Norovirus GII" = "deeppink3",
  "LT-ETEC" = "blue4",
  "Giardia" = "grey70"
)

dotplot_theme <- theme(text = element_text(size = 14),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color = "black"),
  axis.ticks = element_line(color = "black"),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  legend.position = "none",
  panel.background = element_rect(fill = "white"),
  panel.grid = element_line(color = "grey", linewidth = 0.1))
```

## A) Seroconversion rate (SCR) vs. PCR prevalence

```{r}
p1 <- d3 %>% ggplot(aes(x = pcr_prevalence*100,
                         y = ir_per_100,
                         color = pathogen)) +
  geom_point(aes(size = n_children), alpha = 0.7) +
  scale_size(
    range = c(3, 10),
    breaks = c(85, 108, 177),
    labels = c("Esmeraldas (n=85)", "Borbón (n=108)", "Rural (n=177)"),
    name = "Community type"
  ) +
  labs(x = "PCR prevalence (%)",
       y = "IgG seroconversion rate (per 100 py)") +
  scale_color_manual(values = my_colors, name = "") +
  scale_x_continuous(breaks = seq(0, 30, by = 5), limits = c(0, 30)) +
  dotplot_theme +
  coord_cartesian(ylim = c(0, 200)) +
  geom_text(
    data = corr_ci_p1_excl,
    aes(x = -Inf, y = Inf, label = paste0("ρ = ", final_rho)),
    inherit.aes = FALSE,
    hjust = -0.1, vjust = 1.5, size = 5, color = "black"
  ) +  annotate("text", x = 28, y = 10, label = "Giardia excluded from ρ",
           hjust = 1, vjust = 0, size = 3.5, color = "darkgrey", fontface = "italic")

p1
```

## C) SCR vs. Seroprevalence

```{r}
p2 <- d4 %>% ggplot(aes(x = sero_prevalence*100, y = ir_per_100, color = pathogen)) +
  geom_point(aes(size = n_children), alpha = 0.7) +
  labs(x = "IgG seroprevalence (%)",
       y = "IgG seroconversion rate (per 100 py)") +
  scale_color_manual(values = my_colors, name = "") +
  scale_size(
    range = c(3, 10),
    breaks = c(85, 108, 177),
    labels = c("Esmeraldas (n=85)", "Borbón (n=108)", "Rural (n=177)"),
    name = "Community type"
  ) +
  theme(text = element_text(size = 14),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color = "black"),
  axis.ticks = element_line(color = "black"),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  legend.position = "right",
  panel.background = element_rect(fill = "white"),
  panel.grid = element_line(color = "grey", linewidth = 0.1)) +
  geom_text(
    data = corr_ci_p2,
    aes(x = -Inf, y = Inf, label = paste0("ρ = ", final_rho)),
    inherit.aes = FALSE,
    hjust = -0.1, vjust = 1.5, size = 5, color = "black"
  ) +
  xlim(0, 60) +
  coord_cartesian(ylim = c(0, 200))

p2
```

## D) SCR vs. SRR

```{r}
p3 <- d5 %>% ggplot(aes(x = scr_per_100, y = srr_per_100, color = pathogen)) +
  geom_point(aes(size = n_children), alpha = 0.7) +
  labs(x = "Seroconversion rate (per 100 py)",
       y = "Seroreversion rate (per 100 py)") +
  scale_color_manual(values = my_colors, name = "") +
  scale_size(
    range = c(3, 10),
    breaks = c(85, 108, 177),
    labels = c("Esmeraldas (n=85)", "Borbón (n=108)", "Rural (n=177)"),
    name = "Community type"
  ) +
  theme(text = element_text(size = 12),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color = "black"),
  axis.ticks = element_line(color = "black"),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  legend.position = "right",
  panel.background = element_rect(fill = "white"),
  panel.grid = element_line(color = "grey", linewidth = 0.1)) +
  geom_text(
    data = corr_ci_p3,
    aes(x = -Inf, y = Inf, label = paste0("ρ = ", final_rho)),
    inherit.aes = FALSE,
    hjust = -0.1, vjust = 1.5, size = 4, color = "black"
  ) +
  xlim(0, 200) +
  coord_cartesian(ylim = c(0, 300), clip = "off") +
  scale_y_continuous(limits = c(0, 300), breaks = seq(0, 300, by = 50))

p3

## Saved as its own ggplot object -- not folded into the combined figure
## below, so it can be read into a separate R file and composed into a
## supplemental figure there.
saveRDS(p3, here("output/dotplot_scr_vs_srr.rds"))
```

## B) SCR vs. PCR KM cumulative incidence, 24 months

```{r}
p6 <- d9 %>% ggplot(aes(x = pcr_cum_inc_24*100, y = ir_per_100, color = pathogen)) +
  geom_point(aes(size = n_children), alpha = 0.7) +
  labs(x = "PCR cumulative incidence at 24m (%)",
       y = "Seroconversion rate (per 100 py)") +
  scale_color_manual(values = my_colors, name = "") +
  scale_size(
    range = c(3, 10),
    breaks = c(85, 108, 177),
    labels = c("Esmeraldas (n=85)", "Borbón (n=108)", "Rural (n=177)"),
    name = NULL
  ) +
  dotplot_theme +
  coord_cartesian(ylim = c(0, 200)) +
  geom_text(
    data = corr_ci_f,
    aes(x = -Inf, y = Inf, label = paste0("ρ = ", final_rho)),
    inherit.aes = FALSE,
    hjust = -0.1, vjust = 1.5, size = 5, color = "black"
  ) +
  annotate("text", x = 90, y = 2, label = "Giardia excluded from ρ",
           hjust = 1, vjust = 0, size = 3.5, color = "darkgrey", fontface = "italic")

p6
```

# 4) Combined 3-panel figure

Redesigned as a single row of three panels: document panel A (p1, SCR
vs. PCR prevalence, pooled), document panel B (p6, SCR vs. PCR KM
cumulative incidence at 24 months), and document panel C (p2, SCR vs.
seroprevalence, pooled), left to right. Because the combined figure's
panel order now matches the document's panel order exactly, no separate
composite lettering is needed -- the A)/B)/C) tags below are the same
A/B/C as the rest of this document. D (SCR vs. SRR) is not part of this
combined figure; it remains standalone, and is additionally exported on
its own as a saved ggplot object just above.

```{r, fig.width = 15, fig.height = 7}
## plot_layout(guides = "collect") turned out to be unreliable across
## panels built from different source data frames (each panel's color/size
## scale has to be byte-for-byte identical -- same factor levels, same
## order -- for patchwork to recognize them as mergeable; a mismatch
## anywhere produces exactly the "one legend per row" duplication seen
## above). Sidestepping that entirely: extract legends directly via
## cowplot::get_legend() from p1 (A) -- d3's pathogen factor has all 9
## levels explicitly set and its size scale's breaks/labels are
## hardcoded, so it's a safe, always-complete source -- suppress the
## legend completely on all three panels.
library(cowplot)
library(gridExtra)

## legend.key.size / legend.text shrink the extracted legend's own native
## footprint. This matters because draw_plot() (below) does NOT rescale a
## raw grob to fit the box it's given the way it rescales a ggplot -- the
## grob just draws at its native size, and the previous attempt's legend
## was bigger than its box, so its third row got clipped by the page
## edge. Shrinking it here is what actually fixes that, not the box size.
color_legend_source <- p1 + theme(legend.position = "right",
                                   legend.text = element_text(size = 9),
                                   legend.key.size = unit(0.4, "cm"),
                                   legend.spacing.y = unit(0.05, "cm"),
                                   legend.margin = margin(0, 0, 0, 0),
                                   legend.box.margin = margin(0, 0, 0, 0)) +
  guides(size = "none", color = guide_legend(nrow = 2, byrow = TRUE))
color_legend <- cowplot::get_legend(color_legend_source)

size_legend_source <- p1 + theme(legend.position = "right",
                                  legend.text = element_text(size = 9),
                                  legend.key.size = unit(0.4, "cm"),
                                  legend.spacing.y = unit(0.05, "cm"),
                                  legend.margin = margin(0, 0, 0, 0),
                                  legend.box.margin = margin(0, 0, 0, 0)) +
  guides(color = "none", size = guide_legend(nrow = 1, title = NULL))
size_legend <- cowplot::get_legend(size_legend_source)

## Glue the two legend blocks directly together (no gap between them)
## into one compact grob -- gtable_cbind() preserves each block's natural
## width with no padding added in between.
color_legend <- gtable::gtable_add_cols(color_legend, widths = unit(0.5, "cm"))

combined_legend <- gridExtra::gtable_cbind(color_legend, size_legend)
## aspect.ratio = 1 forces each panel to draw as a physical square (same
## on-page length for the x- and y-axis) regardless of how different the
## underlying data ranges are (e.g. 0-200 for SCR vs. 0-30 for PCR
## prevalence in panel A). This is a purely visual constraint -- unlike
## coord_equal(), which scales by data units and would badly distort
## these panels given how different those ranges are.
## Only the first (leftmost) panel, A, keeps its y-axis title. A, B, and
## C all share the same y-axis quantity now (seroconversion rate), so
## labeling it once on the left avoids repeating the same label three
## times; each panel's x-axis is a different quantity (PCR prevalence,
## PCR cumulative incidence, seroprevalence), so all three keep their own
## x-axis title.
panel_row <- wrap_plots(
  p1 + theme(legend.position = "none", aspect.ratio = 1),
  p6 + theme(legend.position = "none", axis.title.y = element_blank(), aspect.ratio = 1),
  p2 + theme(legend.position = "none", axis.title.y = element_blank(), aspect.ratio = 1),
  ncol = 3
) +
  plot_annotation(tag_levels = list(c("A)", "B)", "C)"))) &
  theme(plot.tag = element_text(size = 16, face = "bold"))

## panel_row occupies the top 76% of the canvas; combined_legend sits in
## the bottom-left, sized to 55% of the width and 28% of the height, with
## a small y-buffer (0.02 instead of 0) so it isn't drawn flush against
## the literal bottom edge of the page. These are still the numbers to
## nudge once you see the real render.
combined_dotplots <- ggdraw() +
  draw_plot(panel_row, x = 0, y = 0.24, width = 1, height = 0.76) +
  draw_plot(combined_legend, x = 0.1, y = 0.06, width = 0.55, height = 0.28)

combined_dotplots

ggsave(here("figures/fig4_dotplot_correlations_combined_3panel.tiff"),
       combined_dotplots, height = 7, width = 15)
```

