---
title: "ECoMiD pathogen prevalence by age (PCR and IgG)"
author: "Nikolina Walas (nikolina.walas@ucsf.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
---

## Overview

The purpose of this file is to generate enteric pathogen prevalence by age (in months) of children from the ecomid cohort, as measured by PCR and by serology.

6, 12, 18 PCR results were measured by TAC; 3, 9, 15, 24 measurements were done by GPP.

## Load packages

```{r}
library(here)
source(here("R", "0-config.R"))
```

## Read in files

```{r}
# PCR raw data 
pcr <- readRDS(here("data","ecomid_pcr_analysis_data_2026-06-11.rds"))
print("Safety check: number of childid with pcr results")
length(unique(pcr$childid))

# Serology raw
sero <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_2026-05-19.csv")) 
print("Safety check: number of childid with serology results")
length(unique(sero$childid))
```

## Calculate seroprevalence by age

```{r}
# check unique timepoints for age 
print("Safety check: unique timepoints with serology results")
unique(sero$timepoint)

# summarise proportion positive at each timepoint (this is being calculated at the antigen level...we need it at the pathogen level)
sp <- sero %>%  filter(pathogen == "Campylobacter" | pathogen == "LT-ETEC" | pathogen == "Shigella/EIEC" | 
         pathogen == "Cryptosporidium" | pathogen == "E. histolytica" | pathogen == "Giardia" |
         pathogen == "Norovirus GI" | pathogen == "Norovirus GII" | pathogen == "Salmonella") %>% 
  group_by(childid, timepoint, pathogen, pathogen_lab) %>%
  filter(!is.na(log10mfi)) %>% 
  summarise(
    seropos = max(seropos, na.rm = TRUE),  # max across antigens
    .groups = "drop"
  ) %>% group_by(timepoint, pathogen) %>% summarise(n = sum(!is.na(seropos)),
                  pos = sum(seropos == 1, na.rm = TRUE), 
                  prop = round((pos / n)*100, 2)) %>% 
  rowwise() %>% 
  # estimate prevalence and 95% CIs.
  mutate(ci_low  = prop.test(pos, n)$conf.int[1] * 100,
         ci_high = prop.test(pos, n)$conf.int[2] * 100) 

# set levels 
# updated to accomodate salmonella 
sp$pathogen <- factor(sp$pathogen, levels = c("Cryptosporidium", "Giardia", "E. histolytica", 
                                              "LT-ETEC", "Shigella/EIEC", "Campylobacter", 
                                              "Norovirus GII","Norovirus GI", "Salmonella"))

```

## Calculate PCR prevalence by age

```{r}
# check unique timepoints for age 
print("Safety check: unique timepoints with pcr results")
unique(pcr$visit)

# summarise proportion positive at each timepoint 
pcr_pos <- pcr %>% group_by(visit, pathogen) %>% 
  summarise(n = sum(!is.na(pcrpos)), 
            pos = sum(pcrpos == 1, na.rm = TRUE), 
            prop = round((pos / n)*100, 2)) %>% 
  filter(pathogen == "Campylobacter" | pathogen == "LT-ETEC" | pathogen == "Shigella/EIEC" | 
         pathogen == "Cryptosporidium" | pathogen == "E. histolytica" | pathogen == "Giardia" |
         pathogen == "Norovirus GI" | pathogen == "Norovirus GII" | pathogen == "Salmonella") %>% rowwise() %>% 
  mutate(ci_low  = prop.test(pos, n)$conf.int[1] * 100,
         ci_high = prop.test(pos, n)$conf.int[2] * 100) 

# set levels
pcr_pos$pathogen <- factor(pcr_pos$pathogen, levels = c("Cryptosporidium", "Giardia", "E. histolytica", 
                                              "LT-ETEC", "Shigella/EIEC", "Campylobacter", 
                                              "Norovirus GII","Norovirus GI", "Salmonella"))
```

## Raw age-based prevalence

```{r}
## Seroprevalence by age
sp_plot <- sp %>% 
  ggplot(aes(x = timepoint, y = prop, 
             fill = pathogen, 
             color = pathogen)) + 
  # geom_errorbar(aes(ymin = seroprev_min95*100, ymax = seroprev_max95*100), width = 0.5) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c(cbpal, "red")) +
  scale_x_continuous(breaks = seq(0, 24, by = 2)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) + 
  coord_cartesian(ylim=c(0,100),xlim=c(0,24)) +
  labs(x = "Age (months)", y = "Seroprevalence (%)") +
  theme_minimal()

sp_plot

## pcr prevalence by age
pcr_plot <- pcr_pos %>% 
  ggplot(aes(x = visit, y = prop, 
             fill = pathogen, 
             color = pathogen)) + 
  # geom_errorbar(aes(ymin = seroprev_min95*100, ymax = seroprev_max95*100), width = 0.5) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c(cbpal, "red")) +
  scale_x_continuous(breaks = seq(0, 24, by = 2)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) + 
  coord_cartesian(ylim=c(0,100),xlim=c(0,24)) +
  labs(x = "Age (months)", y = "PCR prevalence (%)") +
  theme_minimal()

pcr_plot

```

## Pair the seroprevalence and the pcr prevalence

```{r}
pcr_sp <- rbind(sp %>% mutate(assay = "IgG"), 
                pcr_pos %>% rename(timepoint = visit) %>% 
                  mutate(assay = "PCR"))  #%>% 
  #mutate(pathogen_class = case_when(pathogen == "Campylobacter" |
  #                                    pathogen == "LT-ETEC" |
  #                                    pathogen == "Shigella" ~ "Bacteria", 
  #                                  pathogen == "Cryptosporidium" |
  #                                    pathogen == "E. histolytica" |
  #                                    pathogen == "Giardia" ~ "Protozoa", 
  #                                  TRUE ~ "Virus"))

# export for further comparative analyses
# write.csv(pcr_sp, here("output/ecomid_pcr_sero_age_based.csv"), row.names = F)

p1 <- pcr_sp %>% 
  ggplot(aes(
    x = timepoint, 
    y = prop,  
    color = assay,   # distinguishes the lines
    group = assay    # ensures separate lines
  )) + 
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("#CC79A7","#009E73")) +
  scale_fill_manual(values = c("#CC79A7","#009E73")) +
  scale_x_continuous(breaks = seq(0, 24, by = 3)) +
  #scale_y_continuous(breaks = seq(0, 100, by = 10)) + 
  coord_cartesian(xlim = c(0, 24), ylim = c(0, 85)) +
  labs(
    x = "Age (months)", 
    y = "Prevalence (%)",
    #color = "Assay"
  ) +
  # CI shading
  geom_ribbon(aes(
    ymin = ci_low,      # if prop is in percent
    ymax = ci_high, 
    fill = assay), alpha = 0.15, color = NA) + 
  theme_minimal() +
  facet_wrap(~ pathogen, ncol = 3, drop = FALSE) + 
  theme(base_size = 12, 
        legend.position = c(0.925,0.9),  # x=0.8 (right), y=0.2 (bottom)
    legend.background = element_rect(fill = alpha("white", 0.7)), # semi-transparent bg
    legend.title = element_blank(),
    legend.text = element_text(size = 12), 
    panel.grid.minor = element_blank(), 
    strip.text = element_text(size = 12),  
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12)
  )
p1

ggsave(here("figures/age_based_prevalence.tiff"), plot = p1, height = 8, width = 10)
```
