---
title: "Serology and PCR overview: supplemental tables and rate figures"
subtitle: "Population-level summary tables and SCR/SRR figures -- Objective 2 of the PCR-vs-serology cleanup"
author: "Nikolina Walas (nikolina.walas@ucsf.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute:
  echo: true
  warning: false
  message: false
---

## File Overview

This file generates the sampling/prevalence
supplemental tables and the seroconversion/seroreversion rate figures for the manuscript.

Output from this document:

1.  Supplemental tables for the ecomid pcr-serology manuscript: sampling
    counts by round/assay, and pathogen-level prevalence by IgG and PCR
2.  Supplemental figures comparing seroconversion and seroreversion
    rates at the community-type level and in binned age groups (\<12
    months and \>= 12 months)

There are **370** unique children included in the serology analysis (the
Table 1 / final enrolled sample set), contributing **1601** individual
samples. There are **370** unique children included in the PCR analysis,
contributing **2231** individual samples. All 370 serology children have
a PCR result at some timepoint.

| Months | PCR assay (n = 370 children) (N = 2231 samples) | Serology Assay (n = 370 children) (N = 1601 samples) |
|----|----|----|
| 3 | X |  |
| 6 | X | X |
| 9 | X | X |
| 12 | X | X |
| 15 | X |  |
| 18 | X | X |
| 24 | X | X |

### Load packages

```{r}
library(here)
source(here("R", "0-config.R"))
```

### Read in files

```{r}
# seroconversion by community type + pathogen
sero_inc_type <- readRDS(here("output/ecomid_pathogen_comm_sero_inc_3_levels.rds"))  %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium") %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
sero_inc_type$community_type2 <- factor(sero_inc_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

# seroconversion by age group
sero_inc_type2 <- readRDS(here("output/ecomid_pathogen_agegroup_sero_inc.rds")) %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium")

# seroprevalence by community type + pathogen
sero_prev_type <- readRDS(here("output/ecomid_sero_prev_rates_3_levels.rds")) %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
sero_prev_type$community_type2 <- factor(sero_prev_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

# pcr prevalence by community type + pathogen
pcr_prev_type <- readRDS(here("output/ecomid_pcr_prev_by_community_3_levels.rds")) %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
pcr_prev_type$community_type2 <- factor(pcr_prev_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

# seroreversion rate by community type + pathogen
sero_rev_type <- readRDS(here("output/ecomid_pathogen_comm_sero_rev_3_levels.rds")) %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium") %>%
  mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
sero_rev_type$community_type2 <- factor(sero_rev_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

# seroreversion by age group
sero_rev_type2 <- readRDS(here("output/ecomid_pathogen_agegroup_sero_rev.rds")) %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium")

# overall measures (by pathogen)
sero_prev_path <- readRDS(here("output/ecomid_sero_prev_rates.rds")) %>%
  filter(community_name == "All communities")

pcr_prev_path <- readRDS(here("output/ecomid_pcr_prev_by_pathogen.rds"))

# raw measures (by timepoint) -- needed only for the sampling-count table below
pcr_all <- readRDS(here("data","ecomid_pcr_analysis_data_2026-08-12.rds"))
sero_all <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_2026-08-26.csv")) %>%
  mutate(community_type = ifelse(str_detect(community_type, "Rural"), "Rural", community_type)) %>%
  mutate(community_type = ifelse(community_type == "Borbon", "Borbón", community_type)) %>%
  filter(antigen_lab != "S. enterica HlyE")
```

# 1) Overall prevalence (supplemental table 1+3)

```{r}
#----------#
# Count of samples at each timepoint
#----------#

# pcr
length(unique(pcr_all$childid))
pcr_sum <- pcr_all %>% group_by(visit, pathogen) %>%
  summarise(n = n()) %>%
  arrange(pathogen) %>%
  ungroup() %>%
  group_by(visit) %>%
  summarise(max_n_pcr = max(n)) %>% mutate(timepoint = paste0("month_", visit)) %>% select(-visit) %>%
  pivot_wider(names_from = timepoint, values_from = max_n_pcr) %>% mutate(collection = "stool") %>%
  select(8, 1:7)

pcr_sum

# serology
igg_sum <- sero_all %>% group_by(timepoint, antigen_lab) %>%
  summarise(n = n()) %>%
  arrange(antigen_lab) %>%
  ungroup() %>%
  group_by(timepoint) %>%
  summarise(max_n_igg = max(n)) %>%
  mutate(timepoint = paste0("month_", timepoint)) %>%
  pivot_wider(names_from = timepoint, values_from = max_n_igg) %>% mutate(collection = "blood") %>%
  select(6, 1:5) %>%
  mutate(month_3 = NA,
         month_15 = NA)
igg_sum

sup_tab_1 <- rbind(pcr_sum, igg_sum) %>% mutate(total_samples = rowSums(across(2:8), na.rm = TRUE))
writexl::write_xlsx(sup_tab_1, here("output/supp_table_1_sampling.xlsx"))

# IQR for manuscript
# pcr
bychild <- pcr_all %>% select(childid, visit) %>% unique() %>% group_by(childid) %>% summarise(n_samples = n())
bychild %>% summarise(median = median(n_samples),
                      mean = mean(n_samples),
                      IQR = IQR(n_samples))
quantile(bychild$n_samples, probs = c(0.25, 0.5, 0.75))

# igg
bychild <- sero_all %>% select(childid, timepoint) %>% unique() %>% group_by(childid) %>% summarise(n_samples = n())
bychild %>% summarise(median = median(n_samples),
                      mean = mean(n_samples),
                      IQR = IQR(n_samples))
quantile(bychild$n_samples, probs = c(0.25, 0.5, 0.75))

#----------#
# pathogen prevalence over all timepoints
#----------#

igg_path <- sero_prev_path %>% mutate(prev = paste0(round(sero_prevalence*100, 1),
                                                    " (",
                                           round(lower_ci*100, 1), ", ",
                                           round(upper_ci*100, 1), ")")) %>%
  select(-sero_prevalence, -lower_ci, -upper_ci, -community_name, -community_type) %>%
  mutate(assay = "igg_prevalence") %>%
  filter(pathogen != "Strongyloides" & pathogen != "T. solium")

pcr_path <- pcr_prev_path %>% mutate(prev = paste0(round(prevalence*100, 1), " (",
                                           round(lower_ci*100, 1), ", ", round(upper_ci*100, 1), ")")) %>%
  select(-prevalence, -lower_ci, -upper_ci, -community_name) %>%
  mutate(assay = "pcr_prevalence") %>%
  filter(pathogen %in% igg_path$pathogen)

s_tab2 <- rbind(igg_path, pcr_path) %>% group_by(pathogen) %>%
  select(pathogen, assay, n_pos, n_total, prev) %>%
  arrange(pathogen, desc(assay))

s_tab2

writexl::write_xlsx(s_tab2, here("output/supp_table_3_prev_pathogen.xlsx"))
```

# 2) Population measures by community

### Seroconversion rate per pathogen

```{r}
#| fig-width: 8
#| fig-height: 8

sero_inc_type %>%
  arrange(community_type2) %>%
  ggplot(aes(x = community_type2, y = ir_per_100, color = community_type2)) +
  geom_point() +
  labs(y = "Seroconversion Rate per 100 Person-Years",
       x = "Community Type",
       title = "Seroconversion Rate Across Community type") +
  scale_color_manual(values = nytpal) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10),
    legend.position = "right",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(color = "grey", linewidth = 0.1)) +
  geom_errorbar(aes(ymin = ir_lo_per_100, ymax = ir_hi_per_100), width = 0.1, linewidth = 0.5) +
  facet_wrap( ~ pathogen, ncol= 4, scales = "free_y")
```

### Seroreversion rate per pathogen

```{r}
#| fig-width: 8
#| fig-height: 8

sero_rev_type$community_type2 <- factor(sero_rev_type$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

sero_rev_type %>%
  arrange(community_type2) %>%
  ggplot(aes(x = community_type2, y = ir_per_100, color = community_type2)) +
  geom_point() +
  labs(y = "Seroreversion Rate per 100 Person-Years",
       x = "Community Type",
       title = "Seroconversion Rate Across Community type") +
  scale_color_manual(values = nytpal) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10),
    legend.position = "right",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(color = "grey", linewidth = 0.1)) +
  geom_errorbar(aes(ymin = ir_lo_per_100, ymax = ir_hi_per_100), width = 0.1, linewidth = 0.5) +
  facet_wrap( ~ pathogen, ncol= 4, scales = "free_y")
```

### SUPP FIG: SCR and SRR composite figures

```{r}
##--------------------------##
##------ BY COMMUNITY ------##
##--------------------------##

combo_df <- sero_rev_type %>% mutate(metric = "Seroreversion")  %>%
  rbind(sero_inc_type %>% mutate(metric = "Seroconversion (force of infection)"))

combo_df$community_type2 <- as.character(combo_df$community_type2)

combo_df$community_type2 <- factor(
  combo_df$community_type2,
  levels = c("Esmeraldas", "Rural", "Borbón")
)

combo_df$pathogen <- factor(
  combo_df$pathogen,
  levels = c("Campylobacter", "LT-ETEC", "Norovirus GII", "Cryptosporidium",
             "Shigella/EIEC", "Norovirus GI", "S. enterica", "Giardia", "E. histolytica")
)

p_combo <- ggplot(
  combo_df,
  aes(
    x = ir*100,
    y = pathogen,
    color = community_type2,
    group = community_type2
  )
) +
  scale_y_discrete(limits = rev(levels(combo_df$pathogen))) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(
    aes(xmin = ir_lo*100, xmax = ir_hi*100),
    position = position_dodge(width = 0.5),
    height = 0.2,
    linewidth = 0.5
  ) +
  scale_color_manual(
    values = nytpal,
    breaks = c("Esmeraldas", "Borbón", "Rural")
  ) +
  facet_wrap(~ metric, scales = "free_x") +
  theme_minimal() +
  labs(y = "", x = "rate per 100 child-years at risk") +
  theme(text = element_text(size = 12),
    legend.position = c(0.9, 0.8),
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", color = "white"),
    panel.border = element_blank()
  )

p_combo <- p_combo +
  ggh4x::facetted_pos_scales(
    x = list(
      metric == "Seroconversion" ~ scale_x_continuous(limits = c(0, 200)),
      metric == "Seroreversion" ~ scale_x_continuous(limits = c(0, 500))
    )
  )

p_combo

ggsave(here("figures/suppfig7_scr_srr_communitytype.tiff"), p_combo, width = 210, height = 120, units = "mm")

##--------------------------##
##-------- BY AGE ----------##
##--------------------------##

combo_df2 <- sero_rev_type2 %>% mutate(metric = "Seroreversion")  %>%
  rbind(sero_inc_type2 %>% mutate(metric = "Seroconversion (force of infection)"))

combo_df2$age_group <- factor(combo_df2$age_group,
                              levels = c("12-24 months", "6-9 months"))

combo_df2$pathogen <- factor(combo_df2$pathogen,
                            levels = c("Campylobacter", "LT-ETEC",
                                       "Norovirus GII", "Cryptosporidium",
                                       "Shigella/EIEC",
                                       "Norovirus GI", "S. enterica",
                                       "Giardia",  "E. histolytica"))

p_combo2 <- ggplot(combo_df2, aes(x = ir*100,
                     y = pathogen,
                     color=age_group,
                     group = age_group )) +
  scale_y_discrete(limits = rev(levels(combo_df2$pathogen))) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = ir_lo*100,
                     xmax = ir_hi*100),
                 position = position_dodge(width = 0.5),
                 height = 0.2,
                 linewidth = 0.5) + scale_color_manual(
  values = nytpal,
  breaks = rev(levels(combo_df2$age_group))
) +
  facet_wrap(~ metric, scales = "free_x") +
  theme_minimal() +
  labs(y = "",
       x = "rate per 100 child-years at risk") +
  theme(
      legend.position = c(0.9, 0.8),
      legend.title = element_blank(),
      legend.background = element_rect(fill = "white", color = "white"),
      panel.border = element_blank())

p_combo2

#ggsave(here("figures/suppfig5_seroconversion_seroreversion_agegroup.tiff"), p_combo2, height = 5, width = 8)

# Age with SCR/SRR alignment 

p3 <- readRDS(here("output/dotplot_scr_vs_srr.rds"))

p3   # renders it as-is

bottom_row <- plot_grid(p3, NULL, ncol = 2, rel_widths = c(10, 1),
                         labels = c("B", ""))
final <- plot_grid(p_combo2, bottom_row, ncol = 1, labels = c("A", ""))
final

ggsave(here("figures/suppfig5_scr_srr_agegroup_dotplot.tiff"), final, width = 220, height = 210, units = "mm")
```

### PCR prevalence per pathogen

```{r}
#| fig-width: 8
#| fig-height: 8

pcr_prev_type %>%
  arrange(community_type2) %>%
  ggplot(aes(x = community_type2, y = pcr_prevalence, color = community_type2)) +
  geom_point() +
  labs(y = "Prevalence",
       x = "",
       title = "PCR Prevalence across community types") +
  scale_color_manual(values = nytpal) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10),
    legend.position = "right",
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(color = "grey", linewidth = 0.1)) +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.1, linewidth = 0.5) +
  facet_wrap( ~ pathogen, ncol= 3, scales = "free_y")
```
