---
title: "ECoMiD pathogen sero-prevalence by urbanicity (3 levels)"
author: "Nikolina Walas (nikolina.walas@ucsf.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
---

## Overview

The purpose of this file is to generate enteric pathogen prevalence by age (in months) and urbanicity of children from the ecomid cohort, as measured by PCR and by IgG serology. The IgG serology is reported in the manuscript, while the PCR is supplemental. 

6, 12, 18 PCR results were measured by TAC; 3, 9, 15, 24 measurements were done by GPP.

## Load packages

```{r}
library(here)
source(here("R", "0-config.R"))
```

## Read in files

```{r}
# PCR raw data 
pcr <- readRDS(here("data","ecomid_pcr_analysis_data_2026-06-11.rds")) %>% 
  mutate(community_type2 = case_when(community_type == "Rural Road" ~ "Rural", 
                                     community_type == "Rural River" ~ "Rural", 
                                     community_type == "Esmeraldas" ~ "Esmeraldas City", 
                                     community_type == "Borbon" ~ "Borbon"))
print("Safety check: number of childid with pcr results")
length(unique(pcr$childid))

# Serology raw
sero <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_2026-05-19.csv")) %>% 
  mutate(community_type2 = case_when(community_type == "Rural Road" ~ "Rural", 
                                     community_type == "Rural River" ~ "Rural", 
                                     community_type == "Esmeraldas" ~ "Esmeraldas City", 
                                     community_type == "Borbon" ~ "Borbon"))
print("Safety check: number of childid with serology results")
length(unique(sero$childid))
```

## Calculate seroprevalence by age

```{r}
# check unique timepoints for age 
print("Safety check: unique timepoints with serology results")
unique(sero$timepoint)

# summarise proportion positive at each timepoint (this is being calculated at the antigen level...we need it at the pathogen level)
sp <- sero %>%  filter(pathogen == "Campylobacter" | 
                         pathogen == "LT-ETEC" | 
                         pathogen == "Shigella/EIEC" | 
                         pathogen == "Cryptosporidium" | 
                         pathogen == "E. histolytica" | 
                         pathogen == "Giardia" |
                         pathogen == "Norovirus GI" | 
                         pathogen == "Norovirus GII" | pathogen == "Salmonella") %>% 
  group_by(childid, timepoint, pathogen, pathogen_lab, community_type2) %>%
  filter(!is.na(log10mfi)) %>% 
  summarise(
    seropos = max(seropos, na.rm = TRUE),  # max across antigens
    .groups = "drop"
  ) %>% group_by(timepoint, pathogen, community_type2) %>% summarise(n = sum(!is.na(seropos)),
                  pos = sum(seropos == 1, na.rm = TRUE), 
                  prop = round((pos / n)*100, 2)) %>% 
  rowwise() %>% 
  # estimate prevalence and 95% CIs.
  mutate(ci_low  = prop.test(pos, n)$conf.int[1] * 100,
         ci_high = prop.test(pos, n)$conf.int[2] * 100) 

# set levels 
# updated to accomodate salmonella 
sp$pathogen <- factor(sp$pathogen, levels = c("Cryptosporidium", "Giardia", "E. histolytica", 
                                              "LT-ETEC", "Shigella/EIEC", "Campylobacter", 
                                              "Norovirus GII","Norovirus GI", "Salmonella"))

```

## Calculate PCR prevalence by age

```{r}
# check unique timepoints for age 
print("Safety check: unique timepoints with pcr results")
unique(pcr$visit)

# summarise proportion positive at each timepoint 
pcr_pos <- pcr %>% group_by(visit, pathogen, community_type2) %>% 
  summarise(n = sum(!is.na(pcrpos)), 
            pos = sum(pcrpos == 1, na.rm = TRUE), 
            prop = round((pos / n)*100, 2)) %>% 
  filter(pathogen == "Campylobacter" | pathogen == "LT-ETEC" | pathogen == "Shigella/EIEC" | 
         pathogen == "Cryptosporidium" | pathogen == "E. histolytica" | pathogen == "Giardia" |
         pathogen == "Norovirus GI" | pathogen == "Norovirus GII" | pathogen == "Salmonella") %>% rowwise() %>% 
  mutate(ci_low  = prop.test(pos, n)$conf.int[1] * 100,
         ci_high = prop.test(pos, n)$conf.int[2] * 100) 

# set levels
pcr_pos$pathogen <- factor(pcr_pos$pathogen, levels = c("Norovirus GII", "LT-ETEC", "Shigella/EIEC", "Campylobacter", "Norovirus GI", "Cryptosporidium", "Giardia", "E. histolytica", "Salmonella"))
```


## Pair the seroprevalence and the pcr prevalence

```{r}
pcr_sp <- rbind(sp %>% mutate(assay = "IgG"), 
                pcr_pos %>% rename(timepoint = visit) %>% 
                  mutate(assay = "PCR"))  
pcr_sp$community_type2 <- factor(pcr_sp$community_type2, 
                                 levels = c("Esmeraldas City", "Borbon", "Rural"))
  
  
p1 <- pcr_sp %>% 
  ggplot(aes(
    x = timepoint, 
    y = prop,  
    color = community_type2,   # distinguishes the lines
    group = community_type2    # ensures separate lines
  )) + 
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("#510000", "#AC112D", "#EC6D47")) +
  scale_fill_manual(values = c( "#510000", "#AC112D", "#EC6D47")) +
  scale_x_continuous(breaks = seq(0, 24, by = 3)) +
  coord_cartesian(xlim = c(3, 24), ylim = c(0, 85)) +
  labs(
    x = "Age (months)", 
    y = "Prevalence (%)",
    #color = "Assay"
  ) +
  # CI shading
  geom_ribbon(aes(
    ymin = ci_low,      # if prop is in percent
    ymax = ci_high, 
    fill = community_type2), alpha = 0.15, color = NA) + 
  theme_minimal() +
  facet_wrap(~ pathogen + assay, ncol = 4, drop = FALSE) + 
  theme(base_size = 12, 
        legend.position = c(0.05,0.9),  # x=0.8 (right), y=0.2 (bottom)
    legend.background = element_rect(fill = alpha("white", 0.7)), # semi-transparent bg
    legend.title = element_blank(),
    legend.text = element_text(size = 12), 
    panel.grid.minor = element_blank(), 
    strip.text = element_text(size = 12),  
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12)
  )
p1

#ggsave(here("figures/age_based_prevalence_byurbanicity.tiff"), plot = p1, height = 10, width = 5)
```
## By assay 
```{r}
## PCR 
p2 <- pcr_sp %>% filter(assay == "PCR") %>% 
  ggplot(aes(
    x = timepoint, 
    y = prop,  
    color = community_type2,   # distinguishes the lines
    group = community_type2    # ensures separate lines
  )) + 
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("#510000", "#AC112D", "#EC6D47")) +
  scale_fill_manual(values = c( "#510000", "#AC112D", "#EC6D47")) +
  scale_x_continuous(breaks = seq(0, 24, by = 3)) +
  coord_cartesian(xlim = c(3, 24), ylim = c(0, 85)) +
  labs(
    x = "Age (months)", 
    y = "Prevalence (%)", 
    title = "PCR prevalence"
  ) +
  # CI shading
  geom_ribbon(aes(
    ymin = ci_low,      # if prop is in percent
    ymax = ci_high, 
    fill = community_type2), alpha = 0.15, color = NA) + 
  theme_minimal() +
  facet_wrap(~ pathogen, ncol = 3, drop = FALSE) + 
  theme(base_size = 12, 
        legend.position = c(0.075,0.9),  # x=0.8 (right), y=0.2 (bottom)
    legend.background = element_rect(fill = alpha("white", 0.7)), # semi-transparent bg
    legend.title = element_blank(),
    legend.text = element_text(size = 12), 
    panel.grid.minor = element_blank(), 
    strip.text = element_text(size = 12),  
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12)
  )
p2

## IgG
p3 <- pcr_sp %>% filter(assay == "IgG") %>% 
  ggplot(aes(
    x = timepoint, 
    y = prop,  
    color = community_type2,   # distinguishes the lines
    group = community_type2    # ensures separate lines
  )) + 
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("#510000", "#AC112D", "#EC6D47")) +
  scale_fill_manual(values = c( "#510000", "#AC112D", "#EC6D47")) +
  scale_x_continuous(breaks = seq(0, 24, by = 3)) +
  coord_cartesian(xlim = c(3, 24), ylim = c(0, 85)) +
  labs(
    x = "Age (months)", 
    y = "IgG Prevalence (%)", 
    title = NULL
  ) +
  # CI shading
  geom_ribbon(aes(
    ymin = ci_low,      # if prop is in percent
    ymax = ci_high, 
    fill = community_type2), alpha = 0.15, color = NA) + 
  theme_minimal() +
  facet_wrap(~ pathogen, ncol = 3, drop = FALSE) + 
  theme(base_size = 12, 
        legend.position = c(0.9,0.9),  # x=0.8 (right), y=0.2 (bottom)
    legend.background = element_rect(fill = alpha("white", 0.7)), # semi-transparent bg
    legend.title = element_blank(),
    legend.text = element_text(size = 12), 
    panel.grid.minor = element_blank(), 
    strip.text = element_text(size = 12),  
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12)
  )
p3

# save figures 
#ggsave(here("figures/age_based_pcr_prevalence_byurbanicity.tiff"), plot = p2, height = 8, width = 10)
ggsave(here("figures/age_based_igg_prevalence_byurbanicity.tiff"), plot = p3, height = 8, width = 10)

```


