---
title: "Giardia, PCR+ analysis"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

## File overview

The purpose of this file is to further investigate the discrepancy in
Giardia PCR and IgG responses for the ECoMiD cohort.

In short, we found a larger than average proportion of PCR+ samples were
IgG- for Giardia compared to all other pathogens in the panel.

The two plausible explanations we would like to investigate whether:

1.  **PCR is overestimating** (different target than serology, perhaps
    capturing Giardia strain incidence that is not captured by serology)

2.  **Luminex IgG is underestimating** (not enough exposure to
    seroconvert, bead issue, ect)

**Objectives of the analysis are**:

1.  For PCR+ data, visualize the raw TAC ct values for PCR results
    (borderline results).
2.  For PCR+ data, visualize IgG MFI values for IgG results (borderline
    results).

**Important to note:**

1.  Childsets are coded PCR/IgG status at the **antigen** level from
    10-ecomid_spaghetti_plots.qmd

2.  Childsets are coded conditional on a PCR+ result occurring before
    the *last occurrence* of an IgG result. This means if the PCR result
    was positive only at 24 months, the childset will be coded as PCR-.
    It also means if the only PCR+ result was at or after the last
    occurrence of an IgG result, the childset will be coded as PCR- (see
    table below). In other words, to classify children, PCR+ results
    were required to occur before the final available serology
    measurement. PCR+ detections observed only at the last serology
    timepoint were excluded, regardless if they were IgG+ or IgG-,
    because no subsequent IgG follow-up was available to assess
    seroconversion.

3.  Due to exclusion criteria: 20 childsets were coded as PCR-, despite
    having a PCR+ result for VSP5_Giardia.

| PCR status | IgG status | Coded as... |
|------------------------|------------------------|------------------------|
| +, at any timepoint before last IgG result | +, at any timepoint | PCR+IgG+ |
| +, at any timepoint before last IgG result | -, at all timepoints | PCR+IgG- |
| -, at all timepoints before last IgG result | +, at any timepoint | PCR-IgG+ |
| -, at all timepoints before last IgG result | -, at all timepoints | PCR-IgG- |

: *PCR+ only counts if it occurs strictly before the last recorded IgG
timepoint, or if a PCR+ at the last IgG timepoint is accompanied by at
least one earlier PCR+ event*

## System setup

```{r}
# Read in files 
library(here)
source(here("R", "0-config.R"))
```

## Read in files

```{r}
# seropositivity individual level
# don't include MFI to avoid confusion with GPP MFI down the line!
sero_pos <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_2026-05-19.csv")) %>% filter(antigen != "hyle") %>% select(-mfi, -log10mfi, -mficut, -mficut_desc) %>% 
  mutate(community_type2 = case_when(community_type == "Rural Road" ~ "Rural", 
                                    community_type == "Rural River" ~ "Rural", 
                                    community_type == "Borbon" ~ "Borbon", 
                                    community_type == "Esmeraldas" ~ "Esmeraldas City"))


# pcr individual level
pcr <- readRDS(here("data/ecomid_pcr_analysis_data_2026-06-11.rds")) %>% ungroup() %>% 
  rename(hhid = childid, 
         sampleid = sample_ID) %>% 
  mutate(childid = substr(sampleid, start = 1, stop = 6))

pcr$community_type <- as.character(pcr$community_type)
pcr <- pcr %>% 
  mutate(community_type2 = case_when(community_type == "Rural Road" ~ "Rural", 
                                    community_type == "Rural River" ~ "Rural", 
                                    community_type == "Borbon" ~ "Borbon", 
                                    community_type == "Esmeraldas" ~ "Esmeraldas City")) %>% rename(timepoint = visit)
pcr$community_type2 <- factor(pcr$community_type2, levels = c("Esmeraldas City", "Borbon", "Rural")) 

# filter it down to the 370 children that were in both cohorts
pcr <- pcr %>% filter(childid %in% sero_pos$childid) 

table(pcr$timepoint)
table(pcr$pcr_assay)

# tac raw results (ct values)
tac <- read_csv(here("../../Data/tac-full/final/TAC_data_pathogensbysamples_CT_20260316.csv")) %>% 
  select(-`...1`, -X, -Collection.Date, -Collection.Month, -Community.type, -Age, -Community, -season, -sex, -HH.ID)

# Luminex PCR: log calls
gpp <- read_rds(here("../../Data/gpp-full/final/gpp_cleaned_files/gpp_final_calls_clean_20260610.rds")) %>% mutate(gpp_target = ifelse(gpp_target == "ETEC_LT", "LT-ETEC", gpp_target)) %>% mutate(timepoint = as.numeric(age)) %>% rename(sampleid = sample)

# Assay categorization (PCR/IgG)
assay <- read_rds(here("output/ecomid_seropos_pcr_alltargets.rds"))
```

## 1) Giardia

### A) TAC: CT values

Data clean: merge the ct values

```{r}
# create merging variable to pcr dataset 
pcr2 <- pcr %>% mutate(id_tp = 
                         paste0(substr(childid, start = 3, stop = 6), 
                                sprintf("%02d", timepoint)))

# pivot tac ct values longer for joining
tac_long <- tac %>%
  pivot_longer(cols = 2:48, names_to = "tac_target", 
               values_to = "ct_value")

# filter down to Giardia 
tac_long <- tac_long %>% filter(tac_target == "Giardia_18S") %>% mutate(pathogen = "Giardia")

# join to Giardia PCR records
pcr_ct <- left_join(pcr2, tac_long, by = c("id_tp", "pathogen")) %>% 
  filter(pathogen == "Giardia", pcr_assay == "tac")  

# merge in IgG seroposivity at each timepoint
igg_ct <- left_join(pcr_ct, 
                    sero_pos %>% select(childid, timepoint, seropos, pathogen) %>% 
                      filter(pathogen == "Giardia"), 
                    by = c("childid", "timepoint", "pathogen"))

# how much missing data (at limit of detection?)
print("No. unique children with CT value and IgG value")
length(unique(igg_ct$childid))

print("CT value, NAs")
table(is.na(igg_ct$ct_value))

print("Missing data: CT values")
igg_ct %>% 
  summarise(
    total_rows = n(),
    n_na = sum(is.na(ct_value)), 
    percent_missing = round(n_na / total_rows * 100, 2)
  )

print("Missing data: IgG")
igg_ct %>% 
  summarise(
    total_rows = n(),
    n_na = sum(is.na(seropos)), 
    percent_missing = round(n_na / total_rows * 100, 2)
  )
```

Many NA values for CT values is likely due to limit of detection for the
assay \*will be for the 6, 12, 18 months samples

```{r}
# label seroposivity and TAC positivity
igg_ct <- igg_ct %>% 
  mutate(seropos2 = ifelse(seropos == 1, "IgG+", "IgG-")) 

unique(igg_ct$timepoint)

igg_ct <- igg_ct %>% mutate(tacpos = ifelse(pcrpos == 1, "TAC+", "TAC-"))

table(igg_ct$tacpos, igg_ct$seropos2)
igg_ct$ct_value <- as.numeric(igg_ct$ct_value)
```

```{r}

# Add in assay categorization that was completed in 10-ecomid_mfi_spaghetti_plots.qmd

# How many of these PCR+ are missing a tac ct value (should be none)
igg_ct %>% 
  filter(pcrpos == 1) %>%
  summarise(
    total_rows = n(),
    n_na = sum(is.na(ct_value)), 
    percent_missing = round(n_na / total_rows * 100, 2)
  )

# merge in childset-level PCR/IgG classification from assay object
igg_ct2 <- igg_ct %>% ungroup() %>% 
  left_join(assay %>% distinct(childid, pathogen, assay_pos), 
            by = c("childid", "pathogen")) %>% 
  filter(pcrpos == 1, 
         assay_pos == "PCR+ IgG+" | 
           assay_pos == "PCR+ IgG-") # filter down to those categorized in the spaghetti plots as PCR+ 
         
# duplicate data and label as "Overall"
igg_ct_overall <- igg_ct2 %>%
  mutate(community_type2 = "Overall")

# combine with original community level data
igg_ct_combined <- bind_rows(igg_ct2, igg_ct_overall)

# assign levels
igg_ct_combined$community_type2 <- 
  factor(igg_ct_combined$community_type2, 
         levels = c("Esmeraldas City", "Borbon", "Rural", "Overall"))

#------------#
# Code Check #
#------------#

# total TAC+ results before assay_pos filter
cat("Total TAC+ results before assay_pos filter")
igg_ct %>% filter(pcrpos == 1) %>% nrow()

# TAC+ results after filtering to PCR+ IgG+ and PCR+ IgG- only
cat("TAC+ results after assay_pos filter")
igg_ct2 %>% nrow()

cat("Total TAC+ data excluded")
(igg_ct %>% filter(pcrpos == 1) %>% nrow()) - (igg_ct2 %>% nrow())
```

#### i) Mann-Whitney U test

```{r}
# repeated measures: number of children with a PCR+ TAC value

cat("Number of children with repeated measures")
igg_ct_combined %>% group_by(childid) %>% 
  summarise(n_tacpos = n()) %>% 
  group_by(n_tacpos) %>% 
  summarise(n_children = n())

# counts for the plot
igg_ct_combined %>% 
  distinct(childid, assay_pos) %>% 
  count(assay_pos) %>% 
  print()

# Community-specific Mann-Whitney 
igg_ct_combined$assay_pos <- factor(igg_ct_combined$assay_pos, 
                                    levels = c("PCR+ IgG-", "PCR+ IgG+"))
levels(igg_ct_combined$assay_pos)

mannwhitney_ct <- igg_ct_combined %>%
  filter(!is.na(ct_value), !is.na(assay_pos)) %>%
  group_by(community_type2) %>%
  summarise(
    test = list(wilcox.test(ct_value ~ assay_pos, conf.int = TRUE)),
    total_n = n(),
    n_IgG_neg = sum(assay_pos == "PCR+ IgG-"),
    n_IgG_pos = sum(assay_pos == "PCR+ IgG+")
  ) %>%
  mutate(
    W = sapply(test, \(x) as.numeric(x$statistic)),
    U = W - (n_IgG_neg * (n_IgG_neg + 1)) / 2,
    p_value = round(sapply(test, \(x) x$p.value), 5),
    conf_low = sapply(test, \(x) x$conf.int[1]),
    conf_high = sapply(test, \(x) x$conf.int[2])
  ) %>%
  select(-test) %>%
  mutate(p_value_print = ifelse(p_value < 0.001, "p < 0.001", paste0("p = ", sprintf("%1.3f", p_value))))

mannwhitney_ct
```

#### ii) CT plot

The analysis here is done for IgG status at the entire childset level
(over the course of the study) and is subset to timepoints when TAC was
positive

```{r}
set.seed(12345) # set seed for consistent point jitter

tac_giardia <- ggplot(
  igg_ct_combined,
  aes(x = assay_pos, y = ct_value)
) + 
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  geom_jitter(
    aes(color = factor(community_type2)),
    alpha = 0.75, position = position_jitter(width = 0.2, height = 0, seed = 12345)
  ) + 
  geom_hline(yintercept = 35, color = "black", linetype = "dashed") +
  facet_grid(~ community_type2) +
  scale_y_reverse(
    limits = c(10, 40),
    breaks = seq(10, 40, 5)
  ) +
  geom_text(data = mannwhitney_ct, aes(x = 1.5, y = 10, label = p_value_print)) + 
  scale_x_discrete(labels = c("IgG-", "IgG+")) +
  scale_color_manual(
    values = c(
      "Esmeraldas City" = nytpal[1],
      "Borbon" = nytpal[2],
      "Rural" = nytpal[3],
      "Overall" = "gray30"
    )
  ) +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "none"
  ) + 
  labs(
    subtitle = "Giardia 18S Ct value for PCR positive samples by TAC (ages 6, 12, 18 months)",
    y = "TAC Ct value", x = "child seroconversion status",
    color = ""
  ) 

tac_giardia <- tac_giardia +
  labs(tag = "A)") +
  theme(
    plot.tag = element_text(size = 14, face = "bold"),
    plot.tag.position = c(0.01, 1)
  )
tac_giardia

ggsave(here("figures/giardia_ct_values_4panel.tiff"), tac_giardia, height=5, width=10)
```

### B) Luminex: MFI values

24 month removed for Luminex

```{r}
# create mfi df
# note: some pathogens have multiple PCR targets so the merge is not 1-to-1;
# Giardia has a single GPP target so this is fine here

gpp <- gpp %>% rename(id_tp = id_time)
                      
pcr_mfi <- left_join(pcr2 %>% 
                       filter(pcr_assay == "gpp") %>% 
                       select(id_tp, pathogen, pcrpos, timepoint, childid), 
                     gpp %>% select(-childid), 
                     by = c("id_tp", "pathogen", "timepoint")) %>% 
  filter(pathogen == "Giardia") 

# join in the seropositivity and community data; fill in community for NA rows due to 3 month and 15 month missing in serology 
igg_mfi <- left_join(pcr_mfi %>% ungroup(), 
                     sero_pos %>% ungroup() %>%
                       filter(pathogen == "Giardia", antigen == "vsp5") %>%
                       select(childid, timepoint, pathogen, seropos) %>% 
                       distinct(),
                     by = c("childid", "timepoint", "pathogen"))

# join in community 
igg_mfi <- left_join(igg_mfi, sero_pos %>% distinct(childid, community, community_name, 
                              community_type2), by = c("childid"))
 
igg_mfi$community_type2 <- factor(igg_mfi$community_type2, levels = c("Esmeraldas City", "Borbon", "Rural"))

 # label seroposivity and GPP positivity 
igg_mfi <- igg_mfi %>% 
  mutate(seropos2 = ifelse(seropos == 1, "IgG+", "IgG-")) 

unique(igg_mfi$timepoint)

igg_mfi <- igg_mfi %>% mutate(gpppos = ifelse(pcrpos == 1, "GPP+", "GPP-"))
table(igg_mfi$gpppos, igg_mfi$seropos2)
igg_mfi$mfi <- as.numeric(igg_mfi$mfi)

print("Missing data: PCR MFI values")
igg_mfi %>% 
  summarise(
    total_rows = n(),
    n_mfi_na = sum(is.na(mfi)), 
    percent_missing = round(n_mfi_na / total_rows * 100, 2)
  )

# merge in the assay classification 
igg_mfi <- igg_mfi %>% ungroup() %>% 
  left_join(assay %>% distinct(childid, pathogen, assay_pos), 
                     by = c("childid", "pathogen"))

# filter to GPP+ samples
# assay_pos status was already assigned excluding the 24 month timepoint/end timepoints
igg_mfi2 <- igg_mfi %>% 
  filter(pcrpos == 1, 
         assay_pos == "PCR+ IgG+" | 
           assay_pos == "PCR+ IgG-")

# duplicate data and label as "Overall"
igg_mfi_overall <- igg_mfi2 %>%
  mutate(community_type2 = "Overall")

# combine with original community level data
igg_mfi_combined <- bind_rows(igg_mfi2, igg_mfi_overall)

# assign levels
igg_mfi_combined$community_type2 <- 
  factor(igg_mfi_combined$community_type2, 
         levels = c("Esmeraldas City", "Borbon", "Rural", "Overall"))

#------------#
# Code Check #
#------------#

# total GPP+ results before assay_pos filter
cat("Total GPP+ results before assay_pos filter")
igg_mfi %>% filter(pcrpos == 1) %>% nrow()

# GPP+ results after filtering to PCR+ IgG+ and PCR+ IgG- only
cat("GPP+ results after assay_pos filter")
igg_mfi2 %>% nrow()

cat("Total GPP+ data excluded")
(igg_mfi %>% filter(pcrpos == 1) %>% nrow()) - (igg_mfi2 %>% nrow())
```

#### i) Mann-Whitney U test

```{r}
# get table of sample distribution 
igg_mfi2 %>% group_by(timepoint) %>% 
  filter(gpp_target_pos == "POS") %>% 
  summarise(n_gpppos = n())
# note that 24 month GPP+ samples are included if they had a previous GPP+ result

# repeated measures: GPP+ timepoints coded as GPP+ after exclusion criteria applied
igg_mfi2 %>% group_by(childid) %>% 
  filter(gpp_target_pos == "POS") %>% 
  summarise(n_pcrpos = n()) %>% 
  group_by(n_pcrpos) %>% 
  summarise(n_children = n())


###-------------------------------------------------------------------###
## Mann-Whitney U test (ignoring repeated measures and community type) ##
###-------------------------------------------------------------------###
igg_mfi_combined$assay_pos <- factor(igg_mfi_combined$assay_pos, levels = c("PCR+ IgG-", "PCR+ IgG+"))
levels(igg_mfi_combined$assay_pos)

cat("Mann Whitney U test, by community, MFI")
mannwhitney_mfi <- igg_mfi_combined %>%
  filter(!is.na(mfi), !is.na(assay_pos)) %>%
  group_by(community_type2) %>%
  summarise(
    test = list(tryCatch(
      wilcox.test(mfi ~ assay_pos, conf.int = TRUE),
      error = function(e) NULL
    )),
    total_n = n(),
    n_IgG_neg = sum(assay_pos == "PCR+ IgG-"),
    n_IgG_pos = sum(assay_pos == "PCR+ IgG+")
  ) %>%
  mutate(
    W          = sapply(test, \(x) if (is.null(x)) NA_real_ else as.numeric(x$statistic)),
    p_value    = sapply(test, \(x) if (is.null(x)) NA_real_ else round(x$p.value, 5)),
    conf_low   = sapply(test, \(x) if (is.null(x)) NA_real_ else x$conf.int[1]),
    conf_high  = sapply(test, \(x) if (is.null(x)) NA_real_ else x$conf.int[2])
  ) %>%
  select(-test) %>%
  mutate(p_value_print = case_when(
    is.na(p_value)   ~ "N/A",
    p_value < 0.001  ~ "p < 0.001",
    TRUE             ~ paste0("p = ", sprintf("%1.3f", p_value))
  ))

mannwhitney_mfi
```

#### ii) MFI plot

```{r}
set.seed(12335) # set seed for consistent jitter

gpp_giardia <- ggplot(
  igg_mfi_combined,
  aes(x = assay_pos, y = log10(mfi))
) + 
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  geom_jitter(
    aes(color = community_type2),
    alpha = 0.75, position = position_jitter(width = 0.2, height = 0, seed = 12345)
  ) + 
  geom_hline(yintercept = log10(200), color = "black", linetype = "dashed") +
  geom_text(data = mannwhitney_mfi, aes(x = 1.5, y = 4, label = p_value_print)) + 
  facet_grid(~ community_type2) +
  scale_y_continuous(
    limits = c(2, 4),
    breaks = seq(2, 4, 0.5)
  ) +
  scale_x_discrete(labels = c("IgG-", "IgG+")) +
  scale_color_manual(
    values = c(
      "Esmeraldas City" = nytpal[1],
      "Borbon" = nytpal[2],
      "Rural" = nytpal[3],
      "Overall" = "gray30"
    )
  ) +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    legend.position = "none"
  ) + 
  labs(
    subtitle = "Giardia MFI values for PCR positive samples by Luminex GPP (ages 3, 9, 15, 24 months)",
    y = "Luminex log10(MFI)", x = "child seroconversion status"
  )

gpp_giardia <- gpp_giardia +
  labs(tag = "B)") +
  theme(
    plot.tag = element_text(size = 14, face = "bold"),
    plot.tag.position = c(0.01, 1)
  )

gpp_giardia
```

### C) Aggregate figure: tac and gpp

```{r fig.height = 7, fig.width = 6}
agg_giardia <- plot_grid(tac_giardia, gpp_giardia, ncol = 1)
agg_giardia

ggsave(here("figures/supp_giardia_mfi_ct_byurbanicity_8panel.tiff"), agg_giardia, device = "tiff", height = 180, width = 200, units = "mm")
```
