---
title: "EcoMiD- measures of association"
subtitle: "MoA of serological and PCR results across community types"
author: "Nikolina Walas (nikolina.walas@ucsf.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
---

## Overview

This file calculates the measure of association between community types (Urban-Esmeraldas, Intermediate-Borbon, Rural- 8 communities across rural river and rural road settings) for outcomes as outlined below:

+----------------------------------+------------------------------------------+-------------------------------------------------------------+----------------------------------------+------------------+
| Outcome measure                  | Method for MoA                           | Method for CI                                               | Confounders                            | Interpretation   |
+==================================+==========================================+=============================================================+========================================+==================+
| PCR prevalence                   | 1\) Log-binomial regression              | Cluster-robust effect standard error (clustered at childid) | Not of interest: looking at net effect | Prevalence ratio |
|                                  |                                          |                                                             |                                        |                  |
|                                  | 2\) Modified poisson model with log link |                                                             |                                        |                  |
+----------------------------------+------------------------------------------+-------------------------------------------------------------+----------------------------------------+------------------+
| Seroconversion rate (per 100 py) | Poisson model                            | Cluster-robust effect standard error                        | Not of interest: looking at net effect | SCR ratio        |
+----------------------------------+------------------------------------------+-------------------------------------------------------------+----------------------------------------+------------------+

## Set up

Load data and packages

```{r}
library(here)
source(here("R", "0-config.R"))
```

Read in files

```{r}
#----------------
# Summary tables
#----------------
# seroconversion by community type + pathogen
sero_inc <- readRDS(here("output/ecomid_pathogen_comm_sero_inc_3_levels.rds"))
sero_inc$community_type2 <- factor(sero_inc$community_type2, levels = c("Esmeraldas", "Borbon", "Rural"))

# seroprevalence by community type + pathogen
sero_prev <- readRDS(here("output/ecomid_sero_prev_rates_3_levels.rds"))
sero_prev$community_type2 <- factor(sero_prev$community_type2, levels = c("Esmeraldas", "Borbon", "Rural"))

# pcr prevalence by community type + pathogen
pcr_prev <- readRDS(here("output/ecomid_pcr_prev_by_community_3_levels.rds"))
pcr_prev$community_type2 <- factor(pcr_prev$community_type2, levels = c("Esmeraldas", "Borbon", "Rural"))

#--------------
# Individual level data
#--------------
# pcr readout
pcr <- readRDS(here("data/ecomid_pcr_analysis_data_2026-08-12.rds")) %>% ungroup() 
pcr$community_type <- as.character(pcr$community_type)
pcr <- pcr %>% mutate(community_type = ifelse(str_detect(community_type, "Rural"), "Rural", community_type))
pcr$community_type <- factor(pcr$community_type, levels = c("Esmeraldas", "Borbon", "Rural"))

# seroconversion table
sero_conv <- readRDS(here("output/ecomid_sero_conv_tbl.rds"))
here()
```

## PCR Prevalence

PCR positivity was calculated according to the xTAG GPP and TAC assay guidelines

```{r}
#| fig-width: 8
#| fig-height: 6

#---------------------
# Overview of data 
#---------------------
summary(pcr_prev)
dim(pcr_prev)
table(is.na(pcr_prev$n_pos))

#----------------------
# Plot PCR prevalence per community across pathogens 
#----------------------

pcr_prev %>% group_by(community_type2) %>% ggplot(aes(x = pathogen, color = community_type2, group = community_type2)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
                position = position_dodge(width = 0.6),
                width = 0.1) +
  geom_point(aes(y= pcr_prevalence), position = position_dodge(width = 0.6),
             size = 2)  +
  scale_color_manual(values=nytpal) +
  scale_y_continuous(breaks=seq(0,0.3,by=0.05),labels=seq(0,0.3,by=0.05)*100) +
  #scale_x_discrete(labels = wrapped_labels) + 
  coord_cartesian(ylim = c(0,0.3)) +
  labs(x="",y="Prevalence (%)", 
       color = "Community", 
       title = "PCR prevalence across communities") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        text = element_text(size = 12), 
        legend.position = c(0.95, 0.95), 
        legend.justification = c("right", "top"), 
        legend.background = element_rect(fill = "white", color = "black", size = 0.2)) 
```

Compare groups using pcr prevalence

Confidence interval was calculated using robust standard errors. Code for the calculation can be found in file 3-ecomid: pcr prevalence

```{r}
#-----------------
# PCR prevalence by community and pathogen 
#-----------------
pcr_prev %>% rename(community = community_type2) %>% select(pathogen, community, n_total, n_pos, pcr_prevalence, lower_ci, upper_ci) %>% kbl(digits=3, caption = "PCR prevalence by community and pathogen") %>%
  kable_styling(bootstrap_options = "striped")
```

### Log-binomial

```{r}
#----------------------------------
# estimate the difference in 
# pcr prevalence using a log-binomial model
# with SEs from model 
# save estimates, 95% CI, and P-value
#----------------------------------
# Log Binomial Model 
fit_results <- pcr %>%
  group_by(pathogen) %>%
  do({
    fit <- glm(pcrpos ~ community_type,
               data = .,
               family = binomial(link = "log"))
    tidy(fit, conf.int = TRUE, exponentiate = TRUE)
  }) %>%
  ungroup()
fit_results

#---------
# Log-binomial plot
#--------

fit_results %>% filter(term != "(Intercept)") %>% 
  mutate(community = ifelse(term == "community_typeBorbon", 
                            "Esmeraldas-Borbon", 
                            "Esmereldas-Rural")) %>% 
  group_by(community) %>% 
  ggplot(aes(x = pathogen, color = community, group = community)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = 0.6),
                width = 0.1) +
  geom_point(aes(y= estimate), position = position_dodge(width = 0.6),
             size = 2)  +
  scale_color_manual(values=nytpal[-1]) +
  scale_y_continuous(breaks=seq(-1,5,by=0.5),labels=seq(-1,5,by=0.5)) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "black")  +
  labs(x="",y="Prevalence Ratio", 
       color = "Community", 
       title = "PCR prevalence ratio across community types", 
       subtitle = "model = log-binomial") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        text = element_text(size = 12), 
        legend.position = c(0.95, 0.95), 
        legend.justification = c("right", "top"), 
        legend.background = element_rect(fill = "white", color = "black", size = 0.2)) 

```

### Modified poisson

```{r}
#-----------------------
# Modified poisson model 
# with robust (sandwich) estimated standard errors; clustered at childid
#-----------------------
fit_results2 <- pcr %>%
  group_by(pathogen) %>%
  group_modify(~ {
    dat <- .x

    # require at least 2 clusters
    if (n_distinct(dat$childid) < 2) {
      return(tibble(
        term = NA_character_,
        estimate = NA_real_,
        conf.low = NA_real_,
        conf.high = NA_real_
      ))
    }

    # modified Poisson (log link)
    fit <- glm(pcrpos ~ community_type,
               data = dat,
               family = poisson(link = "log"))

    # clustered sandwich using the group's child ids
    vcov_cl <- sandwich::vcovCL(fit, cluster = dat$childid)

    coefs <- coef(fit)
    se <- sqrt(diag(vcov_cl))
    z <- qnorm(0.975)

    tibble(
      term = names(coefs),
      estimate = exp(coefs),                         # PRs
      conf.low = exp(coefs - z * se),
      conf.high = exp(coefs + z * se),
      p.value = 2 * pnorm(-abs(coefs / se))          # optional
    )
  }) %>%
  ungroup() %>%
  filter(term != "(Intercept)")

summary(fit_results2)

#------------
# Modified Poisson
#------------

fit_results2 %>%  
  mutate(community = ifelse(term == "community_typeBorbon", 
                            "Esmeraldas-Borbon", 
                            "Esmeraldas-Rural")) %>% 
  group_by(community) %>% 
  ggplot(aes(x = pathogen, color = community, group = community)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = 0.6),
                width = 0.1) +
  geom_point(aes(y= estimate), position = position_dodge(width = 0.6),
             size = 2)  +
  scale_color_manual(values=nytpal[-1]) +
  scale_y_continuous(breaks=seq(-1,5,by=0.5),labels=seq(-1,5,by=0.5)) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "black") + 
  #coord_flip(ylim = c(0,5)) +
  labs(x="",y="Prevalence Ratio", 
       color = "Community", 
       title = "PCR prevalence ratio across community types", 
       subtitle = "model = modified poisson") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        text = element_text(size = 12), 
        legend.position = c(0.95, 0.95), 
        legend.justification = c("right", "top"), 
        legend.background = element_rect(fill = "white", color = "black", size = 0.2)) 

#-----------
# Table 1: PCR prevalence ratios 
#-----------

colnames(fit_results2)

# combine the estimate with the confidence interval
fit_results2$merged <- paste0(round(fit_results2$estimate, 2), " (", round(fit_results2$conf.low, 2),", ", round(fit_results2$conf.high, 2), ")")

# reformat for table 
tbl <- fit_results2 %>% 
  select(pathogen, term, p.value, merged) %>% 
  mutate(p.value = round(p.value, 2), 
         term = ifelse(term == "community_typeBorbon", "Borbon", "Rural")) %>% # round the p-value
  pivot_wider(names_from = term, values_from = c(merged, p.value)) %>% 
  rename(PR_Borbon = merged_Borbon, PR_Rural = merged_Rural) 

# add in community prevalence measures 
add <- pcr_prev %>% rename(community = community_type2) %>% select(pathogen, community, pcr_prevalence, lower_ci, upper_ci) %>% 
  mutate(merged = paste0(round(pcr_prevalence, 3)*100, 
                         " (", round(lower_ci, 3)*100, 
                         ", ", round(upper_ci, 3)*100, ")")) %>% 
  select(pathogen, community, merged) %>% 
  pivot_wider(names_from = community, values_from = merged) %>% select(pathogen, Esmeraldas, Borbon, Rural) 

kable(left_join(tbl, add, by = "pathogen") %>% 
        select(pathogen, Esmeraldas, Borbon, Rural,
               PR_Borbon, p.value_Borbon, PR_Rural, 
               p.value_Rural), 
      caption = "PCR Prevalence", 
      col.names = c("Pathogen", "Esmeraldas (Ref group)", 
                    "Borbon", "Rural", "PR- Borbon (95 CI)", 
                    "p-value", "PR- Rural (95 CI)", "p-value")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE, 
                position = "left", font_size = 14) %>%
  add_header_above(c(" " = 1, "Prevalence (%), (95 CI)" = 3, " " = 4))
```

## SCR ratio

For seroconversion rate ratio; poisson model with robust SE clustered at child-id level was used.

```{r}
sero_conv
sum(is.na(sero_conv$seroconversion))

# overview of SCR (excluding those no longer at risk)
sero_rates <- sero_conv %>%
  filter(atrisk == 1, !is.na(seroconversion), person_time_years > 0) %>% # at risk only 
  group_by(pathogen, community_type) %>%
  summarise(
    total_events = sum(seroconversion),
    total_person_time = sum(person_time_years),
    rate_per_100py = (total_events / total_person_time) * 100,
    .groups = "drop"
  )

#=================================
# Table 2: Serologicial SCR
#=================================
sero_inc2 <- sero_inc %>% arrange(pathogen, community_type2)
sero_inc2 <- sero_inc2 %>% 
  mutate(merged_ir = paste0(round(ir, 2), " (", round(ir_lo, 2), ", ", round(ir_hi, 2), ")"),
         merged_ir100 = paste0(round(ir_per_100, 2), " (", round(ir_lo_per_100, 2), " ,", 
         round(ir_hi_per_100, 2), ")"), 
         total_py = round(total_py, 2)) %>% select(pathogen, community_type2, total_events, n_children, total_py, merged_ir100) 

sero_inc2 %>% kable(caption = "Seroconversion Rates", 
      col.names = c("Pathogen", "Community", "Total Events", "Children (N)", "Total person-years", "Seroconversion rate (per 100 py); (95CI)")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE, 
                position = "left", font_size = 14)

#====================
# Model Fit
#====================

df <- sero_conv %>%
  filter(atrisk != 0) # filter those that are no longer at risk (lead_seropositive = 1)
df <- df %>%
  filter(!is.na(seroconversion)) #filter out NA values for seroconversion (usually 24 months)
df <- df %>%
  filter(person_time_years > 0) # filter out 0 person-time contribution (usually 24 months)

# fit to model for SCRR
fit_results_scrr <- df %>%
  group_by(pathogen) %>%
  group_modify(~ {
    dat <- .x

    # require at least 2 clusters
    if (n_distinct(dat$childid) < 2) {
      return(tibble(
        term = NA_character_,
        estimate = NA_real_,
        conf.low = NA_real_,
        conf.high = NA_real_
      ))
    }

    # modified Poisson (log link)
    fit <- glm(seroconversion ~ community_type,
               data = dat,
               family = poisson(link = "log"), 
               offset = log(person_time_years))

    # clustered sandwich using the group's child ids
    vcov_cl <- sandwich::vcovCL(fit, cluster = dat$childid)

    coefs <- coef(fit)
    se <- sqrt(diag(vcov_cl))
    z <- qnorm(0.975)

    tibble(
      term = names(coefs),
      estimate = exp(coefs),                         # PRs
      conf.low = exp(coefs - z * se),
      conf.high = exp(coefs + z * se),
      p.value = 2 * pnorm(-abs(coefs / se))          # optional
    )
  }) %>%
  ungroup()

# make smaller table 
fit_results_scrr %>%
  mutate(
    estimate = round(estimate, 2),
    conf.low = round(conf.low, 2),
    conf.high = round(conf.high, 2),
    p.value = signif(p.value, 3)
  ) %>% 
  mutate(p.value = round(p.value, 2))

# make merged estimate + CI column                                
fit_results_scrr$merged <- paste0(round(fit_results_scrr$estimate, 2), " (", round(fit_results_scrr$conf.low, 2),", ", round(fit_results_scrr$conf.high, 2), ")")

```

Plot results and make a table as well

```{r}
#----------------
# Plot- SCR Ratio 
#----------------
fit_results_scrr %>%  
  mutate(community = ifelse(term == "community_typeBorbon", 
                            "Esmeraldas-Borbon", 
                            "Esmereldas-Rural")) %>% 
  group_by(community) %>% filter(term != "(Intercept)") %>% 
  ggplot(aes(x = pathogen, color = community, group = community)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = 0.6),
                width = 0.1) +
  geom_point(aes(y= estimate), position = position_dodge(width = 0.6),
             size = 2)  +
  coord_cartesian(ylim = c(0,22)) +
  scale_color_manual(values=nytpal[-1]) +
  scale_y_continuous(breaks=seq(0,22,by=2),labels=seq(0,22,by=2)) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "black") + 
  labs(x="",y="Seroconversion Rate Ratio", 
       color = "Community", 
       title = "SCR ratio across community types", 
       subtitle = "model = poisson") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        text = element_text(size = 12), 
        legend.position = c(0.95, 0.95), 
        legend.justification = c("right", "top"), 
        legend.background = element_rect(fill = "white", color = "black", size = 0.2)) 

#------------------
# SCR ratio summary table  
#------------------

# reformat for table 
tbl2 <- fit_results_scrr %>% 
  filter(term != "(Intercept)") %>%
  select(pathogen, term, merged, p.value) %>% 
  mutate(p.value = round(p.value, 2), 
         term = ifelse(term == "community_typeBorbon", "Borbon", "Rural")) %>% # round the p-value
  pivot_wider(names_from = term, values_from = c(merged, p.value)) %>% 
  rename(SCRR_Borbon = merged_Borbon, SCRR_Rural = merged_Rural) 

# seroconversion rates table
add <- sero_rates %>% select(pathogen, community_type, rate_per_100py) %>% 
  mutate(rate_per_100py = round(rate_per_100py, 2)) %>% 
        pivot_wider(names_from = community_type, values_from = rate_per_100py)

kable(left_join(tbl2, add, by = "pathogen") %>% select(pathogen, Esmeraldas, Borbon, Rural, SCRR_Borbon, p.value_Borbon, SCRR_Rural, p.value_Rural), caption = "Seroconversion", 
      caption.short = "reference group is Esmeraldas", 
      col.names = c("Pathogen", "Esmeraldas (Ref group)", "Borbon", "Rural", "SCR Ratio- Borbon (95 CI)", "p-value", "SCR Ratio- Rural (95 CI)", "p-value")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE, 
                position = "left", font_size = 14) %>%
  add_header_above(c(" " = 1, "Seroconversion rate (per 100 py)" = 3, " " = 4))
```

## Export datasets

```{r}
# Seroconversion rate ratio
write.csv(fit_results_scrr, file = here("output/ecomid_scrr.csv"))
saveRDS(fit_results_scrr, file = here("output/ecomid_scrr.rds"))

# PCR prevalence ratio
write.csv(fit_results2, file = here("output/ecomid_pr.csv"))
saveRDS(fit_results2, file = here("output/ecomid_pr.rds"))

```
