---
title: "ECoMiD: measure of association figure generation"
author: "Nikolina Walas (nikolina.walas@ucsf.edu) and Julianna Colado (juliannacolado@berkeley.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true

execute:
  echo: true
---

## Overview

The purpose of this file is to create a large / aggregated figure for the manuscript that includes 4 panes per pathogen (raw SCR, SCRR, raw PCR Prevalence, PCRPrevR) across community types. Data used in the plots is generated in the pcr-serology workflow.

## Set up

Load data and packages

```{r}
library(here)
library(purrr)
source(here("R", "0-config.R"))
```

### Read in files

```{r}
#----------------
# Summary tables
#----------------
# seroconversion by community type + pathogen
sero_inc <- readRDS(here("output/ecomid_pathogen_comm_sero_inc_3_levels.rds")) %>% mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
sero_inc$community_type2 <- factor(sero_inc$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

# pcr prevalence by community type + pathogen
pcr_prev <- readRDS(here("output/ecomid_pcr_prev_by_community_3_levels.rds")) %>% mutate(community_type2 = ifelse(community_type2 == "Borbon", "Borbón", community_type2))
pcr_prev$community_type2 <- factor(pcr_prev$community_type2, levels = c("Esmeraldas", "Borbón", "Rural"))

#-----------------
# Measures of Association
#-----------------
# RR
scrr <- readRDS(here("output/ecomid_scrr.rds"))

# RR
prevr <- readRDS(here("output/ecomid_pr.rds"))


# set margins for the plots in the following code chunks
small_margins <- theme(
  plot.margin = margin(0, 0, 0, 0),  # top, right, bottom, left; 
  panel.spacing.y = unit(-2, "pt")
)

```

### 1) Seroconversion rate plots

```{r}
####--------------------------------####
####--------SCR PLOTS---------------####
####--------------------------------####

# get vector of unique pathogen names
pathogens <- unique(sero_inc$pathogen)
pathogens_clean <- str_replace_all(pathogens, " ", "_")
# create a named list of plots, one per pathogen
plot_list <- map(pathogens, function(p){

  sero_inc %>%
    filter(pathogen == p) %>%
    ggplot(aes(x = community_type2, y = ir_per_100, color = community_type2)) +
    geom_point() + 
    labs(
      y = "SCR (per 100 py)",
      x = "",
      title = paste(p) # title for the entire plot panel 
    ) +
    scale_color_manual(values = nytpal) +
    theme(
      #plot.title = element_text(hjust = -Inf), 
      axis.text.x = element_blank(),
      legend.position = "right",
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(color = "grey", linewidth = 0.1), 
      plot.margin = margin(0, 0, 0, 0),
      panel.spacing.y = unit(0, "pt"),
      strip.background = element_blank(),
      strip.text = element_text(margin = margin(0, 0, 0, 0))
    ) +
    geom_errorbar(aes(ymin = ir_lo_per_100, ymax = ir_hi_per_100), width = 0.1, linewidth = 0.5) + theme(
        axis.text.x = element_blank(),
        text = element_text(size = 12),
        legend.position = "none"
      ) +
  coord_fixed(ratio = 0.02, ylim = c(0, 200))  + small_margins

}) 

# name each list element using scr_[pathogen]
names(plot_list) <- paste0("scr_", pathogens_clean)

# Optionally assign each plot as an object in the global environment:
list2env(plot_list, envir = .GlobalEnv)

```

### 2) PCR Prevalence Plots

```{r}
####--------------------------------####
####--------Prev PLOTS--------------####
####--------------------------------####
# get the unique pathogens
pathogens <- unique(pcr_prev$pathogen)
pathogens_clean <- str_replace_all(pathogens, " ", "_")
# create a plot for each pathogen
pcr_plot_list <- map(pathogens, function(p){

  pcr_prev %>%
    filter(pathogen == p) %>%
    ggplot(aes(x = community_type2, y = pcr_prevalence*100, color = community_type2,)) +
    geom_point() + 
    labs(
      y = "PCR Prevalence",
      x = ""
      #title = paste("PCR Prevalence:", p)
    ) +
    scale_color_manual(values = nytpal) + 
    theme(
      axis.text.x = element_blank(),
      legend.position = "none", 
      panel.background = element_rect(fill = "white"),
      panel.grid = element_line(color = "grey", linewidth = 0.1), 
       plot.margin = margin(0, 0, 0, 0),
      panel.spacing.y = unit(0, "pt"),
      strip.background = element_blank(),
      strip.text = element_text(margin = margin(0, 0, 0, 0))
    ) +
    geom_errorbar(aes(ymin = lower_ci*100, ymax = upper_ci*100), 
                  width = 0.1, linewidth = 0.5) +
  coord_fixed(ratio = 0.13, ylim = c(0, 30)) + small_margins  # equal height/width scaling
  
})

# name the plots like pcr_[pathogen]
names(pcr_plot_list) <- paste0("pcr_", pathogens_clean)

# optionally push all plots into the global environment
list2env(pcr_plot_list, envir = .GlobalEnv)

```

### 3) SCR Ratio

```{r}
# Get list of unique pathogens
pathogens <- unique(scrr$pathogen)

# Clean pathogen names so object names have no spaces
pathogens_clean <- str_replace_all(pathogens, " ", "_")

# Create a plot for each pathogen
scrr_plot_list <- map(pathogens, function(p){

  scrr %>%  
    filter(pathogen == p) %>% 
    mutate(
      community = case_when(
        term == "(Intercept)" ~ "Esmeraldas", 
        term == "community_typeBorbon" ~ "Borbón",
        term == "community_typeRural"  ~ "Rural",
        TRUE ~ NA_character_
      ),
      # Force x-axis levels (Esmeraldas stays even if dropped from plot)
      community = factor(community, 
                         levels = c("Esmeraldas", "Borbón", "Rural"))
    ) %>%
    
    # REMOVE "Esmeraldas" from data so it does not plot  
    filter(community != "Esmeraldas") %>%
    
    ggplot(aes(x = community, color = community, group = community)) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                    position = position_dodge(width = 0.6),
                    width = 0.1) +
      geom_point(aes(y = estimate),
                 position = position_dodge(width = 0.6),
                 size = 2) +
      
      # FREE Y-AXIS (removed coord_cartesian)
      # ggplot will choose appropriate limits
      
      scale_color_manual(values = nytpal[c(2, 3)]) +
      
      scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    scale_x_discrete(drop = FALSE) +
      
      geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
      
      labs(
        x = "",
        y = "RR",
        color = "Community"
        #title = paste("SCRR:", p),
      ) +
      
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 12),
        legend.position = "none", 
         plot.margin = margin(0, 0, 0, 0),
      panel.spacing.y = unit(0, "pt"),
      strip.background = element_blank(),
      strip.text = element_text(margin = margin(0, 0, 0, 0))) + 
    coord_fixed(ratio = 1.2, ylim = c(0, 3.5)) 
})

# Name each plot object scrr_[pathogen]
names(scrr_plot_list) <- paste0("scrr_", pathogens_clean)

# Save each plot into the global environment
list2env(scrr_plot_list, envir = .GlobalEnv)


# Need to manually make one for E. histolytica due to range being wide 

scrr_E._histolytica <-  scrr %>%  
    filter(pathogen == "E. histolytica") %>% 
    mutate(
      community = case_when(
        term == "(Intercept)" ~ "Esmeraldas", 
        term == "community_typeBorbon" ~ "Borbón",
        term == "community_typeRural"  ~ "Rural",
        TRUE ~ NA_character_
      ),
      # Force x-axis levels (Esmeraldas stays even if dropped from plot)
      community = factor(community, 
                         levels = c("Esmeraldas", "Borbón", "Rural"))
    ) %>%
    
    # REMOVE "Esmeraldas" from data so it does not plot  
    filter(community != "Esmeraldas") %>%
    
    ggplot(aes(x = community, color = community, group = community)) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                    position = position_dodge(width = 0.6),
                    width = 0.1) +
      geom_point(aes(y = estimate),
                 position = position_dodge(width = 0.6),
                 size = 2) +
      
      scale_color_manual(values = nytpal[c(2, 3)]) +
      scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    scale_x_discrete(drop = FALSE) +
      
      geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
      
      labs(
        x = "",
        y = "RR",
        color = "Community"
        #title = paste("SCRR:", p),
      ) +
      
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 12),
        legend.position = "none"
      ) + 
    coord_fixed(ratio = 0.25, ylim = c(0, 18)) + small_margins

```

### 4) PCR Prevalence Ratio

```{r}
# Get list of unique pathogens
pathogens <- unique(prevr$pathogen)

# Clean pathogen names so object names have no spaces
pathogens_clean <- str_replace_all(pathogens, " ", "_")

# Create a plot for each pathogen
pr_plot_list <- map(pathogens, function(p){

  prevr %>%  
    filter(pathogen == p) %>% 
    mutate(
      community = case_when(
        term == "(Intercept)" ~ "Esmeraldas", 
        term == "community_typeBorbon" ~ "Borbón",
        term == "community_typeRural"  ~ "Rural",
        TRUE ~ NA_character_
      ),
      # Force x-axis levels (Esmeraldas stays even if dropped from plot)
      community = factor(community, 
                         levels = c("Esmeraldas", "Borbón", "Rural"))
    ) %>%
    
    # REMOVE "Esmeraldas" from data so it does not plot  
    filter(community != "Esmeraldas") %>%
    
    ggplot(aes(x = community, color = community, group = community)) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                    position = position_dodge(width = 0.6),
                    width = 0.1) +
      geom_point(aes(y = estimate),
                 position = position_dodge(width = 0.6),
                 size = 2) +
      
      # FREE Y-AXIS (removed coord_cartesian)
      # ggplot will choose appropriate limits
      
      scale_color_manual(values = nytpal[c(2, 3)]) +
      
      scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    scale_x_discrete(drop = FALSE) +
      
      geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
      
      labs(
        x = "",
        y = "PR",
        color = "Community"
        #title = paste("PR:", p)
      ) +
      
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 12),
        legend.position = "none", 
         plot.margin = margin(0, 0, 0, 0),
      panel.spacing.y = unit(0, "pt"),
      strip.background = element_blank(),
      strip.text = element_text(margin = margin(0, 0, 0, 0))) + 
    coord_fixed(ratio = 1.2, ylim = c(0, 3.5)) + small_margins
      
})

# Name each plot object scrr_[pathogen]
names(pr_plot_list) <- paste0("prevr_", pathogens_clean)

# Save each plot into the global environment
list2env(pr_plot_list, envir = .GlobalEnv)

`prevr_LT-ETEC`

# Manually do one for E. histolytica #
prevr_E._histolytica <- prevr %>%  
    filter(pathogen == "E. histolytica") %>% 
    mutate(
      community = case_when(
        term == "(Intercept)" ~ "Esmeraldas", 
        term == "community_typeBorbon" ~ "Borbón",
        term == "community_typeRural"  ~ "Rural",
        TRUE ~ NA_character_
      ),
      # Force x-axis levels (Esmeraldas stays even if dropped from plot)
      community = factor(community, 
                         levels = c("Esmeraldas", "Borbón", "Rural"))
    ) %>%
    
    # REMOVE "Esmeraldas" from data so it does not plot  
    filter(community != "Esmeraldas") %>%
    
    ggplot(aes(x = community, color = community, group = community)) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                    position = position_dodge(width = 0.6),
                    width = 0.1) +
      geom_point(aes(y = estimate),
                 position = position_dodge(width = 0.6),
                 size = 2) +
      
      # FREE Y-AXIS (removed coord_cartesian)
      # ggplot will choose appropriate limits
      
      scale_color_manual(values = nytpal[c(2, 3)]) +
      
      scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    scale_x_discrete(drop = FALSE) +
      
      geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
      
      labs(
        x = "",
        y = "PR",
        color = "Community"
        #title = paste("PR:", p)
      ) +
      
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 12),
        legend.position = "none") + 
    coord_fixed(ratio = 1.0, ylim = c(0, 4.5)) + small_margins

```

### 5) Merge the plots

\*Decided to just manually make the composite figure as was running into issues doing it in a loop

```{r}
# Get name of overlapping pathogens 
names <- sero_inc %>% filter(sero_inc$pathogen %in% pcr_prev$pathogen)
unique(names$pathogen)
library(patchwork)

# 1) Camplobacter
camp <- (scr_Campylobacter + pcr_Campylobacter) / (scrr_Campylobacter + prevr_Campylobacter) + plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
#ggsave(here("figures/camp_composite.tiff"), camp, height = 5, width = 4)

# 2) LT-ETEC
ltetec <- (`scr_LT-ETEC` + `pcr_LT-ETEC`) / (`scrr_LT-ETEC` + `prevr_LT-ETEC`) + plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
ltetec
#ggsave(here("figures/ltetec_composite.tiff"), ltetec, height = 5, width = 4)

# 3) Shigella
shig <- (`scr_Shigella/EIEC` + `pcr_Shigella/EIEC`) / (`scrr_Shigella/EIEC` + `prevr_Shigella/EIEC`) +
   plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
#ggsave(here("figures/shigella_composite.tiff"), shig, height = 5, width = 4)

# 4) Norovirus GI
noro1 <- (scr_Norovirus_GI + pcr_Norovirus_GI) / (scrr_Norovirus_GI + prevr_Norovirus_GI) +
   plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
noro1
#ggsave(here("figures/noro1_composite.tiff"), noro1, height = 5, width = 4)

# 5) Cryptosporidium
cryp <- (scr_Cryptosporidium + pcr_Cryptosporidium) / (scrr_Cryptosporidium + prevr_Cryptosporidium) +  plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
cryp
#ggsave(here("figures/cryp_composite.tiff"), cryp, height = 5, width = 4)

# 6) E. histolytica
ehist <- (scr_E._histolytica + pcr_E._histolytica) / (scrr_E._histolytica + prevr_E._histolytica) +  plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
ehist
#ggsave(here("figures/ehist_composite.tiff"), ehist, height = 5, width = 4)

# 7) Giardia
giardia <- (scr_Giardia + pcr_Giardia) / (scrr_Giardia + prevr_Giardia) +  plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
giardia
#ggsave(here("figures/giardia_composite.tiff"), giardia, height = 5, width = 4)

# 8) Norovirus GII
noro2 <- (scr_Norovirus_GII + pcr_Norovirus_GII) / (scrr_Norovirus_GII + prevr_Norovirus_GII) +  plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
noro2

# 9) S. enterica
sal <- (scr_S._enterica + pcr_S._enterica) / (scrr_S._enterica + prevr_S._enterica) +  plot_layout(heights = c(1, 1)) &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
sal


#ggsave(here("figures/noro2_composite.tiff"), noro2, height = 5, width = 4)

```

### 6) Create master figures

```{r}
library(patchwork)
plot_list <- list(ltetec, shig, camp, noro2, cryp, giardia, ehist, noro1)

master <- wrap_plots(plot_list, ncol = 4) +       # 4 columns
  plot_layout(guides = "collect") 
# Save
#ggsave(
#  filename = "../figures/figure1_large_horizontal.tiff",
#  plot = master,
#  width = 15, #23
#  height = 11, #15
#  dpi = 300
#)

master

```

### 7) Master figure for ...

#### A) Bacteria

```{r}
plot_list <- list(ltetec, shig, camp)

master_b <- wrap_plots(plot_list, ncol = 3) +       # 4 columns
  plot_layout(guides = "collect") &
  theme(
    plot.margin = margin(0, 0, 0, 0),
    plot.spacing = unit(0, "pt")
  )
  
# Save
#ggsave(
#  filename = "../figures/figure1a_large_horizontal_bacteria.tiff",
#  plot = master_b,
#  width = 10, #23
#  height = 5, #15
#  dpi = 300
#)

master_b

```

#### B) Protozoa

```{r}
plot_list <- list(cryp, giardia, ehist)

master_p <- wrap_plots(plot_list, ncol = 3) +       # 4 columns
  plot_layout(guides = "collect") &
  theme(
    plot.margin = margin(0, 0, 0, 0)
  )
# Save
#ggsave(
#  filename = "../figures/figure1a_large_horizontal_protozoa.tiff",
#  plot = master_p,
#  width = 10, #23
#  height = 5, #15
#  dpi = 300
#)

master_p
```

#### C) Virus +  (for balance)

```{r}
plot_list <- list(noro2, noro1, sal)

master_v <- wrap_plots(plot_list, ncol = 3) +       
  plot_layout(guides = "collect") 
# Save
#ggsave(
#  filename = "../figures/figure1a_large_horizontal_virus.tiff",
#  plot = master_v,
#  width = 10, #23
#  height = 5, #15
#  dpi = 300
#)

master_v
```

### 8) Vertical master

```{r}
layout <- c(
  area(t = 1, l = 1, b = 1, r = 3),   # master_b: full 3 columns
  area(t = 2, l = 1, b = 2, r = 3),   # master_p: full 3 columns
  area(t = 3, l = 1, b = 3, r = 3)    # master_v: full 3 columns to add in 
)

vert <- master_p / master_b / master_v + 
  plot_layout(design = layout)

vert
ggsave(
  filename = "../figures/fig5_moa_large_vertical.tiff",
  plot = vert,
  width = 230, 
  height = 275, 
  units = "mm"
)
```
