---
title: "EcoMID: PCR prevalence"
subtitle: "PCR prevalence by GPP and TAC assay"
author: "Nikolina Walas (nikolina.walas@ucsf.edu) and Juilanna Colado (juliannacolado@berkeley.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
---

## File Overview

This file summarizes the prevalence of 12 enteric pathogens, as measured by the GPP and TAC assays for the ecomid birth cohort. It was updated 2025/10/21 to include the 15 month timepoint from the GPP assay.

There are 484 unique children in the most recent dataset.
There are 2503 unique samples per pathogen. 

There is data for 3, 6, 9, 12, 15, 18, 24 month samples.

It is important to note there is only TAC-PCR data available for Salmonella. This pathogen is only measured at 6, 12, 18 months.

### Load packages

```{r}
library("here")
source(here("R", "0-config.R"))
```

### Read in files

```{r}
# read in data
data <- read_rds(here("data/ecomid_pcr_analysis_data_2026-06-11.rds"))

# overview of unique childids in most recent file
print("Unique childid's")
unique(data$childid) %>% length()

print("Unique sampleid's")
unique(data$sample_ID) %>% length()

# overview of timepoints in most recent file
print("timepoints")
unique(data$visitf)
print("unique childid's per time point")
data %>% distinct(visitf, childid) %>% group_by(visitf) %>% summarise(n = n())
```

## Prevalence by pathogen

```{r}
prev_overall <- data %>%
  group_by(pathogen) %>%
  summarise(
    n_pos = sum(pcrpos, na.rm = TRUE),
    n_total = sum(!is.na(pcrpos)),
    prevalence = n_pos / n_total
  ) %>%
  arrange(desc(prevalence))

prev_overall
```

## Prevalence per community

```{r}
prev_by_community <- data %>%
  group_by(pathogen, community_name) %>%
  summarise(
    n_pos = sum(pcrpos, na.rm = TRUE),
    n_total = sum(!is.na(pcrpos)),
    prevalence = n_pos / n_total
  ) %>%
  arrange(pathogen, desc(prevalence))
prev_by_community

```

## Robust SE by pathogen

```{r}
robust_prev_overall <- data %>%
  group_by(pathogen) %>%
  group_map(~{
    model <- glm(pcrpos ~ 1, family = binomial(), data = .x)
    se <- sqrt(diag(vcovCL(model, cluster = .x$childid)))[1]
    est <- coef(model)[1]
    
    tibble(
      pathogen = .y$pathogen,
      prevalence = plogis(est),
      lower_ci = plogis(est - 1.96 * se),
      upper_ci = plogis(est + 1.96 * se),
      n = nrow(.x)
    )
  }) %>%
  bind_rows()

final_prev_overall <- prev_overall %>%
  left_join(robust_prev_overall, by = "pathogen") %>%
  select(-prevalence.y) %>%          # remove duplicate
  select(-n) %>%                     # remove duplicate
  rename(prevalence = prevalence.x) %>%
  arrange(pathogen)

final_prev_overall
```

## Robust SE by pathogen & community

```{r}
robust_prev_by_community <- data %>%
  group_by(pathogen, community_name) %>%
  group_map(~{
    model <- glm(pcrpos ~ 1, family = binomial(), data = .x)
    se <- sqrt(diag(vcovCL(model, cluster = .x$childid)))[1]
    est <- coef(model)[1]
    
    tibble(
      pathogen = .y$pathogen,
      community_name = .y$community_name,
      prevalence = plogis(est),
      lower_ci = plogis(est - 1.96 * se),
      upper_ci = plogis(est + 1.96 * se),
      n = nrow(.x)
    )
  }) %>%
  bind_rows()

final_prev_by_community <- prev_by_community %>%
  left_join(robust_prev_by_community, by = c("pathogen", "community_name")) %>%
  select(-prevalence.y) %>%     # remove duplicates
  select(-n) %>%                # remove duplicates
  rename(prevalence = prevalence.x) %>%
  arrange(pathogen, community_name)
  

```

## Format

Stack final_prev_overall & final_prev_by_community

Save RDS/CVS

```{r}
# reformat for final export
final_prev_overall$community_name <- "All communities"

# final prevalence table 
final_prev_overall <- final_prev_overall %>%
  select(pathogen, community_name, n_pos, n_total, prevalence, lower_ci, upper_ci) 

# final prevalence table 
saveRDS(final_prev_overall, file = paste0("../output/ecomid_pcr_prev_by_pathogen.rds"))
write.csv(final_prev_overall, file = paste0("../output/ecomid_pcr_prev_by_pathogen.csv"))

# recode to include community type and pathogen label that is same as serology datasets
combined_df <- rbind(final_prev_overall, final_prev_by_community) %>% rename(pcr_prevalence = prevalence) %>% 
  mutate(community_type = case_when(community_name == "Esmeraldas" ~ "Esmeraldas",
      community_name %in% c("Maldonado", "Colon Eloy", "Timbire", "Selva Alegre") ~ 
        "Rural Road",
      community_name %in% c("Zancudo", "Colon de Onzole", "Santo Domingo", "San Francisco") ~ 
        "Rural River",
      community_name == "Borbon" ~ "Borbon", 
      TRUE ~ "All communities"))

# view output
combined_df %>%   
  kable("html", 
        caption = "Prevalence by pathogen and community", 
        align = "c") %>%      # center-align columns
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = "#D3D3D3") %>% # header row styling
  column_spec(1, bold = TRUE, border_right = TRUE)  

saveRDS(combined_df, file = paste0("../output/ecomid_pcr_prev_by_community.rds"))
write.csv(combined_df, file = paste0("../output/ecomid_pcr_prev_by_community.csv"))

```

## Recalculations of 3-level community (Rural, Esmereldas, and Borbon) scale

### Prevalence per community (3-level)

```{r}
# recode data to 3 level community scale 
data2 <- data %>% mutate(community_type2 = case_when(
      community_name == "Esmeraldas" ~ "Esmeraldas",
      community_name == "Borbon" ~ "Borbon",
      community_name %in% c("Maldonado", "Colon Eloy", "Timbire", "Zancudo", "Selva Alegre", "Zancudo", "Colon de Onzole", "Santo Domingo", "San Francisco") ~ "Rural"))

# prevalence 
prev_by_community2 <- data2 %>%
  group_by(pathogen, community_type2) %>%
  summarise(
    n_pos = sum(pcrpos, na.rm = TRUE),
    n_total = sum(!is.na(pcrpos)),
    prevalence = n_pos / n_total
  ) %>%
  arrange(pathogen, desc(prevalence))

# robust se prevalence
robust_prev_by_community2 <- data2 %>%
  group_by(pathogen, community_type2) %>%
  group_map(~{
    model <- glm(pcrpos ~ 1, family = binomial(), data = .x)
    se <- sqrt(diag(vcovCL(model, cluster = .x$childid)))[1]
    est <- coef(model)[1]
    
    tibble(
      pathogen = .y$pathogen,
      community_type2 = .y$community_type2,
      prevalence = plogis(est),
      lower_ci = plogis(est - 1.96 * se),
      upper_ci = plogis(est + 1.96 * se),
      n = nrow(.x)
    )
  }) %>%
  bind_rows()

final_prev_by_community2 <- prev_by_community2 %>%
  left_join(robust_prev_by_community2, by = c("pathogen", "community_type2")) %>%
  select(-prevalence.y) %>%     # remove duplicates
  select(-n) %>%                # remove duplicates
  rename(pcr_prevalence = prevalence.x) %>%
  arrange(pathogen, community_type2)

saveRDS(final_prev_by_community2, file = paste0("../output/ecomid_pcr_prev_by_community_3_levels.rds"))
write.csv(final_prev_by_community2, file = paste0("../output/ecomid_pcr_prev_by_community_3_levels.csv"))
  
```
