---
title: "Seroincidence"
subtitle: "Seroprevalence and seroconversion of ecomid birth cohort against 11 enteric pathogens"
author: "Nikolina Walas (nikolina.walas@ucsf.edu) and Juilanna Colado (juliannacolado@berkeley.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
---

## File Overview

This file calculates the seroprevalence and seroincidence rate (seroconversion rate) for 11 enteric pathogens per community in the ecomid trial, as analyzed by multiplex serology at USF-Q.

Seroconversion included seroreversions, or children that seroconverted twice, and person-time was accounted for in the interim time between the two seroconversions.

There are 1606 unique sample results per pathogen

There are 370 unique children

Code was originally written by Julianna Colado, and updated to include most recent sample testing results by Nikolina Walas

### Load packages

```{r}
library(here)
source(here("R", "0-config.R"))
library(readxl)
library(foreach)
```

### Read in files

1.  Read in file, modified to most recent updated dataset with community attributes

2.  Read in child-visit-level dataset for DOB + dbs/stool/svy visitdates

```{r}
# most recent datasets are from September 10th, 2025
sys.date <- format(Sys.Date(), "%Y%m%d")

data <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_2026-05-19.csv"))

child_visit_data <- read_csv(here("..", "..", "Data", "ecomid-field", "final",
                                  "ecomid_child_visit_info_2026-06-15.csv")) %>% 
  mutate(community_name = ifelse(community_name == "Colon", "Colon de Onzole", community_name))

# overview of unique childids in most recent file
print("Unique childid's")
unique(data$childid) %>% length()

# overview of timepoints in most recent file
print("timepoints")
unique(data$timepoint)

```

## Seroconvergence Table

```{r}
sero_conv_tbl <- data %>%
  # Filter to enteric pathogens only & remove cholera & schistosoma (undefined threhold cut off)
  filter(group == "Enterics") %>%
  filter(pathogen != "Cholera") %>%
  filter(pathogen != "Schistosoma") %>%
  
  # filter to remove samples/targets that are NA for seropositivity (updated 10/13)
  filter(!is.na(mfi)) %>%
  
  # Step 1: Collapse from antigen level to pathogen level
  group_by(childid, timepoint, pathogen, pathogen_lab) %>%
  summarise(
    seropos = max(seropos, na.rm = TRUE),  # max across antigens
    .groups = "drop"
  ) %>%
  
  
  # Step 2: Identify seroconversion 
    # create atrisk, lead_seropos, lead_timepoint, & agediff attributes
  arrange(childid, pathogen, timepoint) %>%
  group_by(childid, pathogen, pathogen_lab) %>%
  mutate(
    atrisk = ifelse(seropos == 0, 1, 0),
    lead_seropos = lead(seropos),
    seroconversion = ifelse(atrisk == 1 & seropos == 0 & lead_seropos == 1, 1, 0)
    # lead_timepoint = lead(timepoint),
    # agediff = lead_timepoint - timepoint
    # person_time = case_when(
    #  atrisk == 0 ~ 0,
    #  atrisk == 1 & seroconversion == 0 ~ agediff,
    #  atrisk == 1 & seroconversion == 1 ~ agediff / 2
    )  %>%
  ungroup() %>%
  select(childid, pathogen, pathogen_lab, timepoint, seropos, atrisk,
         lead_seropos, seroconversion) %>%
  arrange(childid, pathogen, timepoint)

#------------------------------------
# join child_visit_data with sero_conv_tbl
# identify missing dbs_visitdates
#------------------------------------

cv_attrs <- child_visit_data %>%
  mutate(
    timepoint       = as.numeric(timepoint),
    dob             = as_date(dob),
    visitdate_dbs   = as_date(visitdate_dbs),
    visitdate_stool = as_date(visitdate_stool),
    visitdate_svy   = as_date(visitdate_svy),
    targetdate      = as_date(targetdate)
  ) %>%
  select(childid, timepoint, dob, visitdate_dbs, visitdate_stool, visitdate_svy, 
         agedays_dbs, agedays_stool, agedays_svy, targetdate, community_name) %>%
  distinct(childid, timepoint, .keep_all = TRUE) 

joined <- sero_conv_tbl %>%
  mutate(timepoint = as.numeric(timepoint)) %>%
  left_join(cv_attrs, by = c("childid", "timepoint"))

#------------------------------------
# update sero_conv_tbl adding: lead_agedays, agedays_diff, person_time
#------------------------------------
sero_conv_tbl2 <- joined %>%
  mutate(
    # ensure Date types
    dob            = as_date(dob),
    visitdate_dbs  = as_date(visitdate_dbs),
    visitdate_stool= as_date(visitdate_stool),
    visitdate_svy  = as_date(visitdate_svy),
    targetdate     = as_date(targetdate),
    
    #recomputing agedays_* safety check
    agedays_dbs   = coalesce(agedays_dbs,
                             if_else(!is.na(visitdate_dbs)  & !is.na(dob),
                                     as.integer(visitdate_dbs  - dob), NA_integer_)),
    agedays_stool = coalesce(agedays_stool,
                             if_else(!is.na(visitdate_stool)& !is.na(dob),
                                     as.integer(visitdate_stool- dob), NA_integer_)),
    agedays_svy   = coalesce(agedays_svy,
                             if_else(!is.na(visitdate_svy)  & !is.na(dob),
                                     as.integer(visitdate_svy  - dob), NA_integer_)),
    
    # fallback from targetdate if no observed visit date
    agedays_from_target = if_else(!is.na(targetdate) & !is.na(dob),
                                  as.integer(targetdate - dob), NA_integer_),
    
    # best-available age in days for this visit (dbs > stool > svy > target)
    agedays_best = coalesce(agedays_dbs, agedays_stool, agedays_svy, agedays_from_target)
  ) %>%
  arrange(childid, pathogen, agedays_best, timepoint) %>%   # order chronologically
  group_by(childid, pathogen, pathogen_lab) %>%
  mutate(
    lead_agedays_best = lead(agedays_best),
    age_diff_days     = lead_agedays_best - agedays_best,
    person_time_days = case_when(
      atrisk == 0 ~ 0,
      atrisk == 1 & seroconversion == 0 ~ age_diff_days,
      atrisk == 1 & seroconversion == 1 ~ age_diff_days / 2
    ),
    person_time_years  = person_time_days / 365.25
  ) %>%
  ungroup() %>%
  select(
    childid, dob, 
    visitdate_dbs, visitdate_stool, visitdate_svy, 
    pathogen, pathogen_lab, timepoint,targetdate, 
    seropos, atrisk, lead_seropos, seroconversion, 
    agedays_dbs, agedays_stool, agedays_svy, agedays_from_target,
    agedays_best, lead_agedays_best, age_diff_days,
    person_time_days, person_time_years, community_name
  ) %>%
  arrange(childid, pathogen, timepoint) %>% 
  mutate(community_type = case_when(community_name == "Esmeraldas" ~ "Esmeraldas", 
                                    community_name == "Borbon" ~ "Borbon", 
                                    TRUE ~ "Rural"))
sero_conv_tbl2$community_type <- factor(sero_conv_tbl2$community_type, levels = c("Esmeraldas", "Borbon", "Rural"))
```

## Prevalence per pathogen

```{r}
prev_overall <- sero_conv_tbl2 %>%
  ungroup() %>% 
  group_by(pathogen) %>%
  summarise(
    n_pos = sum(seropos == 1, na.rm = TRUE),
    n_total = sum(!is.na(seropos)), 
    prevalence = n_pos / n_total
  ) %>%
  arrange(desc(prevalence))

prev_overall 
```

## Prevalence per pathogen & community

```{r}
prev_by_community <- data %>%
  filter(group == "Enterics") %>%
  filter(pathogen != "Cholera") %>%
  filter(pathogen != "Schistosoma") %>%
  filter(!is.na(log10mfi)) %>% 
   group_by(childid, timepoint, pathogen, pathogen_lab, community_name, community_type) %>%
  summarise(
    seropos = max(seropos, na.rm = TRUE),  # max across antigens
    .groups = "drop"
  ) %>% 
  group_by(pathogen, community_name, community_type) %>%
  summarise(
    n_pos = sum(seropos, na.rm = TRUE),
    n_total = sum(!is.na(seropos)),
    prevalence = n_pos / n_total
  ) %>%
  arrange(pathogen, desc(prevalence))

prev_by_community 
```

## Prevalence robust SE: overall pathogen

```{r}
robust_prev_overall <- sero_conv_tbl2 %>%
  group_by(pathogen) %>%
  group_map(~{
    model <- glm(seropos ~ 1, family = binomial(), data = .x) # intercept only to get log-odds of seroprevalence for that pathogen 
    se <- sqrt(diag(vcovCL(model, cluster = .x$childid)))[1] # accounts for intra-cluster correlation among children with the same childid — so observations within a child aren’t treated as independent. (get cluster robust standard error)
    est <- coef(model)[1] # estimated intercept
    
    tibble(
      pathogen = .y$pathogen,
      prevalence = plogis(est),
      lower_ci = plogis(est - 1.96 * se), # Wald-type confidence interval around the logit estimate and back transform to probability 
      upper_ci = plogis(est + 1.96 * se), # Wald-type confidence interval around the logit estimat and back transform to probability
      n = nrow(.x)
    )
  }) %>%
  bind_rows()

final_prev_overall <- prev_overall %>%
  left_join(robust_prev_overall, by = "pathogen") %>%
  select(-prevalence.y) %>%          # remove duplicate
  select(-n) %>%                     # remove duplicate
  rename(prevalence = prevalence.x) %>%
  arrange(pathogen)
```

## Prevalence robust SE: by community & pathogen

```{r}
robust_prev_by_community <- data %>%
  filter(group == "Enterics", !pathogen %in% c("Cholera","Schistosoma")) %>%
  filter(!is.na(log10mfi)) %>% 
   group_by(childid, timepoint, pathogen, pathogen_lab, community_name, community_type) %>%
  summarise(
    seropos = max(seropos, na.rm = TRUE),  # max across antigens
    .groups = "drop"
  ) %>% 
  group_by(pathogen, community_name) %>%
  summarise(
    n   = sum(!is.na(seropos)),
    pos = sum(seropos == 1, na.rm = TRUE),
    prevalence = pos / n,
    # exact (Clopper–Pearson) CI works for 0 or n successes
    lower_ci = binom.test(pos, n)$conf.int[1],
    upper_ci = binom.test(pos, n)$conf.int[2],
    .groups = "drop"
  )

final_prev_by_community <- prev_by_community %>%
  left_join(robust_prev_by_community, by = c("pathogen", "community_name")) %>%
  select(-prevalence.y) %>%     # remove duplicates
  select(-n, -pos) %>%                # remove duplicates
  rename(prevalence = prevalence.x) %>%
  arrange(pathogen, community_name)
```

## Bootstrap: seroconversion per pathogen

```{r}
# seroconv incidence per pathogen
# Build child × pathogen summary
#------------------------------------
child_pathogen <- sero_conv_tbl2 %>%
  group_by(pathogen, childid) %>%
  summarise(
    events = sum(seroconversion, na.rm = TRUE),   # cap at 1 if you want first-ever
    py     = sum(person_time_years, na.rm = TRUE),          # person-time in YEARS
    .groups = "drop"
  )                                   # only contributors

#------------------------------------
# helper function: calc seroreversion rate 
#------------------------------------
est_ir <- function(seroconv, pt) {
  nsc  <- sum(seroconv, na.rm = TRUE)
  ptar <- sum(pt,        na.rm = TRUE)
  if (ptar <= 0) return(NA_real_)
  nsc / ptar
  }

#------------------------------------
# set up bootstrap resample childid and calc IR overall pathogens
#------------------------------------
set.seed(235235)
nboot <- 1000

pathogens <- sort(unique(child_pathogen$pathogen))
boot_results <- vector("list", length(pathogens))

for (k in seq_along(pathogens)) {
  p  <- pathogens[k]
  df <- child_pathogen %>% filter(pathogen == p)
  
  # 1) Observed IR for this pathogen 
  ir_obs <- est_ir(df$events, df$py)
  
  # 2) resampling matrix of child IDs (clusters) for pathogen k
  idlist    <- unique(df$childid)
  n_ids     <- length(idlist)
  boot_samp <- matrix(
    sample(idlist, size = n_ids * nboot, replace = TRUE),
    nrow = n_ids, ncol = nboot
    )
  # 3) Bootstrap loop: rebuild dataset per replicate via left_join 
  ir_boots <- foreach(i = 1:nboot, .combine = c) %do% {
    di <- left_join(
      tibble(childid = boot_samp[, i]),
      df, by = "childid",                 # duplicates keep whole child trajectory
      relationship = "many-to-many"
      )
    iri <- est_ir(di$events, di$py)
    return(iri)
  }
  
  boot_results[[k]] <- tibble(
    pathogen = p,
    ir_obs = ir_obs,
    ir_lo  = quantile(ir_boots, 0.025, na.rm = TRUE),
    ir_hi  = quantile(ir_boots, 0.975, na.rm = TRUE),
    ir_obs_per_100 = 100 * ir_obs,
    ir_lo_per_100  = 100 * ir_lo,
    ir_hi_per_100  = 100 * ir_hi
  )
}

#examine distribution
hist(ir_boots, breaks = 40)

boot_summary <- bind_rows(boot_results)

```

## Bootstrap: seroconversion per pathogen by community

```{r}
child_comm <- data %>%
  select(childid, community_name, community_type) %>% 
  rename(community_type2 = community_type) %>% 
  distinct()

child_pathogen_comm <- sero_conv_tbl2 %>%
  left_join(child_comm %>% select(-community_name), by = "childid") %>%
  group_by(pathogen, community_name, community_type, childid) %>%
  summarise(
    # events per child (cap at 1 if you only want first-ever seroconversion)
    events = sum(seroconversion, na.rm = TRUE),
    # person-time in YEARS
    py     = sum(person_time_years, na.rm = TRUE),
    .groups = "drop_last"
  )
   

# 3) Bootstrap (child-cluster) within each pathogen × community
set.seed(19475)
nboot <- 1000
keys <- child_pathogen_comm %>%
  distinct(pathogen, community_name, community_type) %>%
  arrange(pathogen, community_name, community_type)
boot_results_pc <- vector("list", nrow(keys))



for (i in seq_len(nrow(keys))) {
  p  <- keys$pathogen[i]
  cm <- keys$community_name[i]
  df <- child_pathogen_comm %>% filter(pathogen == p, community_name == cm)
  
  n_children <- n_distinct(df$childid)
  total_events <- sum(df$events, na.rm = TRUE)
  total_py     <- sum(df$py,     na.rm = TRUE)
  
  # Observed IR for this stratum
  ir_obs <- est_ir(df$events, df$py)
  
  # skip stratum if not enough available info
  if (n_children == 0 || is.na(ir_obs) || total_py <= 0) {
    boot_results_pc[[i]] <- tibble(
      pathogen = p, community_name = cm,
      n_children, total_events, total_py,
      ir = ir_obs, ir_per_100 = 100 * ir_obs,
      ir_lo = NA_real_, ir_hi = NA_real_,
      ir_lo_per_100 = NA_real_, ir_hi_per_100 = NA_real_
    )
    next  
  }
  
  # Matrix of resampled child IDs (clusters)
  idlist <- unique(df$childid)
  n_ids  <- length(idlist)
  boot_samp <- matrix(
    sample(idlist, size = n_ids * nboot, replace = TRUE),
    nrow = n_ids, ncol = nboot
  )
  
  ir_boots <- foreach(b = 1:nboot, .combine = c) %do% {
    di <- left_join(
      tibble(childid = boot_samp[, b]),
      df, by = "childid" 
    )
    est_ir(di$events, di$py)
  }
  
  # Precompute quantiles, then per-100
  lo <- as.numeric(quantile(ir_boots, 0.025, na.rm = TRUE))
  hi <- as.numeric(quantile(ir_boots, 0.975, na.rm = TRUE))


  boot_results_pc[[i]] <- tibble(
    pathogen = p, community_name = cm,
    n_children, total_events, total_py,
    ir = ir_obs, ir_per_100 = 100 * ir_obs,
    ir_lo = lo, ir_hi = hi,
    ir_lo_per_100 = 100 * lo,
    ir_hi_per_100 = 100 * hi
  )
}


#examine distribution
hist(ir_boots, breaks = 30)

boot_summary_pc <- bind_rows(boot_results_pc) %>%
  arrange(pathogen, community_name) %>% 
  left_join(data %>% # manually add back in community_type labels
              select(pathogen, community_name, 
                     community_type), 
            by = c("pathogen", "community_name")) %>% unique()

```

## Bootstrap: seroconversion per pathogen by age group

```{r}
child_agegroup <- data %>%
  select(childid, community_name, timepoint) %>%
  distinct() %>% 
  mutate(age_group = ifelse(timepoint < 12, "6-9 months", "12-24 months"))

child_pathogen_age <- sero_conv_tbl2 %>%
  left_join(child_agegroup, by = c("childid","timepoint")) %>%
  group_by(pathogen, age_group, childid) %>%
  summarise(
    # events per child (cap at 1 if you only want first-ever seroreversion)
    events = sum(seroconversion, na.rm = TRUE),
    # person-time in YEARS
    py     = sum(person_time_years, na.rm = TRUE),
    .groups = "drop_last"
  )
   
# 3) Bootstrap (child-cluster) within each pathogen × community
set.seed(960712)
nboot <- 1000
keys <- child_pathogen_age %>%
  distinct(pathogen, age_group) %>%
  arrange(pathogen, age_group)
boot_results_ag <- vector("list", nrow(keys))



for (i in seq_len(nrow(keys))) {
  p  <- keys$pathogen[i]
  ag <- keys$age_group[i]
  df <- child_pathogen_age %>% filter(pathogen == p, age_group == ag)
  
  n_children <- n_distinct(df$childid)
  total_events <- sum(df$events, na.rm = TRUE)
  total_py     <- sum(df$py,     na.rm = TRUE)
  
  # Observed IR for this stratum
  ir_obs <- est_ir(df$events, df$py)
  
  # skip stratum if not enough available info
  if (n_children == 0 || is.na(ir_obs) || total_py <= 0) {
    boot_results_ag[[i]] <- tibble(
      pathogen = p, age_group = ag,
      n_children, total_events, total_py,
      ir = ir_obs, ir_per_100 = 100 * ir_obs,
      ir_lo = NA_real_, ir_hi = NA_real_,
      ir_lo_per_100 = NA_real_, ir_hi_per_100 = NA_real_
    )
    next  
  }
  
  # Matrix of resampled child IDs (clusters)
  idlist <- unique(df$childid)
  n_ids  <- length(idlist)
  boot_samp <- matrix(
    sample(idlist, size = n_ids * nboot, replace = TRUE),
    nrow = n_ids, ncol = nboot
  )
  
  ir_boots <- foreach(b = 1:nboot, .combine = c) %do% {
    di <- left_join(
      tibble(childid = boot_samp[, b]),
      df, by = "childid" 
    )
    est_ir(di$events, di$py)
  }
  
  # Precompute quantiles, then per-100
  lo <- as.numeric(quantile(ir_boots, 0.025, na.rm = TRUE))
  hi <- as.numeric(quantile(ir_boots, 0.975, na.rm = TRUE))


  boot_results_ag[[i]] <- tibble(
    pathogen = p, age_group = ag,
    n_children, total_events, total_py,
    ir = ir_obs, ir_per_100 = 100 * ir_obs,
    ir_lo = lo, ir_hi = hi,
    ir_lo_per_100 = 100 * lo,
    ir_hi_per_100 = 100 * hi
  )
}


#examine distribution
hist(ir_boots, breaks = 30)

boot_summary_ag <- bind_rows(boot_results_ag) %>%
  arrange(pathogen, age_group) %>%
  distinct()

# export
write.csv(boot_summary_ag, file = here("output/ecomid_pathogen_agegroup_sero_inc.csv"))
saveRDS(boot_summary_ag, file = here("output/ecomid_pathogen_agegroup_sero_inc.rds"))
```

## Seroconversion rate calculation

Daily and annually

```{r}
sero_inc_rates <- sero_conv_tbl2 %>%
  group_by(pathogen) %>%
  summarise(
    total_seroconversions = sum(seroconversion, na.rm = TRUE),
    total_person_time_days = sum(person_time_days, na.rm = TRUE),  # includes person-time for those not at risk? 
    total_person_time_years = sum(person_time_years, na.rm = TRUE),
    
    incidence_rate_per_day = total_seroconversions / total_person_time_days,
    incidence_rate_per_year = total_seroconversions / total_person_time_years,
    
    inc_per_100_child_years = 100 * incidence_rate_per_year
    
  ) %>%
  ungroup()

# check how many seroreversions there are 
result <- sero_conv_tbl2 %>% 
  arrange(childid, pathogen, timepoint) %>%
  group_by(childid, pathogen) %>%
  summarise(
    pattern = paste(seroconversion, collapse = "")
  ) %>%
  filter(grepl("10+1", pattern)) %>%   # 1, then one or more 0s, then 1
  #select(childid, pathogen) %>%
  unique() %>% as.data.frame() %>% rename(childid = 1)

print("Number of seroreversions (childid-pathogen level)")
length(unique(result$childid))

print("Number of unique children with a seroreversion for at least 1 pathogen")
result %>% select(childid) %>% unique() %>% nrow()

print("Top pathogens for multiple seroconversions")
result %>% select(pathogen) %>% 
  group_by(pathogen) %>% 
  summarise(n_children = n()) %>% 
  arrange(desc(n_children))

sero_inc_by_path <- sero_inc_rates %>%
  left_join(boot_summary, by = "pathogen") %>%
  select(pathogen, total_seroconversions, total_person_time_years, ir_obs_per_year = ir_obs, ir_lo_per_year = ir_lo, ir_hi_per_year = ir_hi, ir_obs_per_100, ir_lo_per_100, ir_hi_per_100)
```

## Seroconversion table

```{r}
#------------------------------------
# export seroconversion table, 
# prevalence & 95% CI overall and by community,
# seroincidence rates CI csv & rds
#------------------------------------
final_prev_overall$community_name <- "All communities"
final_prev_overall$community_type <- "All communities"
final_prev_overall <- final_prev_overall %>%
  select(pathogen, community_name, community_type, n_pos, n_total, prevalence, lower_ci, upper_ci)

combined_prev_df <- rbind(final_prev_overall, final_prev_by_community) 


# rename prevalence to seroprevalence so as not to get confused during later analysis and merges 

saveRDS(combined_prev_df %>% rename(sero_prevalence = prevalence), file = here("output/ecomid_sero_prev_rates.rds"))
write.csv(combined_prev_df %>% rename(sero_prevalence = prevalence), file = here("output/ecomid_sero_prev_rates.csv"))


saveRDS(sero_conv_tbl2, file = here("output/ecomid_sero_conv_tbl.rds"))
write.csv(sero_conv_tbl2, file = here("output/ecomid_sero_conv_tbl.csv"))


write.csv(sero_inc_by_path, file = here("output/ecomid_pathogen_sero_inc.csv"))
saveRDS(sero_inc_by_path, file = here("output/ecomid_pathogen_sero_inc.rds"))


write.csv(boot_summary_pc, file = here("output/ecomid_pathogen_comm_sero_inc.csv"))
saveRDS(boot_summary_pc, file = here("output/ecomid_pathogen_comm_sero_inc.rds"))

```

## Aggregated to 3-community levels

For ease of analysis in downstream, repeat the seroconvergence, prevalence and bootstrapping for different grouping levels.

### Prevalence per pathogen & community level 2

```{r}
# create higher level community variable 
data2 <- data %>% mutate(community_type2 = case_when(
      community_name == "Esmeraldas" ~ "Esmeraldas",
      community_name == "Borbon" ~ "Borbon",
      community_name %in% c("Maldonado", "Colon Eloy", "Timbire", "Zancudo", "Selva Alegre", "Zancudo", "Colon de Onzole", "Santo Domingo", "San Francisco") ~ "Rural"))

prev_by_community2 <- data2 %>%
  filter(group == "Enterics") %>%
  filter(pathogen != "Cholera") %>%
  filter(pathogen != "Schistosoma") %>%
  filter(!is.na(mfi)) %>% 
  group_by(childid, timepoint, community_type2, pathogen, pathogen_lab) %>%
  summarise(
    seropos = max(seropos, na.rm = TRUE),  # max across antigens
    .groups = "drop"
  ) %>%
  group_by(pathogen, community_type2) %>%
  summarise(
    n_pos = sum(seropos, na.rm = TRUE),
    n_total = sum(!is.na(seropos)),
    prevalence = n_pos / n_total, 
    # exact (Clopper–Pearson) CI works for 0 or n successes
    lower_ci = binom.test(n_pos, n_total)$conf.int[1],
    upper_ci = binom.test(n_pos, n_total)$conf.int[2],
    .groups = "drop"
  ) %>%
  arrange(pathogen, desc(prevalence))
prev_by_community2
```

### Export prevalence per pathogen & community level (3 levels)

```{r}
# export 
saveRDS(prev_by_community2 %>% rename(sero_prevalence = prevalence), file = here("output/ecomid_sero_prev_rates_3_levels.rds"))
write.csv(prev_by_community2 %>% rename(sero_prevalence = prevalence), file = here("output/ecomid_sero_prev_rates_3_levels.csv"))
```

### Bootstrap sero-conversion per pathogen & community levels

Repeat steps as before but with new aggregated level

```{r}
child_comm2 <- data2 %>%
  select(childid, community_type2) %>%
  distinct()

child_pathogen_comm2 <- sero_conv_tbl2 %>%
  left_join(child_comm2, by = "childid") %>%
  group_by(pathogen, community_type2, childid) %>%
  summarise(
    # events per child (cap at 1 if you only want first-ever seroconversion)
    events = sum(seroconversion, na.rm = TRUE),
    # person-time in YEARS
    py     = sum(person_time_years, na.rm = TRUE),
    .groups = "drop_last"
  )
   

# 3) Bootstrap (child-cluster) within each pathogen × community
set.seed(19373)
nboot <- 1000
keys <- child_pathogen_comm2 %>%
  distinct(pathogen, community_type2) %>%
  arrange(pathogen, community_type2)
boot_results_pc <- vector("list", nrow(keys))



for (i in seq_len(nrow(keys))) {
  p  <- keys$pathogen[i]
  cm <- keys$community_type2[i]
  df <- child_pathogen_comm2 %>% filter(pathogen == p, community_type2 == cm)
  
  n_children <- n_distinct(df$childid)
  total_events <- sum(df$events, na.rm = TRUE)
  total_py     <- sum(df$py,     na.rm = TRUE)
  
  # Observed IR for this stratum
  ir_obs <- est_ir(df$events, df$py)
  
  # skip stratum if not enough available info
  if (n_children == 0 || is.na(ir_obs) || total_py <= 0) {
    boot_results_pc[[i]] <- tibble(
      pathogen = p, community_type2 = cm,
      n_children, total_events, total_py,
      ir = ir_obs, ir_per_100 = 100 * ir_obs,
      ir_lo = NA_real_, ir_hi = NA_real_,
      ir_lo_per_100 = NA_real_, ir_hi_per_100 = NA_real_
    )
    next  
  }
  
  # Matrix of resampled child IDs (clusters)
  idlist <- unique(df$childid)
  n_ids  <- length(idlist)
  boot_samp <- matrix(
    sample(idlist, size = n_ids * nboot, replace = TRUE),
    nrow = n_ids, ncol = nboot
  )
  
  ir_boots <- foreach(b = 1:nboot, .combine = c) %do% {
    di <- left_join(
      tibble(childid = boot_samp[, b]),
      df, by = "childid" 
    )
    est_ir(di$events, di$py)
  }
  
  # Precompute quantiles, then per-100
  lo <- as.numeric(quantile(ir_boots, 0.025, na.rm = TRUE))
  hi <- as.numeric(quantile(ir_boots, 0.975, na.rm = TRUE))


  boot_results_pc[[i]] <- tibble(
    pathogen = p, community_type2 = cm,
    n_children, total_events, total_py,
    ir = ir_obs, ir_per_100 = 100 * ir_obs,
    ir_lo = lo, ir_hi = hi,
    ir_lo_per_100 = 100 * lo,
    ir_hi_per_100 = 100 * hi
  )
}


# examine distribution
hist(ir_boots, breaks = 30)

boot_summary_pc <- bind_rows(boot_results_pc) %>%
  arrange(pathogen, community_type2)

# export 

write.csv(boot_summary_pc, file = here("output/ecomid_pathogen_comm_sero_inc_3_levels.csv"))
saveRDS(boot_summary_pc, file = here("output/ecomid_pathogen_comm_sero_inc_3_levels.rds"))
```
