---
title: "sero_reversions_ecomid"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
---

## File overview

This file calculates the seroreversion rate for all pathogens included in the pcr-serology enteric comparison study 

### Load packages

```{r}
library(here)
source(here("R", "0-config.R"))
library(readxl)
library(foreach)
```

### Read in files

1.  Read in file, modified to most recent updated dataset with community attributes

2.  Read in child-visit-level dataset for DOB + dbs/stool/svy visitdates

```{r}
data <- read_csv(here("..","..",
                      "Data","serology","final","ecomid_serology_datasets",
                      "ecomid_sero_antigen_seropos_long_2026-08-26.csv")) %>% 
  filter(antigen_lab != "S. enterica HlyE") # remove way upstream please

child_visit_data <- read_csv(here("..", "..", "Data", "ecomid-field", "final",
                                  "ecomid_child_visit_info_2026-06-15.csv")) %>% 
  mutate(community_name = ifelse(community_name == "Colon", "Colon de Onzole", community_name))

# overview of unique childids in most recent file
print("Unique childid's")
unique(data$childid) %>% length()

# overview of timepoints in most recent file
print("timepoints")
unique(data$timepoint)

```

## Seroreversion Table

```{r}
sero_rev_tbl <- data %>%
  # Filter to enteric pathogens only & remove cholera & schistosoma (undefined threhold cut off; also remove HylE that is specific for paratyphi)
  filter(group == "Enterics") %>%
  filter(pathogen != "Cholera") %>%
  filter(pathogen != "Schistosoma") %>%
  filter(antigen_lab != "Salmonella HylE") %>%
  
  # filter to remove samples/targets that are NA for seropositivity (updated 10/13)
  filter(!is.na(mfi))

# safety check: what antigens are being included? 
  unique(sero_rev_tbl$antigen_lab)

  
  # Step 1: Collapse from antigen level to pathogen level
sero_rev_tbl <- sero_rev_tbl %>%  group_by(childid, timepoint, pathogen, pathogen_lab) %>%
  summarise(
    seropos = max(seropos, na.rm = TRUE),  # max across antigens
    .groups = "drop"
  ) %>%
  
  
  # Step 2: Identify seroconversion 
    # create atrisk, lead_seropos, lead_timepoint, & agediff attributes
  arrange(childid, pathogen, timepoint) %>%
  group_by(childid, pathogen, pathogen_lab) %>%
  mutate(
    atrisk = ifelse(seropos == 1, 1, 0), # only at risk if was seropositive and now losing it
    lead_seropos = lead(seropos),
    seroreversion = ifelse(atrisk == 1 & seropos == 1 & lead_seropos == 0, 1, 0)
    )  %>%
  ungroup() %>%
  select(childid, pathogen, pathogen_lab, timepoint, seropos, atrisk,
         lead_seropos, seroreversion) %>%
  arrange(childid, pathogen, timepoint)

#------------------------------------
# join child_visit_data with sero_conv_tbl
# identify missing dbs_visitdates
#------------------------------------

cv_attrs <- child_visit_data %>%
  mutate(
    timepoint       = as.numeric(timepoint),
    dob             = as_date(dob),
    visitdate_dbs   = as_date(visitdate_dbs),
    visitdate_stool = as_date(visitdate_stool),
    visitdate_svy   = as_date(visitdate_svy),
    targetdate      = as_date(targetdate)
  ) %>%
  select(childid, timepoint, dob, visitdate_dbs, visitdate_stool, visitdate_svy, 
         agedays_dbs, agedays_stool, agedays_svy, targetdate, community_name) %>%
  distinct(childid, timepoint, .keep_all = TRUE) 

joined <- sero_rev_tbl %>%
  mutate(timepoint = as.numeric(timepoint)) %>%
  left_join(cv_attrs, by = c("childid", "timepoint"))

#------------------------------------
# update sero_conv_tbl adding: lead_agedays, agedays_diff, person_time
#------------------------------------
sero_rev_tbl2 <- joined %>%
  mutate(
    # ensure Date types
    dob            = as_date(dob),
    visitdate_dbs  = as_date(visitdate_dbs),
    visitdate_stool= as_date(visitdate_stool),
    visitdate_svy  = as_date(visitdate_svy),
    targetdate     = as_date(targetdate),
    
    #recomputing agedays_* safety check
    agedays_dbs   = coalesce(agedays_dbs,
                             if_else(!is.na(visitdate_dbs)  & !is.na(dob),
                                     as.integer(visitdate_dbs  - dob), NA_integer_)),
    agedays_stool = coalesce(agedays_stool,
                             if_else(!is.na(visitdate_stool)& !is.na(dob),
                                     as.integer(visitdate_stool- dob), NA_integer_)),
    agedays_svy   = coalesce(agedays_svy,
                             if_else(!is.na(visitdate_svy)  & !is.na(dob),
                                     as.integer(visitdate_svy  - dob), NA_integer_)),
    
    # fallback from targetdate if no observed visit date
    agedays_from_target = if_else(!is.na(targetdate) & !is.na(dob),
                                  as.integer(targetdate - dob), NA_integer_),
    
    # best-available age in days for this visit (dbs > stool > svy > target)
    agedays_best = coalesce(agedays_dbs, agedays_stool, agedays_svy, agedays_from_target)
  ) %>%
  arrange(childid, pathogen, agedays_best, timepoint) %>%   # order chronologically
  group_by(childid, pathogen, pathogen_lab) %>%
  mutate(
    lead_agedays_best = lead(agedays_best),
    age_diff_days     = lead_agedays_best - agedays_best,
    person_time_days = case_when(
      atrisk == 0 ~ 0,
      atrisk == 1 & seroreversion == 0 ~ age_diff_days,
      atrisk == 1 & seroreversion == 1 ~ age_diff_days / 2
    ),
    person_time_years  = person_time_days / 365.25
  ) %>%
  ungroup() %>%
  select(
    childid, dob, 
    visitdate_dbs, visitdate_stool, visitdate_svy, 
    pathogen, pathogen_lab, timepoint,targetdate, 
    seropos, atrisk, lead_seropos, seroreversion, 
    agedays_dbs, agedays_stool, agedays_svy, agedays_from_target,
    agedays_best, lead_agedays_best, age_diff_days,
    person_time_days, person_time_years, community_name
  ) %>%
  arrange(childid, pathogen, timepoint) %>% 
  mutate(community_type = case_when(community_name == "Esmeraldas" ~ "Esmeraldas", 
                                    community_name == "Borbon" ~ "Borbon", 
                                    TRUE ~ "Rural"))

sero_rev_tbl2$community_type <- factor(sero_rev_tbl2$community_type, levels = c("Esmeraldas", "Borbon", "Rural"))
```

## Distribution of seroreversion events per pathogen, by child

```{r}
sero_count <- sero_rev_tbl2 %>%
  group_by(pathogen, childid) %>%
  summarise(
    n_rev = sum(seroreversion, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  count(pathogen, n_rev) %>% rename(n_children = n) %>% 
  mutate(n_rev = paste0(n_rev, "_reversions")) %>% 
  pivot_wider(
    names_from = n_rev,
    values_from = n_children,
    values_fill = 0
  )
```

## Bootstrap: seroreversion per pathogen

```{r}
# seroreversion per pathogen
# Build child × pathogen summary
#------------------------------------
child_pathogen <- sero_rev_tbl2 %>%
  group_by(pathogen, childid) %>%
  summarise(
    events = sum(seroreversion, na.rm = TRUE),   # cap at 1 if you want first-ever
    py     = sum(person_time_years, na.rm = TRUE),          # person-time in YEARS
    .groups = "drop"
  )                                   # only contributors

#------------------------------------
# helper function: calc incidence rate
#------------------------------------
est_ir <- function(serorev, pt) {
  nsc  <- sum(serorev, na.rm = TRUE)
  ptar <- sum(pt,        na.rm = TRUE)
  if (ptar <= 0) return(NA_real_)
  nsc / ptar
  }

#------------------------------------
# set up bootstrap resample childid and calc IR overall pathogens
#------------------------------------
set.seed(19373)
nboot <- 1000

pathogens <- sort(unique(child_pathogen$pathogen))
boot_results <- vector("list", length(pathogens))

for (k in seq_along(pathogens)) {
  p  <- pathogens[k]
  df <- child_pathogen %>% filter(pathogen == p)
  
  # 1) Observed SRR for this pathogen 
  ir_obs <- est_ir(df$events, df$py)
  
  # 2) resampling matrix of child IDs (clusters) for pathogen k
  idlist    <- unique(df$childid)
  n_ids     <- length(idlist)
  boot_samp <- matrix(
    sample(idlist, size = n_ids * nboot, replace = TRUE),
    nrow = n_ids, ncol = nboot
    )
  # 3) Bootstrap loop: rebuild dataset per replicate via left_join 
  ir_boots <- foreach(i = 1:nboot, .combine = c) %do% {
    di <- left_join(
      tibble(childid = boot_samp[, i]),
      df, by = "childid",                 # duplicates keep whole child trajectory
      relationship = "many-to-many"
      )
    iri <- est_ir(di$events, di$py)
    return(iri)
  }
  
  boot_results[[k]] <- tibble(
    pathogen = p,
    ir_obs = ir_obs,
    ir_lo  = quantile(ir_boots, 0.025, na.rm = TRUE),
    ir_hi  = quantile(ir_boots, 0.975, na.rm = TRUE),
    ir_obs_per_100 = 100 * ir_obs,
    ir_lo_per_100  = 100 * ir_lo,
    ir_hi_per_100  = 100 * ir_hi
  )
}

#examine distribution
hist(ir_boots, breaks = 40)

boot_summary <- bind_rows(boot_results)

```
## Bootstrap: seroconversion per pathogen by community

```{r}
child_comm <- data %>%
  select(childid, community_name, community_type) %>% 
  rename(community_type2 = community_type) %>% 
  distinct()

child_pathogen_comm <- sero_rev_tbl2 %>%
  left_join(child_comm %>% select(-community_name), by = "childid") %>%
  group_by(pathogen, community_name, community_type, childid) %>%
  summarise(
    # events per child (cap at 1 if you only want first-ever seroconversion)
    events = sum(seroreversion, na.rm = TRUE),
    # person-time in YEARS
    py     = sum(person_time_years, na.rm = TRUE),
    .groups = "drop_last"
  )
   

# 3) Bootstrap (child-cluster) within each pathogen × community
set.seed(39485)
nboot <- 1000
keys <- child_pathogen_comm %>%
  distinct(pathogen, community_name, community_type) %>%
  arrange(pathogen, community_name, community_type)
boot_results_pc <- vector("list", nrow(keys))


for (i in seq_len(nrow(keys))) {
  p  <- keys$pathogen[i]
  cm <- keys$community_name[i]
  df <- child_pathogen_comm %>% filter(pathogen == p, community_name == cm)
  
  n_children <- n_distinct(df$childid)
  total_events <- sum(df$events, na.rm = TRUE)
  total_py     <- sum(df$py,     na.rm = TRUE)
  
  # Observed IR for this stratum
  ir_obs <- est_ir(df$events, df$py)
  
  # skip stratum if not enough available info
  if (n_children == 0 || is.na(ir_obs) || total_py <= 0) {
    boot_results_pc[[i]] <- tibble(
      pathogen = p, community_name = cm,
      n_children, total_events, total_py,
      ir = ir_obs, ir_per_100 = 100 * ir_obs,
      ir_lo = NA_real_, ir_hi = NA_real_,
      ir_lo_per_100 = NA_real_, ir_hi_per_100 = NA_real_
    )
    next  
  }
  
  # Matrix of resampled child IDs (clusters)
  idlist <- unique(df$childid)
  n_ids  <- length(idlist)
  boot_samp <- matrix(
    sample(idlist, size = n_ids * nboot, replace = TRUE),
    nrow = n_ids, ncol = nboot
  )
  
  ir_boots <- foreach(b = 1:nboot, .combine = c) %do% {
    di <- left_join(
      tibble(childid = boot_samp[, b]),
      df, by = "childid" 
    )
    est_ir(di$events, di$py)
  }
  
  # Precompute quantiles, then per-100
  lo <- as.numeric(quantile(ir_boots, 0.025, na.rm = TRUE))
  hi <- as.numeric(quantile(ir_boots, 0.975, na.rm = TRUE))


  boot_results_pc[[i]] <- tibble(
    pathogen = p, community_name = cm,
    n_children, total_events, total_py,
    ir = ir_obs, ir_per_100 = 100 * ir_obs,
    ir_lo = lo, ir_hi = hi,
    ir_lo_per_100 = 100 * lo,
    ir_hi_per_100 = 100 * hi
  )
}


#examine distribution
hist(ir_boots, breaks = 30)

boot_summary_pc <- bind_rows(boot_results_pc) %>%
  arrange(pathogen, community_name) %>% 
  left_join(data %>% # manually add back in community_type labels
              select(pathogen, community_name, 
                     community_type), 
            by = c("pathogen", "community_name")) %>% unique()

```

## Seroreversion rate calculation

Daily and annually

```{r}
sero_rev_rates <- sero_rev_tbl2 %>%
  group_by(pathogen) %>%
  summarise(
    total_seroreversions = sum(seroreversion, na.rm = TRUE),
    total_person_time_days = sum(person_time_days, na.rm = TRUE),  # includes person-time for those not at risk? 
    total_person_time_years = sum(person_time_years, na.rm = TRUE),
    
    reversion_rate_per_day = total_seroreversions / total_person_time_days,
    reversion_rate_per_year = total_seroreversions / total_person_time_years,
    
    rev_per_100_child_years = 100 * reversion_rate_per_year
    
  ) %>%
  ungroup()

# check how many seroreversions there are (at least one)
result <- sero_rev_tbl2 %>% 
  arrange(childid, pathogen, timepoint) %>%
  group_by(childid, pathogen) %>%
  summarise(
    pattern = paste(seroreversion, collapse = "")
  ) %>%
  filter(grepl("1", pattern)) %>%   # 1, then one or more 0s, then 1
  #select(childid, pathogen) %>%
  unique() %>% as.data.frame() %>% rename(childid = 1)

print("Number of seroreversions (childid-pathogen level)")
length(unique(result$childid))

print("Number of unique children with a seroreversion for at least 1 pathogen")
result %>% select(childid) %>% unique() %>% nrow()

# check how many seroreversions there are (at least one)
result2 <- sero_rev_tbl2 %>% 
  arrange(childid, pathogen, timepoint) %>%
  group_by(childid, pathogen) %>%
  summarise(
    pattern = paste(seroreversion, collapse = "")
  ) %>%
  filter(grepl("10+1", pattern)) %>%   # 1, then one or more 0s, then 1
  #select(childid, pathogen) %>%
  unique() %>% as.data.frame() %>% rename(childid = 1)

print("Top pathogens for multiple seroreversions")
result2 %>% select(pathogen) %>% 
  group_by(pathogen) %>% 
  summarise(n_children = n()) %>% 
  arrange(desc(n_children))

sero_rev_by_path <- sero_rev_rates %>%
  left_join(boot_summary, by = "pathogen") %>%
  select(pathogen, total_seroreversions, total_person_time_years, ir_obs_per_year = ir_obs, ir_lo_per_year = ir_lo, ir_hi_per_year = ir_hi, ir_obs_per_100, ir_lo_per_100, ir_hi_per_100)
```

## Seroreversion table, community level

```{r}
#------------------------------------
# export seroreversion table, 
# seroincidence rates CI csv & rds
#------------------------------------

sys.date <- format(Sys.Date(), "%Y%m%d")

saveRDS(sero_rev_tbl2, file = here("output/ecomid_sero_rev_tbl.rds"))
write.csv(sero_rev_tbl2, file = here("output/ecomid_sero_rev_tbl.csv"))


write.csv(sero_rev_by_path, file = here("output/ecomid_pathogen_sero_rev.csv"))
saveRDS(sero_rev_by_path, file = here("output/ecomid_pathogen_sero_rev.rds"))


write.csv(boot_summary_pc, file = here("output/ecomid_pathogen_comm_sero_rev.csv"))
saveRDS(boot_summary_pc, file = here("output/ecomid_pathogen_comm_sero_rev.rds"))

```

## Community levels (Esmeraldas, Borbon, Rural)


For ease of analysis in downstream, repeat the serocreversion, prevalence and bootstrapping for different grouping levels.

### Recode community leves

```{r}
# create higher level community variable 
data2 <- data %>% mutate(community_type2 = case_when(
      community_name == "Esmeraldas" ~ "Esmeraldas",
      community_name == "Borbon" ~ "Borbon",
      community_name %in% c("Maldonado", "Timbire", "Selva Alegre", "Colon Eloy",
                            "Zancudo", "Santo Domingo", "Colon de Onzole", "San Francisco") ~ "Rural"))
```

### Bootstrap sero-reversion per pathogen & community levels

Repeat steps as before but with new aggregated level

```{r}
child_comm2 <- data2 %>%
  select(childid, community_type2) %>%
  distinct()

child_pathogen_comm2 <- sero_rev_tbl2 %>%
  left_join(child_comm2, by = "childid") %>%
  group_by(pathogen, community_type2, childid) %>%
  summarise(
    # events per child (cap at 1 if you only want first-ever seroreversion)
    events = sum(seroreversion, na.rm = TRUE),
    # person-time in YEARS
    py     = sum(person_time_years, na.rm = TRUE),
    .groups = "drop_last"
  )
   

# 3) Bootstrap (child-cluster) within each pathogen × community
set.seed(234897)
nboot <- 1000
keys <- child_pathogen_comm2 %>%
  distinct(pathogen, community_type2) %>%
  arrange(pathogen, community_type2)

boot_results_pc <- vector("list", nrow(keys))

for (i in seq_len(nrow(keys))) {
  p  <- keys$pathogen[i]
  cm <- keys$community_type2[i]
  df <- child_pathogen_comm2 %>% filter(pathogen == p, community_type2 == cm)
  
  n_children <- n_distinct(df$childid)
  total_events <- sum(df$events, na.rm = TRUE)
  total_py     <- sum(df$py,     na.rm = TRUE)
  
  # Observed IR for this stratum
  ir_obs <- est_ir(df$events, df$py)
  
  # skip stratum if not enough available info
  if (n_children == 0 || is.na(ir_obs) || total_py <= 0) {
    boot_results_pc[[i]] <- tibble(
      pathogen = p, community_type2 = cm,
      n_children, total_events, total_py,
      ir = ir_obs, ir_per_100 = 100 * ir_obs,
      ir_lo = NA_real_, ir_hi = NA_real_,
      ir_lo_per_100 = NA_real_, ir_hi_per_100 = NA_real_
    )
    next  
  }
  
  # Matrix of resampled child IDs (clusters)
  idlist <- unique(df$childid)
  n_ids  <- length(idlist)
  boot_samp <- matrix(
    sample(idlist, size = n_ids * nboot, replace = TRUE),
    nrow = n_ids, ncol = nboot
  )
  
  ir_boots <- foreach(b = 1:nboot, .combine = c) %do% {
    di <- left_join(
      tibble(childid = boot_samp[, b]),
      df, by = "childid" 
    )
    est_ir(di$events, di$py)
  }
  
  # Precompute quantiles, then per-100
  lo <- as.numeric(quantile(ir_boots, 0.025, na.rm = TRUE))
  hi <- as.numeric(quantile(ir_boots, 0.975, na.rm = TRUE))


  boot_results_pc[[i]] <- tibble(
    pathogen = p, community_type2 = cm,
    n_children, total_events, total_py,
    ir = ir_obs, ir_per_100 = 100 * ir_obs,
    ir_lo = lo, ir_hi = hi,
    ir_lo_per_100 = 100 * lo,
    ir_hi_per_100 = 100 * hi
  )
}


# examine distribution
hist(ir_boots, breaks = 30)

boot_summary_pc <- bind_rows(boot_results_pc) %>%
  arrange(pathogen, community_type2)

# export 
write.csv(boot_summary_pc, file = here("output/ecomid_pathogen_comm_sero_rev_3_levels.csv"))
saveRDS(boot_summary_pc, file = here("output/ecomid_pathogen_comm_sero_rev_3_levels.rds"))
```

## Age-based seroreversion (under 12 months, over 12 months)
```{r}
child_agegroup <- data2 %>%
  select(childid, community_type2, timepoint) %>%
  distinct() %>% 
  mutate(age_group = ifelse(timepoint < 12, "6-9 months", "12-24 months"))

child_pathogen_age <- sero_rev_tbl2 %>%
  left_join(child_agegroup, by = c("childid", "timepoint")) %>%
  group_by(pathogen, age_group, childid) %>%
  summarise(
    # events per child (cap at 1 if you only want first-ever seroreversion)
    events = sum(seroreversion, na.rm = TRUE),
    # person-time in YEARS
    py     = sum(person_time_years, na.rm = TRUE),
    .groups = "drop_last"
  )
   

# 3) Bootstrap (child-cluster) within each pathogen × community
set.seed(234897)
nboot <- 1000
keys <- child_pathogen_age %>%
  distinct(pathogen, age_group) %>%
  arrange(pathogen, age_group)

boot_results_pc <- vector("list", nrow(keys))

for (i in seq_len(nrow(keys))) {
  p  <- keys$pathogen[i]
  ag <- keys$age_group[i]
  df <- child_pathogen_age %>% filter(pathogen == p, age_group == ag)
  
  n_children <- n_distinct(df$childid)
  total_events <- sum(df$events, na.rm = TRUE)
  total_py     <- sum(df$py,     na.rm = TRUE)
  
  # Observed IR for this stratum
  ir_obs <- est_ir(df$events, df$py)
  
  # skip stratum if not enough available info
  if (n_children == 0 || is.na(ir_obs) || total_py <= 0) {
    boot_results_pc[[i]] <- tibble(
      pathogen = p, age_group = ag,
      n_children, total_events, total_py,
      ir = ir_obs, ir_per_100 = 100 * ir_obs,
      ir_lo = NA_real_, ir_hi = NA_real_,
      ir_lo_per_100 = NA_real_, ir_hi_per_100 = NA_real_
    )
    next  
  }
  
  # Matrix of resampled child IDs (clusters)
  idlist <- unique(df$childid)
  n_ids  <- length(idlist)
  boot_samp <- matrix(
    sample(idlist, size = n_ids * nboot, replace = TRUE),
    nrow = n_ids, ncol = nboot
  )
  
  ir_boots <- foreach(b = 1:nboot, .combine = c) %do% {
    di <- left_join(
      tibble(childid = boot_samp[, b]),
      df, by = "childid" 
    )
    est_ir(di$events, di$py)
  }
  
  # Precompute quantiles, then per-100
  lo <- as.numeric(quantile(ir_boots, 0.025, na.rm = TRUE))
  hi <- as.numeric(quantile(ir_boots, 0.975, na.rm = TRUE))


  boot_results_pc[[i]] <- tibble(
    pathogen = p, age_group = ag,
    n_children, total_events, total_py,
    ir = ir_obs, ir_per_100 = 100 * ir_obs,
    ir_lo = lo, ir_hi = hi,
    ir_lo_per_100 = 100 * lo,
    ir_hi_per_100 = 100 * hi
  )
}


# examine distribution
hist(ir_boots, breaks = 30)

boot_summary_pc <- bind_rows(boot_results_pc) %>%
  arrange(pathogen, age_group)

# export 
write.csv(boot_summary_pc, file = here("output/ecomid_pathogen_agegroup_sero_rev.csv"))
saveRDS(boot_summary_pc, file = here("output/ecomid_pathogen_agegroup_sero_rev.rds"))

```




