---
title: "ECoMiD: Cumulative incidence summary datasets (PCR & IgG)"
subtitle: "Data-generation script -- Objective 1 of the PCR-vs-serology cleanup"
author: "Nikolina Walas (nikolina.walas@ucsf.edu)"
date: "updated: `r Sys.time()`"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## File overview

This is the **data-generation** script in the cleaned PCR-vs-serology
workflow. It reads the raw PCR and IgG data, builds the pathogen-level
point prevalence and Kaplan-Meier cumulative incidence datasets used
throughout the rest of this workflow, and exports them to `output/` so
that downstream visualization scripts do not need to re-derive them from
raw data (per the project convention that scripts don't source each
other directly, but pass data forward via files written to `output/`).

This file was updated on 2026/08/12 to restrict the PCR cohort to the same 370 children 
in the serology dataset. 

**Run this script first.** Two other cleaned documents in this workflow
depend on its exports:

-   `12-ecomid_dotplot_correlations.qmd` (Objective 3) uses the
    community-level, 24-month PCR/IgG export for the seroprevalence-
    vs-PCR-cumulative-incidence dot plot, and the community-stratified
    PCR and IgG Kaplan-Meier round-level exports for the PCR-cum.-inc.-
    vs-IgG-cum.-inc. dot plot (panel G).
-   `17b-ecomid_age_based_visualization.qmd` (Objective 4) uses the
    pathogen-level point prevalence and KM exports for the final
    4-measure summary figure and the KM sensitivity plots.

`9-ecomid_sero_pcr_summary.qmd` (Objective 2) does **not** depend on
this file -- it reads its own pre-computed population-level RDS files
(outputs of upstream scripts 2 and 3), independent of the KM work done
here.

This file generates cumulative incidence datasets at two levels:

1.  **Pathogen x assay** (PCR and IgG), pooled across all community
    types -- Kaplan-Meier cumulative incidence, with left-truncation/
    censoring diagnostics, an internal-missing-as-event upper-bound
    sensitivity analysis, and a Shigella/EIEC 6-month IgG exclusion
    sensitivity comparison.
2.  **Pathogen x community_type2** (Esmeraldas, Borbón, Rural) --
    Kaplan-Meier cumulative incidence for both PCR and IgG, matching
    the community-stratified build already in use for the correlation
    dot plots. IgG is also still separately summarized at this level
    as simple point seroprevalence, kept for panel E's existing
    seroprevalence-vs-PCR-cumulative-incidence comparison; the new
    IgG KM build instead feeds panel G's PCR-cumulative-incidence-vs-
    IgG-cumulative-incidence comparison.

### Specifications

-   The nominal reporting grid is now **assay-specific**
    (`pcr_rounds` / `igg_rounds`) instead of one shared `km_rounds`
    vector. The old vector also had an erroneous `21`-month entry
    that didn't correspond to any real visit round; that's gone. IgG
    additionally has no 3- or 15-month round, so those are no longer
    reported for IgG (they were previously just the prior round's
    estimate held forward, which implied IgG granularity that doesn't
    exist).
-   Added `km_rounds_override` + `rounds_for()`: a lookup that returns
    a pathogen/assay's exact reporting rounds, defaulting to the
    assay's normal grid unless a pathogen/assay pair has its own
    entry. Currently it holds one entry: **S. enterica PCR is only
    reported at 6, 12, and 18 months** (no real PCR testing at 3, 9,
    15, or 24 months for that pathogen), applied to both the
    pathogen-level and community-level PCR exports. Add more entries
    here if other pathogen/assay combinations turn out to have the
    same kind of gap.

## System setup

```{r}
library(here)
source(here("R", "0-config.R"))
library(survival)
library(broom)
```

## Read in raw data & build the pooled long dataset

```{r}
## ------------------------------------------------------------------------
## Point & cumulative prevalence, PCR vs serology, at the pathogen level
## ECoMiD -- restricted to the 9 pathogens assayed by both platforms
## ------------------------------------------------------------------------
## Inputs (edit paths as needed):
##   - PCR:      ecomid_pcr_analysis_data_2026-06-11.rds
##                 one row per child x visit x pathogen, already pathogen-level
##                 (pcrpos = 0/1/NA)
##   - Serology: ecomid_sero_antigen_seropos_long_2026-05-19.rds
##                 one row per child x timepoint x pathogen x ANTIGEN
##                 (seropos = 0/1/NA) -- some pathogens have >1 antigen
##                 (Campylobacter, Cryptosporidium, Norovirus GII, Salmonella)
## ------------------------------------------------------------------------

pcr_raw  <- readRDS(here("data/ecomid_pcr_analysis_data_2026-08-12.rds"))

sero_raw <- readRDS(here("../../data/serology/final/ecomid_serology_datasets/ecomid_sero_antigen_seropos_long_2026-08-26.rds")) %>%
  filter(antigen != "hlye")

# check childid alignment
length(unique(pcr_raw$childid))
length(unique(sero_raw$childid))
all(pcr_raw$childid %in% sero_raw$childid)

# The 9 pathogens with both a PCR target and a serology antigen
overlap_pathogens <- c(
  "Campylobacter", "Cryptosporidium", "E. histolytica", "Giardia",
  "LT-ETEC", "Norovirus GI", "Norovirus GII", "S. enterica", "Shigella/EIEC"
)

## ------------------------------------------------------------------------
## 1. PCR -- already at the pathogen level (one row per child/visit/pathogen,
##    regardless of whether tac or gpp was the assay run on that sample)
## ------------------------------------------------------------------------
pcr_long <- pcr_raw %>%
  filter(pathogen %in% overlap_pathogens) %>%
  transmute(
    childid,
    timepoint = as.numeric(visit),
    pathogen,
    assay  = "PCR",
    result = pcrpos
  )

## ------------------------------------------------------------------------
## 2. Serology -- collapse antigen -> pathogen: positive if positive to AT
##    LEAST ONE antigen for that pathogen (hlye excluded for Salmonella);
##    NA only if every antigen for that pathogen/child/timepoint is missing
##    (i.e. genuinely untested)
## ------------------------------------------------------------------------
any_pos <- function(x) {
  if (all(is.na(x))) return(NA_real_)
  as.numeric(any(x == 1, na.rm = TRUE))
}

sero_long <- sero_raw %>%
  filter(pathogen %in% overlap_pathogens) %>%
  group_by(childid, timepoint, pathogen) %>%
  summarise(result = any_pos(seropos), .groups = "drop") %>%
  mutate(
    timepoint = as.numeric(timepoint),
    assay     = "IgG"
  ) %>%
  select(childid, timepoint, pathogen, assay, result)

## ------------------------------------------------------------------------
## 3. Combine
## ------------------------------------------------------------------------
combined <- bind_rows(pcr_long, sero_long) %>%
  arrange(pathogen, assay, childid, timepoint)

## ------------------------------------------------------------------------
## Shared constants reused throughout this file
## ------------------------------------------------------------------------
## PCR and IgG were not collected on the same visit schedule (IgG has no
## 3- or 15-month round), so each assay's KM step function is read off
## only at rounds it actually has -- see km_at_rounds() below.
pcr_rounds      <- c(3, 6, 9, 12, 15, 18, 24)
igg_rounds      <- c(6, 9, 12, 18, 24)
age_axis_breaks <- scale_x_continuous(breaks = c(3, 6, 9, 12, 15, 18, 24))

## Pathogen x assay exceptions to the normal reporting grid above -- some
## pathogen/assay pairs were only actually tested at a subset of their
## assay's normal rounds. Currently: S. enterica PCR only has real data at
## 6, 12, and 18 months (no 3-, 9-, 15-, or 24-month PCR testing for that
## pathogen), so it gets its own explicit round set rather than pcr_rounds.
## NOTE: key must match the pathogen label used in overlap_pathogens/the
## dataframe (currently "S. enterica", not "Salmonella") or the override
## silently fails to match and this falls back to the full pcr_rounds grid.
## Add more entries here (keyed "pathogen.assay") if other pairs need it.
km_rounds_override <- list(
  "S. enterica.PCR" = c(6, 12, 18)
)

## Resolves the reporting grid for a given pathogen x assay: the override
## above if one exists, otherwise the assay's normal grid.
rounds_for <- function(pathogen, assay) {
  key <- paste(pathogen, assay, sep = ".")
  if (!is.null(km_rounds_override[[key]])) return(km_rounds_override[[key]])
  if (assay == "PCR") pcr_rounds else igg_rounds
}
```

## Point prevalence (pathogen x assay x round)

```{r}
## ------------------------------------------------------------------------
## Point prevalence: positives / tested, per pathogen x assay x timepoint
## CI: prop.test() per pathogen x assay x timepoint group -- each child
## contributes exactly one row to a given timepoint's tally here, so
## there's no repeated-measures/clustering to adjust for at this stage.
## ------------------------------------------------------------------------
point_prevalence <- combined %>%
  group_by(pathogen, assay, timepoint) %>%
  summarise(
    n_tested   = sum(!is.na(result)),
    n_pos      = sum(result == 1, na.rm = TRUE),
    point_prev = n_pos / n_tested,
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(ci_low  = prop.test(n_pos, n_tested)$conf.int[1],
         ci_high = prop.test(n_pos, n_tested)$conf.int[2]) %>%
  ungroup()

point_prevalence %>% arrange(pathogen, assay, timepoint) %>% print(n = 30)
```

## Naive cumulative prevalence (proportion-ever-positive; reference only)

Kept for reference and exported below, but this is **not** the
"cumulative incidence" measure plotted elsewhere in this workflow -- that
uses the Kaplan-Meier estimator built in the next section, which
properly accounts for left-truncated entry and loss to follow-up rather
than treating every "ever tested" child identically regardless of how
much of the study they were actually observed for.

```{r}
## ------------------------------------------------------------------------
## Assumption: a child's cumulative status carries forward across rounds
## (a positive at an earlier visit still counts as positive later), and a
## round with no recorded result for that child/pathogen just doesn't add
## information -- it neither resets nor advances their status.
##
## Exclusion: the 6-month Shigella/EIEC IgG readout is recoded to NA here
## (this pipeline only) to account for maternal antibody contribution.
## This has to happen BEFORE cum_pos/cum_tested are computed via cummax(),
## not as a filter afterward -- cummax() carries a positive forward to
## every later timepoint for that child, so filtering the 6-month row out
## only after the fact would hide the point at 6 months without removing
## its influence on 9-month-plus cumulative estimates.
## ------------------------------------------------------------------------
combined_cumulative <- combined %>%
  mutate(result = if_else(
    pathogen == "Shigella/EIEC" & assay == "IgG" & timepoint == 6,
    NA_real_, result
  ))

cumulative_prevalence <- combined_cumulative %>%
  arrange(childid, pathogen, assay, timepoint) %>%
  group_by(childid, pathogen, assay) %>%
  mutate(
    cum_pos    = cummax(replace_na(result, 0)),
    cum_tested = cummax(as.numeric(!is.na(result)))
  ) %>%
  ungroup() %>%
  filter(cum_tested == 1) %>%
  group_by(pathogen, assay, timepoint) %>%
  summarise(
    n_ever_tested = n(),
    n_ever_pos    = sum(cum_pos),
    cum_prev      = n_ever_pos / n_ever_tested,
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(ci_low  = prop.test(n_ever_pos, n_ever_tested)$conf.int[1],
         ci_high = prop.test(n_ever_pos, n_ever_tested)$conf.int[2]) %>%
  ungroup()

cumulative_prevalence %>% arrange(pathogen, assay, timepoint) %>% print(n = 30)
```

## Kaplan-Meier: time to first positive, pathogen x assay level

Point and naive cumulative prevalence, as calculated above, describe
each round as a cross-section and are sensitive to exactly which
children were tested that round. Because children enter the cohort at
different ages, miss individual rounds without dropping out, and in some
cases are lost to follow-up before the study ends, we instead estimate
time to first positive (PCR or IgG) using the Kaplan-Meier estimator,
treating each child's observed testing history as a single
entry-to-event or entry-to-censoring interval rather than a series of
independent rounds.

**Missingness handling rules** (per CLAUDE.md conventions):

-   leading NA (before the first observed round) -> left-truncation;
    entry = first observed round, nothing assumed before it
-   internal NA (an observed round follows later) -> carried forward /
    skipped over; never itself fires an event or a censor
-   trailing NA run (nothing observed after it, and no prior positive)
    -> right-censored at the FIRST missed round
-   positive anywhere from entry through the last observed round,
    including AT entry (baseline positive) -> event at that round
-   data ends on an observed (negative) round, no trailing NA ->
    administrative right-censoring at that last observed round
-   never tested at all (100% NA for that pathogen/assay) -> dropped;
    no entry time to anchor on

Because carrying an internal missing round forward as negative is
itself an assumption, we also fit a sensitivity version in which
internal missing rounds are instead treated as the event itself, giving
an upper bound on cumulative incidence. Trailing missing rounds are
still censored in this sensitivity version.

### Build (tstart, tstop, event) intervals

```{r}
EPS <- 1e-4  # tiny nudge so a same-round entry+event doesn't create a
             # zero-length (tstart, tstop] interval, which Surv() rejects

code_child_interval <- function(timepoint, result, na_as_event = FALSE) {

  ord       <- order(timepoint)
  timepoint <- timepoint[ord]
  result    <- result[ord]

  obs <- which(!is.na(result))
  if (length(obs) == 0) {
    return(tibble(tstart = numeric(0), tstop = numeric(0),
                  event = numeric(0), entry = numeric(0)))
  }

  first_obs  <- obs[1]
  last_obs   <- max(obs)
  entry_time <- timepoint[first_obs]

  window   <- first_obs:last_obs

  is_event <- if (na_as_event) {
    result[window] == 1 | is.na(result[window])
  } else {
    result[window] == 1
  }
  fire <- which(is_event)

  if (length(fire) > 0) {
    tstop <- timepoint[window][fire[1]]
    event <- 1
  } else if (last_obs < length(timepoint)) {
    tstop <- timepoint[last_obs + 1]
    event <- 0
  } else {
    tstop <- timepoint[last_obs]
    event <- 0
  }

  tstart <- entry_time
  if (tstop <= tstart) tstop <- tstart + EPS

  tibble(tstart = tstart, tstop = tstop, event = event, entry = entry_time)
}

## complete() forces the full child x round grid, per pathogen x assay, so
## code_child_interval() always sees the true trailing-NA run out to the
## last round anyone in the cohort was seen for that pathogen/assay.
combined_complete <- combined_cumulative %>%
  group_by(pathogen, assay) %>%
  complete(childid, timepoint) %>%
  ungroup()

## Sensitivity comparator: identical pipeline, but WITHOUT the Shigella/
## EIEC 6-month IgG exclusion, so we can see how much that one recoding
## decision actually moves the curve.
combined_complete_unexcluded <- combined %>%
  group_by(pathogen, assay) %>%
  complete(childid, timepoint) %>%
  ungroup()

build_km_intervals <- function(data, na_as_event = FALSE) {
  data %>%
    arrange(childid, pathogen, assay, timepoint) %>%
    group_by(childid, pathogen, assay) %>%
    group_modify(~ code_child_interval(.x$timepoint, .x$result, na_as_event)) %>%
    ungroup()
}

km_data_main            <- build_km_intervals(combined_complete,            na_as_event = FALSE)
km_data_upper           <- build_km_intervals(combined_complete,            na_as_event = TRUE)
km_data_main_unexcluded <- build_km_intervals(combined_complete_unexcluded, na_as_event = FALSE)
```

### Diagnostic: entry/event/censor counts per pathogen x assay

```{r}
km_diagnostic_counts <- km_data_main %>%
  group_by(pathogen, assay) %>%
  summarise(
    n_children  = n(),
    n_event     = sum(event == 1),
    n_censored  = sum(event == 0),
    .groups = "drop"
  ) %>%
  arrange(pathogen, assay)

km_diagnostic_counts %>% print(n = 30)
```

### Fit Kaplan-Meier curves by pathogen x assay

```{r}
fit_km_by_pathogen <- function(km_data) {
  km_data %>%
    group_by(pathogen) %>%
    group_modify(~ broom::tidy(survfit(
      Surv(tstart, tstop, event) ~ assay, data = .x,
      error     = "greenwood",
      conf.type = "log"
    ))) %>%
    ungroup() %>%
    mutate(assay = sub("^assay=", "", strata))
}

km_main_tidy <- fit_km_by_pathogen(km_data_main) %>%
  mutate(
    cum_inc = 1 - estimate,
    ci_low  = 1 - conf.high,
    ci_high = 1 - conf.low,
    bound   = "Main (internal NA = negative)"
  )

km_upper_tidy <- fit_km_by_pathogen(km_data_upper) %>%
  mutate(
    cum_inc = 1 - estimate,
    bound   = "Upper bound (internal NA = event)"
  )

## Sensitivity comparison for the Shigella/EIEC 6-month IgG exclusion --
## only expected to visibly differ from km_main_tidy for the Shigella/EIEC
## IgG panel; any divergence elsewhere would flag the exclusion leaking
## somewhere it shouldn't.
km_unexcluded_tidy <- fit_km_by_pathogen(km_data_main_unexcluded) %>%
  mutate(
    cum_inc = 1 - estimate,
    bound   = "Sensitivity: Shigella 6-month IgG included"
  )
```

### Extract cumulative incidence & n_at_risk at each nominal visit round

`broom::tidy(survfit(...))` only returns a row where an event or a
censor actually happened. That's fine for the cumulative incidence
estimate/CI (holding the last known value forward is a correct read of
the step function), but NOT for `n.risk`, which also only updates at
event/censor rows. `n_at_risk` is instead computed directly from each
child's (tstart, tstop] interval.

```{r}
km_at_rounds <- function(df, rounds) {
  df <- df %>% mutate(time = round(time, 3)) %>% arrange(time)

  grid <- tibble(time = rounds, cum_inc = NA_real_, ci_low = NA_real_,
                 ci_high = NA_real_, .is_round = TRUE)
  actual <- df %>%
    select(time, cum_inc, ci_low, ci_high) %>%
    mutate(.is_round = FALSE)

  bind_rows(actual, grid) %>%
    arrange(time, .is_round) %>%
    fill(cum_inc, ci_low, ci_high, .direction = "down") %>%
    filter(.is_round) %>%
    mutate(
      cum_inc = replace_na(cum_inc, 0),
      ci_low  = replace_na(ci_low, 0),
      ci_high = replace_na(ci_high, 0)
    ) %>%
    select(time, cum_inc, ci_low, ci_high)
}

n_at_risk_by_round <- function(df, rounds) {
  tibble(
    time      = rounds,
    n_at_risk = sapply(rounds, function(r) sum(df$tstart <= r & r <= df$tstop))
  )
}

## Each pathogen x assay group is read off its own reporting grid via
## rounds_for() (assay default, or the pathogen-specific override) rather
## than one shared set of rounds. group_modify()'s .y carries the grouping
## key (here: pathogen, assay) for the current group, since .x itself
## drops the grouping columns.
km_main_rounds_est <- km_main_tidy %>%
  group_by(pathogen, assay) %>%
  group_modify(~ km_at_rounds(.x, rounds = rounds_for(.y$pathogen, .y$assay))) %>%
  ungroup()

km_main_n_at_risk <- km_data_main %>%
  group_by(pathogen, assay) %>%
  group_modify(~ n_at_risk_by_round(.x, rounds = rounds_for(.y$pathogen, .y$assay))) %>%
  ungroup()

## Trim each pathogen x assay line to start at its own first entry round,
## rather than plotting/exporting every nominal round from 3 months on.
km_first_entry <- km_data_main %>%
  group_by(pathogen, assay) %>%
  summarise(first_entry = min(tstart), .groups = "drop")

km_main_rounds <- km_main_rounds_est %>%
  left_join(km_main_n_at_risk, by = c("pathogen", "assay", "time")) %>%
  left_join(km_first_entry, by = c("pathogen", "assay")) %>%
  filter(time >= first_entry) %>%
  select(-first_entry)

km_main_rounds %>% arrange(pathogen, assay, time) %>% print(n = 30)
```

### Wide export: cumulative incidence by assay, one row per pathogen x round

```{r}
km_by_assay_wide <- km_main_rounds %>%
  select(pathogen, assay, time, cum_inc, ci_low, ci_high, n_at_risk) %>%
  pivot_wider(
    names_from  = assay,
    values_from = c(cum_inc, ci_low, ci_high, n_at_risk),
    names_glue  = "{.value}_{assay}"
  ) %>%
  arrange(pathogen, time)

km_by_assay_wide %>% print(n = 30)
```

## Kaplan-Meier: PCR cumulative incidence by community type (pathogen x community_type2)

PCR, following the community-stratified build already validated for the
correlation dot plots. A parallel IgG Kaplan-Meier build follows in the
next section (added to support panel G's PCR-cum.-inc.-vs-IgG-cum.-inc.
correlation in `12-ecomid_dotplot_correlations.qmd`; previously deferred
as a larger piece of new analysis). IgG is additionally still summarized
at the community level as simple point seroprevalence further below,
which panel E continues to use.

```{r}
## community_type lives in sero_raw, at the child level -- pulled in here
## just for this section. community_type2 collapses "Rural Road"/"Rural
## River" into "Rural"; everything else (including "Borbon") passes
## through unchanged.
child_community <- sero_raw %>%
  distinct(childid, community_type) %>%
  mutate(community_type2 = case_when(
    community_type == "Rural Road"  ~ "Rural",
    community_type == "Rural River" ~ "Rural",
    TRUE ~ community_type
  )) %>%
  select(childid, community_type2)

community_n <- child_community %>%
  count(community_type2, name = "n_children")

pcr_by_comm <- pcr_long %>%
  inner_join(child_community, by = "childid") %>%
  group_by(pathogen, community_type2) %>%
  complete(childid, timepoint) %>%
  ungroup()

km_data_by_comm <- pcr_by_comm %>%
  arrange(childid, pathogen, community_type2, timepoint) %>%
  group_by(childid, pathogen, community_type2) %>%
  group_modify(~ code_child_interval(.x$timepoint, .x$result, na_as_event = FALSE)) %>%
  ungroup()
```

### Diagnostic: does every pathogen have all 3 community types?

```{r}
## If this shows only 1-2 rows for any pathogen instead of 3, that
## pathogen's PCR data doesn't actually span all community types, which is
## a real data question, not something to paper over.
km_data_by_comm %>% distinct(pathogen, community_type2) %>% count(pathogen) %>% print(n = 30)
```

### Fit & extract PCR community-level KM at each nominal round

```{r}
## broom::tidy(survfit(...)) only includes a `strata` column when more than
## one community_type2 level is actually present in the data being fit --
## reconstruct it from the single level present if needed.
km_tidy_by_comm <- km_data_by_comm %>%
  group_by(pathogen) %>%
  group_modify(~ {
    tidied <- broom::tidy(survfit(
      Surv(tstart, tstop, event) ~ community_type2, data = .x,
      error = "greenwood", conf.type = "log"
    ))
    if (!"strata" %in% names(tidied)) {
      tidied$strata <- paste0("community_type2=", unique(.x$community_type2))
    }
    tidied
  }) %>%
  ungroup() %>%
  mutate(
    community_type2 = sub("^community_type2=", "", strata),
    cum_inc         = 1 - estimate,
    ci_low          = 1 - conf.high,
    ci_high         = 1 - conf.low
  )

## Reuse the same km_at_rounds()/n_at_risk_by_round() helpers defined above,
## just grouped by community_type2 instead of assay -- this is what makes
## the community-level export "time based" (every nominal round) rather
## than a single 24-month snapshot. This section is PCR only, so rounds_for()
## is always called with assay = "PCR" -- it still applies the S. enterica
## override (6/12/18 only), same as the pooled export above.
km_pcr_by_comm_rounds_est <- km_tidy_by_comm %>%
  group_by(pathogen, community_type2) %>%
  group_modify(~ km_at_rounds(.x, rounds = rounds_for(.y$pathogen, "PCR"))) %>%
  ungroup()

km_pcr_by_comm_n_at_risk <- km_data_by_comm %>%
  group_by(pathogen, community_type2) %>%
  group_modify(~ n_at_risk_by_round(.x, rounds = rounds_for(.y$pathogen, "PCR"))) %>%
  ungroup()

km_pcr_by_comm_first_entry <- km_data_by_comm %>%
  group_by(pathogen, community_type2) %>%
  summarise(first_entry = min(tstart), .groups = "drop")

km_pcr_by_comm_rounds <- km_pcr_by_comm_rounds_est %>%
  left_join(km_pcr_by_comm_n_at_risk, by = c("pathogen", "community_type2", "time")) %>%
  left_join(km_pcr_by_comm_first_entry, by = c("pathogen", "community_type2")) %>%
  filter(time >= first_entry) %>%
  select(-first_entry)

km_pcr_by_comm_rounds %>% arrange(pathogen, community_type2, time) %>% print(n = 30)
```

## Kaplan-Meier: IgG cumulative incidence by community type (pathogen x community_type2)

Mirrors the PCR community-stratified build directly above -- same
`code_child_interval()` / `survfit(Surv(tstart, tstop, event) ~
community_type2)` machinery, just sourced from IgG instead of PCR.
Sourced from `combined_cumulative` (not the raw `sero_long`/`combined`)
filtered to `assay == "IgG"`, so the Shigella/EIEC 6-month IgG exclusion
carries through here exactly as it does for the pathogen-level IgG KM
above -- using `combined`/`sero_long` directly would silently reintroduce
that excluded timepoint into this community-level build.

```{r}
igg_by_comm <- combined_cumulative %>%
  filter(assay == "IgG") %>%
  inner_join(child_community, by = "childid") %>%
  group_by(pathogen, community_type2) %>%
  complete(childid, timepoint) %>%
  ungroup()

km_data_igg_by_comm <- igg_by_comm %>%
  arrange(childid, pathogen, community_type2, timepoint) %>%
  group_by(childid, pathogen, community_type2) %>%
  group_modify(~ code_child_interval(.x$timepoint, .x$result, na_as_event = FALSE)) %>%
  ungroup()
```

### Diagnostic: does every pathogen have all 3 community types (IgG)?

```{r}
km_data_igg_by_comm %>% distinct(pathogen, community_type2) %>% count(pathogen) %>% print(n = 30)
```

### Fit & extract IgG community-level KM at each nominal round

```{r}
## Same strata-reconstruction fallback as the PCR fit above.
km_tidy_igg_by_comm <- km_data_igg_by_comm %>%
  group_by(pathogen) %>%
  group_modify(~ {
    tidied <- broom::tidy(survfit(
      Surv(tstart, tstop, event) ~ community_type2, data = .x,
      error = "greenwood", conf.type = "log"
    ))
    if (!"strata" %in% names(tidied)) {
      tidied$strata <- paste0("community_type2=", unique(.x$community_type2))
    }
    tidied
  }) %>%
  ungroup() %>%
  mutate(
    community_type2 = sub("^community_type2=", "", strata),
    cum_inc         = 1 - estimate,
    ci_low          = 1 - conf.high,
    ci_high         = 1 - conf.low
  )

## rounds_for() is called with assay = "IgG" here, so this reads off the
## igg_rounds grid (no 3- or 15-month round) rather than pcr_rounds.
km_igg_by_comm_rounds_est <- km_tidy_igg_by_comm %>%
  group_by(pathogen, community_type2) %>%
  group_modify(~ km_at_rounds(.x, rounds = rounds_for(.y$pathogen, "IgG"))) %>%
  ungroup()

km_igg_by_comm_n_at_risk <- km_data_igg_by_comm %>%
  group_by(pathogen, community_type2) %>%
  group_modify(~ n_at_risk_by_round(.x, rounds = rounds_for(.y$pathogen, "IgG"))) %>%
  ungroup()

km_igg_by_comm_first_entry <- km_data_igg_by_comm %>%
  group_by(pathogen, community_type2) %>%
  summarise(first_entry = min(tstart), .groups = "drop")

km_igg_by_comm_rounds <- km_igg_by_comm_rounds_est %>%
  left_join(km_igg_by_comm_n_at_risk, by = c("pathogen", "community_type2", "time")) %>%
  left_join(km_igg_by_comm_first_entry, by = c("pathogen", "community_type2")) %>%
  filter(time >= first_entry) %>%
  select(-first_entry)

km_igg_by_comm_rounds %>% arrange(pathogen, community_type2, time) %>% print(n = 30)
```

## IgG point seroprevalence by community type (pathogen x community_type2 x round)

Not Kaplan-Meier -- simple point prevalence (positives / tested) at each
round, per pathogen x community type. Paired with the PCR KM export above
for panel E's community-level correlation work in
`12-ecomid_dotplot_correlations.qmd`. (Panel G instead uses the IgG KM
build in the section above.)

```{r}
igg_seroprev_by_comm_rounds <- combined %>%
  filter(assay == "IgG") %>%
  inner_join(child_community, by = "childid") %>%
  group_by(pathogen, community_type2, timepoint) %>%
  summarise(
    n_pos          = sum(result == 1, na.rm = TRUE),
    n_tested       = sum(!is.na(result)),
    seroprevalence = n_pos / n_tested,
    .groups = "drop"
  ) %>%
  rename(time = timepoint)

igg_seroprev_by_comm_rounds %>% arrange(pathogen, community_type2, time) %>% print(n = 30)
```

## 24-month snapshot: PCR cumulative incidence vs. IgG seroprevalence, by community

Convenience export for the correlation dot plot in
`12-ecomid_dotplot_correlations.qmd` (Objective 3, plot E) -- reads the
fitted step function's value at (or just before) 24 months for PCR, and
IgG point seroprevalence at exactly 24 months, both by pathogen x
community type. (Plot G's PCR-cum.-inc.-vs-IgG-cum.-inc. snapshot is
built directly in that file from the two round-level KM exports below,
the same way plot F already reads the PCR one -- no separate snapshot
export is needed here for G.)

```{r}
pcr_cuminc_24_by_comm <- km_tidy_by_comm %>%
  filter(time <= 24) %>%
  group_by(pathogen, community_type2) %>%
  slice_max(time, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(pathogen, community_type2, pcr_cum_inc = cum_inc)

igg_seroprev_24_by_comm <- igg_seroprev_by_comm_rounds %>%
  filter(time == 24) %>%
  transmute(pathogen, community_type2, seroprevalence)

pcr_igg_24mo_by_community <- inner_join(
  igg_seroprev_24_by_comm, pcr_cuminc_24_by_comm,
  by = c("pathogen", "community_type2")
) %>%
  left_join(community_n, by = "community_type2")

pcr_igg_24mo_by_community %>% print(n = 30)
```

## Export datasets

```{r}
## Pathogen-level (both assays)
write_csv(point_prevalence,      here("output/ecomid_point_prevalence_pcr_vs_serology.csv"))
write_csv(cumulative_prevalence, here("output/ecomid_cumulative_prevalence_pcr_vs_serology.csv"))
write_csv(km_by_assay_wide,      here("output/ecomid_km_cumulative_incidence_pcr_vs_igg_by_round.csv"))

saveRDS(km_main_tidy,       here("output/ecomid_km_main_tidy.rds"))
saveRDS(km_upper_tidy,      here("output/ecomid_km_upper_tidy.rds"))
saveRDS(km_unexcluded_tidy, here("output/ecomid_km_unexcluded_tidy.rds"))
saveRDS(km_main_rounds,     here("output/ecomid_km_main_rounds.rds"))

## Community-level (PCR KM + IgG KM + IgG point prevalence)
saveRDS(km_pcr_by_comm_rounds,        here("output/ecomid_km_pcr_by_community_round.rds"))
## Per-child PCR survival intervals (tstart, tstop, event), by community --
## exported so 10b-ecomid_dotplot_correlations.qmd's panel B can bootstrap
## by resampling children and refitting Kaplan-Meier per replicate, without
## rebuilding intervals from raw testing rounds on every one of 2000
## replicates (the slow, non-vectorized group_modify() step above). A
## child's own interval doesn't depend on which other children land in a
## given bootstrap resample, so it's exported once here rather than
## recomputed downstream.
saveRDS(km_data_by_comm,              here("output/ecomid_km_intervals_pcr_by_community.rds"))
saveRDS(km_igg_by_comm_rounds,        here("output/ecomid_km_igg_by_community_round.rds"))
saveRDS(igg_seroprev_by_comm_rounds,  here("output/ecomid_igg_seroprevalence_by_community_round.rds"))
saveRDS(pcr_igg_24mo_by_community,    here("output/ecomid_pcr_igg_24mo_by_community.rds"))

## Diagnostics (kept for review, not consumed downstream)
write_csv(km_diagnostic_counts, here("output/ecomid_km_diagnostic_counts_pathogen_assay.csv"))
```
