---
title: "ECoMiD Serology, threshold assignment figure"
date: "updated: `r Sys.time()`"
author: "Nikolina Walas (nikoewalas@gmail.com"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
---

# File description:

The purpose of this file is to generate the supplemental figure of MFI thresholds, as determined for the enteric serology antigen panel.

Thresholds were determined by generating a pre-exposed endemic cohort as outlined in the steps below:

1\) Identify children who had a 100-fold increase in MFI between two timepoints

2\) Take the serology observation at the first timepoint before the 100-fold MFI increase

3\) Take the mean + 3SD of that population

Thresholds for 4 antigen targets were determined by the CDC (Cp17, Cp23, LecA and VSP-5) using ROC curves. We identified the threshold using the pre-exposed endemic cohort, but used the ROC thresholds for final analysis.

Importantly, SalB target did not have enough data to generate a pre-exposed endemic cohort, and therefore the threshold was determined using the mean + 3\*SD of the MFI values on the log10 scale for all samples

## System setup

```{r}
# Read in files 
library(here)
source(here("R", "0-config.R"))

# Read in data
d1 <- read_rds(here("data/ecomid_sero_antigen_seropos_enteric_cutoffs_plotting_2026-05-19.rds"))

d1 <- d1 %>% mutate()
#--------------------------
# custom log labels
#--------------------------
log10labs4 <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)
```

## Stacked barplot

```{r}
#--------------------------------#
# Stacked barplot
#--------------------------------#

# add in variable for the two age groups
d1 <- d1 %>%
  mutate(time_grp = if_else(timepoint <= 12, "6-12 months", "18-24 months"))
d1$time_grp <- factor(d1$time_grp, levels = c("18-24 months", "6-12 months"))

# plot
pstack_time <- ggplot(d1, aes(x = log10mfi, fill = time_grp)) +
  facet_wrap(~antigen_lab, nrow = 3, ncol = 5) +
  geom_histogram(
    binwidth = 0.2,
    position = "stack",
    color = NA,
    alpha = 0.8
  )  +
  geom_vline(aes(xintercept = log10(mficut), color = "ROC", linetype = "ROC")) +
  geom_vline(aes(xintercept = log10(pecut), color = "pre-exposed endemic cohort", linetype = "pre-exposed endemic cohort")) +
  scale_x_continuous(limits = c(0, 4.5), breaks = 0:4, labels = log10labs4) +
  labs(
    x = "Luminex IgG (MFI-bg)",
    y = "Sample count",
    fill = NULL,
    color = "",
    linetype = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c(
    "6-12 months" = "deepskyblue4",
    "18-24 months" = "lightblue"
  ),
  breaks = c("6-12 months", "18-24 months")) +
  scale_color_manual(values = c(
    "ROC" = "black",
    "pre-exposed endemic cohort" = "black"
  )) +
  scale_linetype_manual(values = c(
    "ROC" = "solid",
    "pre-exposed endemic cohort" = "dashed"
  )) 

pstack_time

ggsave(filename = here("figures","plot_igg_dists_stackedbar.tiff"), pstack_time, device="tiff",width = 210, height = 150, units = "mm")
```
